### 2020-05-15 free-theorems-nesteds/issues-with-fold.tex
writeup of fold examples / bush reverse investigation

% For double-blind review submission, w/o CCS and ACM Reference (max
% submission space)
\documentclass[acmsmall,review,anonymous]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2020}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations
%\citestyle{acmnumeric}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8]{inputenc}
\usepackage{ccicons}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amscd}
%\usepackage{MnSymbol}
\usepackage{xcolor}

\usepackage{bbold}
\usepackage{url}
\usepackage{upgreek}
%\usepackage{stmaryrd}

\usepackage{lipsum}
\usepackage{tikz-cd}
\usetikzlibrary{cd}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}

\usepackage{bussproofs}
\EnableBpAbbreviations

\input{macros}
\input{mytheorem}

\theoremstyle{definition}
\newtheorem{exmpl}{Example}

\renewcommand{\greyout}[1]{} %{{\color{gray} {#1}}} -- toggle to remove greyed text

\newcommand{\emptyfun}{{[]}}
\newcommand{\cal}{\mathcal}
%\newcommand{\fold}{\mathit{fold}}
\newcommand{\F}{\mathcal{F}}
\renewcommand{\G}{\mathcal{G}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\P}{\mathcal{A}}
\newcommand{\pred}{\mathsf{Fam}}
\newcommand{\env}{\mathsf{Env}}
\newcommand{\set}{\mathsf{Set}}
\renewcommand{\S}{\mathcal S}
\renewcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\A}{\mathcal{A}}
\renewcommand{\id}{\mathit{id}}
\newcommand{\map}{\mathsf{map}}
\newcommand{\pid}{\underline{\mathit{id}}}
\newcommand{\pcirc}{\,\underline{\circ}\,}
\newcommand{\pzero}{\underline{0}}
\newcommand{\pone}{\underline{1}}
\newcommand{\psum}{\,\underline{+}\,}
%\newcommand{\inl}{\mathit{inL}\,}
%\newcommand{\inr}{\mathit{inR}\,}
\newcommand{\pinl}{\underline{\mathit{inL}}\,}
\newcommand{\pinr}{\underline{\mathit{inR}}\,}
\newcommand{\ptimes}{\,\underline{\times}\,}
\newcommand{\ppi}{\underline{\pi_1}}
\newcommand{\pppi}{\underline{\pi_2}}
\newcommand{\pmu}{\underline{\mu}}

\title[Free Theorems for Nested Types]{Free Theorems for 
Nested Types} %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Patricia Johann and Andrew Polonsky}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{Department1}              %% \department is recommended
  \institution{Appalachian State University}            %% \institution is required
%  \streetaddress{Street1 Address1}
%  \city{City1}
%  \state{State1}
%  \postcode{Post-Code1}
%  \country{Country1}                    %% \country is recommended
}
\email{johannp@appstate.edu, andrew.polonsky@gmail.com}          %% \email is recommended

\begin{document}

\begin{abstract}
\end{abstract}

\maketitle
%% sstart 
\section{Limitations of folds and issues defining \texttt{bushreverse}}

We would like to define a function \texttt{bushreverse} which reverses a bush. 
Bushes are not linear, so it is not immediately clear what reversing a bush should even mean. 
First consider the flattening of bushes as a natural transformation from Bush to List (see bush2list). 
The flattening of a reversed bush should be the reversal (on lists) of the flattening of the original bush, 
i.e., \texttt{forall b. bush2list (bushreverse b) = reverse (bush2list b)} 
or \texttt{bush2list $\circ$ bushreverse = reverse $\circ$ bush2list}.
We would also like \texttt{bushreverse} to be an involution on bushes.

Clearly this function will have a recursive definition, so to define it in our calculus we will use a fold. 
Before defining \texttt{bushreverse} as a fold, let's consider how to define list reversal as a fold. There are several ways to do this. 
One method is to replace nil with nil and replace cons with a function which inserts an element at the back of a list.
This cons replacement function can be defined in several ways, e.g.,

  \texttt{$\lambda \; x \; xs \to$ xs ++ [ x ]}, or

  \texttt{$\lambda \; x \to$ foldr (::) [ x ] }.

The first version makes use of the append function on lists. 
The second version simply replaces the nil constructor with a singleton list. 
The first cannot be defined in our system because we cannot define an append function for bushes (see next section). 
The second cannot be defined because the fold for bushes requires a fully polymorphic algebra, 
so the only possible replacement for bnil is bnil in this context. Writing the insertion function with foldr is
much simpler than the case for bushes because there is only one nil constructor in a list, but a bush
can have many occurences of bnil.
We can also give a recursive definition without using append or foldr:
\begin{align*}
  &\texttt{insert : $\forall$ \{A : Set\} $\to$ A $\to$ List A $\to$ List A} \\
  &\texttt{insert x [] = [ x ] } \\
  &\texttt{insert x (y :: ys) = y :: (insert x xs) } 
\end{align*}

and we can define an analogous function for bushes:
\begin{align*}
  &\texttt{binsert : $\forall$ \{A : Set\} $\to$ A $\to$ Bush (Bush A) $\to$ Bush A} \\
  &\texttt{binsert x bnil = bcons x bnil } \\
  &\texttt{binsert x (bcons bnil bbbx) = bcons x (binsert bnil bbbx)} \\
  &\texttt{binsert x (bcons (bcons y bx) bbbx) = bcons y (binsert (binsert x bx) bbbx)}
\end{align*}
and feed this as an argument to the fold for bushes to get a reversal function: 
\begin{align*}
  &\texttt{bushreverse : $\forall$ \{A : Set\} $\to$ Bush A $\to$ Bush A} \\
  &\texttt{bushreverse \{A\} = bfold \{A\} \{Bush\} bnil binsert}
\end{align*}

Note that the first explicit argument to bfold is completely determined by the types. 
The argument must have type \texttt{($\forall$ \{A : Set\} $\to$ Bush A)} and the only inhabitant of this type is bnil.
Folds for (non first-order) nested types will require arguments of the form \texttt{($\forall$ \{A : Set\} $\to$ F A)} whenever 
the type has constructors with no arguments, and due to the universal quantification, the only inhabitants of this type 
will be these constructors themselves (e.g. pnil, bnil) with some other possible combinations 
if there are constructors of type \texttt{($\forall$ \{A : Set\} $\to$ F A $\to$ F A $\to$ F A)}, for example.

This restriction means that the behavior of our reversal function must be specified entirely by binsert. 
The question is, can we define binsert without arbitrary term level recursion? 
If we try to define it as a fold over the \texttt{Bush (Bush A)} parameter, then we have no 
way to include x in that (fully polymorphic) fold. We could even try to define it as a fold over a 
bush built up from x such as \texttt{(bcons x bnil)}, but the same issue arises.

The fundamental issue here is that the binsert function takes two arguments, 
and there is no way to combine them together with a single fold. We can fold
over each argument separately, but then we still have results to combine.

\subsection{What can we define with folds over nested types?}
These issues are not specific to the Bush datatype; they are inherent limitations of folds
over nested types. In fact we are quite restricted in the types of functions we can define recursively. 
To define a function using a fold, its domain must be a fixed point, therefore
types such as \texttt{($\forall$ \{A : Set\} $\to$ (Bush A $\times$ Bush A) $\to$ Bush A)} cannot be defined as
a fold over the pair. Of course, we can fold over each bush in the pair individually, but we cannot interleave
the computations in any way because the polymorphism required by the folds prohibits this.

Additionally, the
return type of the result of the fold must be a functor. So even if we curry the previous type to get
\texttt{($\forall$ \{A : Set\} $\to$ Bush A $\to$ (Bush A $\to$ Bush A))}, we can not define it as a fold
over the first bush because \texttt{(Bush A $\to$ Bush A)} is not a functor.
These two restrictions indicate that we cannot define many interesting functions which have
multiple nested types as arguments.

The restriction stretches even further though, because sometimes the algebras supplied to a fold
must have a recursive definition involving nested types. This is the case with \texttt{bushreverse}
because even though it has a suitable type, \texttt{($\forall$ \{A : Set\} $\to$ Bush A $\to$ Bush A)},
the binsert function has a type which makes it incompatible with fold.

\subsection{Generalized Folds}
We can make our calculus slightly more expressive using generalized folds \cite{?}.
However, these generalized folds are mostly helpful for defining \textit{reductions}, 
functions which take data of a monomorphic type as input and have a monomorphic return type. 
Specifically, generalized folds relax the constraint on the domain of the function being defined as a fold.
Instead of requiring the domain to be a fixed point, the domain can be a fixed point composed with another functor. 
However, this does not help us in defining functions which have multiple nested types as arguments.

\end{document}