### 2021-04-15 induction-paper/ENTCS-GADT-induction.tex
revised sections 1 through 4, added abstract and related work

\documentclass[9pt]{entcs}
\usepackage{entcsmacro}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{ccicons}
\usepackage{verbatim}
\usepackage{amscd}
\usepackage{MnSymbol}
\usepackage{xcolor}
\usepackage{bbold}
\usepackage{url}
\usepackage{upgreek}
\usepackage{stmaryrd}
\usepackage{lipsum}
\usepackage{tikz-cd}
\usetikzlibrary{cd}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usepackage{bussproofs}
\EnableBpAbbreviations

\sloppy
% The following is enclosed to allow easy detection of differences in
% ascii coding.
% Upper-case    A B C D E F G H I J K L M N O P Q R S T U V W X Y Z
% Lower-case    a b c d e f g h i j k l m n o p q r s t u v w x y z
% Digits        0 1 2 3 4 5 6 7 8 9
% Exclamation   !           Double quote "          Hash (number) #
% Dollar        $           Percent      %          Ampersand     &
% Acute accent  '           Left paren   (          Right paren   )
% Asterisk      *           Plus         +          Comma         ,
% Minus         -           Point        .          Solidus       /
% Colon         :           Semicolon    ;          Less than     <
% Equals        =3D           Greater than >          Question mark ?
% At            @           Left bracket [          Backslash     \
% Right bracket ]           Circumflex   ^          Underscore    _
% Grave accent  `           Left brace   {          Vertical bar  |
% Right brace   }           Tilde        ~

\bibliographystyle{entcs}

\def\lastname{Johann, Ghiorzi, and Jeffries}

\input{macros}

\begin{document}
\begin{frontmatter}
  \vspace*{-0.1in}
  \title{(Deep) Induction for GADTs\vspace*{-0.1in}}
  \author{Patricia Johann~~~~}%\thanksref{JohannEmail}}
  \author{Enrico Ghiorzi~~~~}%\thanksref{GhiorziEmail}}
  \author{Daniel Jeffries}%\thanksref{JeffriesEmail}}
%  \thanks[JohannEmail]{Email: \href{mailto:johannp@appstate.edu} {\texttt{\normalshape johannp@appstate.edu}}}
%  \thanks[GhiorziEmail]{Email: \href{mailto:ghiorzie@appstate.edu} {\texttt{\normalshape ghiorzie@appstate.edu}}}
%  \thanks[JeffriesEmail]{Email: \href{mailto:jeffriesd@appstate.edu} {\texttt{\normalshape jeffriesd@appstate.edu}}}
  \address{$\mathtt{\{johannp,ghiorzie,jeffriesd\}@appstate.edu}$\\
    Appalachian State University}

\vspace*{-0.1in}

  \begin{abstract} 
Deep data types are data types that are defined in terms of other such
data types, including, in the case of truly nested types, themselves.
Deep induction is an extension of structural induction that traverses
{\em all} of the structure present in a structure of such a type,
propagating suitable predicates to {\em all} of the data contained in
that structure. Deep induction has been shown to be the form of
induction most suitable for applications involving deep nested
types. In this paper we show how to extend deep induction to a robust
class of deep GADTs that are not truly nested. We also show that it
cannot be extended to truly nested GADTs.
\end{abstract}

%\begin{keyword}
%  Please list keywords from your paper here, separated by commas.
%\end{keyword}

\end{frontmatter}

\vspace*{-0.1in}

\section{Introduction}\label{sec:intro}

%\vspace*{-0.09in}

Induction is one of the most important techniques available for
working with advanced data types, so it is both inevitable and
unsurprising that it plays an essential role in modern proof
assistants. In the proof assistant Coq~\cite{coq20}, for example,
functions and predicates over advanced types are defined inductively,
and almost all non-trivial proofs of their properties are either
proved by induction outright or rely on lemmas that are. Every time a
new inductive data type is declared in Coq, an induction rule is
automatically generated for it.

The data types handled by Coq are (possibly mutually inductive)
polynomial ADTs, and the induction rules automatically generated for
them are the expected ones for standard structural induction. It has
long been understood, however, that these rules are too weak to be
genuinely useful for so-called {\em deep ADTs}~\cite{jp20}, i.e., ADTs
that are (possibly mutually inductively) defined in terms of (other)
such ADTs.\footnote{Such data types are called nested inductive types
  by Chlipala~\cite{chl}, reflecting the fact that ``inductive type''
  means ``ADT''~in~Coq.} Consider, for example, the following type of
rose trees, here coded in Agda and defined in terms of the standard
type of lists:

\vspace*{-0.075in}

\[\begin{array}{l}
\mathsf{data\, Rose\, : Set \to Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;empty\, :\, Rose\,A}\\
\mathsf{\;\;\;\;\;\;\;\;node\,\,\,\, :\, A \to List\,(Rose\,A) \to Rose\,A} 
\end{array}\]
\noindent
The induction rule Coq automatically generates for rose trees is

\vspace*{-0.075in}

\[\begin{array}{l}
\mathsf{\forall\, (a : Set)\,(P : Rose\,a \to Set) \to P\,empty \to}\\
\mathsf{\hspace*{0.15in}
 (\forall\, (x : a)\,(ts :
  List\,(Rose\,a)) \to P\,(node\,x\,ts)) \to 
  \forall \,(x : Rose\,a) \to P\,x}
\end{array}\]
\noindent
Unfortunately, this is neither the induction rule we intuitively
expect, nor is it expressive enough to prove even basic properties of
rose trees that ought to be amenable to inductive proof. What is
needed here is an enhanced notion of induction that, when specialized
to rose trees, will propagate the predicate $\mathsf{P}$ through the
outer list structure and to the rose trees sitting inside
$\mathsf{node}$'s list argument. More generally, this enhanced notion
of induction should traverse {\em all} of the structure present in a
data element, propagating suitable predicates to {\em all} of the data
contained in the structure. With data types becoming ever more
advanced, and with deeply structured such types becoming ever more
%increasingly
ubiquitous in formalizations, it is critically important that proof
assistants
%like Coq, Agda, and Lean 
be able to automatically generate genuinely useful induction
rules for data types that go well beyond traditional ADTs. Such
data types include (truly) nested types~\cite{bm98}\footnote{A truly
  nested type is a nested type that is defined over itself.},
generalized algebraic data types
(GADTs)~\cite{ch03,pvww06,sp04,xcc03}, more richly indexed
families~\cite{ch88}, and deep variants of all of these.

\pagebreak

{\em Deep induction}~\cite{jp20} is a generalization of structural
induction that fits this bill exactly. Whereas structural induction
rules induct over only the top-level structure of data, leaving any
data internal to the top-level structure untouched, deep induction
rules induct over {\em all} of the structured data present. The key
idea is to parameterize induction rules not just over a predicate over
the top-level data type being considered, but also over additional
custom predicates on the types of primitive data they contain. These
custom predicates are then lifted to predicates on any internal
structures containing these data, and the resulting predicates on
these internal structures are lifted to predicates on any internal
structures containing structures at the previous level, and so on,
until the internal structures at all levels of the data type
definition, including the top level, have been so
processed. Satisfaction of a predicate by the data at one level of a
structure is then conditioned upon satisfaction of the appropriate
predicates by {\em all} of the data at the preceding level.

Deep induction was shown in~\cite{jp20} to be the form of induction
most appropriate to nested types (including ADTs) that are defined
over, or mutually recursively with, other such types (including,
possibly, themselves). Deep induction delivers the following genuinely
useful induction rule for rose trees:

\vspace*{-0.075in}

\begin{equation}\label{eq:rose}
\begin{array}{l}
\mathsf{\forall \,(a : Set)\,(P : Rose\,a \to Set)\,(Q : a \to Set)
  \to P\,empty \to}\\ 
\mathsf{\hspace*{0.15in}(\forall \,(x : a)\, (ts :
  List\,(Rose\,a))\to Q\,x \to List^\land\,P\,ts \to P\,(node\,x\,ts)) \to}\\
\mathsf{\hspace*{0.15in}\forall \,(x :
  Rose\,a) \to Rose^\land\,Q\, x \to P\,x} 
\end{array}
\end{equation}

\noindent
%Here, $\mathsf{List^\land}$ lifts its predicate argument $\mathsf{P}$
%on data of type $\mathsf{Rose\,a}$ to a predicate on data of type
%$\mathsf{List\,(Rose\,a)}$ asserting that $\mathsf{P}$ holds for every
%element of its argument list, and similarly for the lifting
%$\mathsf{Rose^\land}$ and the predicate $\mathsf{Q}$.
Here, $\mathsf{List^\land}$ (resp., $\mathsf{Rose^\land}$) lifts its
predicate argument $\mathsf{P}$ (resp., $\mathsf{Q}$) on data of type
$\mathsf{Rose\,a}$ (resp., $\mathsf{a}$) to a predicate on data of
type $\mathsf{List\,(Rose\,a)}$ (resp., $\mathsf{Rose\,a}$) asserting
that $\mathsf{P}$ (resp., $\mathsf{Q}$) holds for every element of its
list (resp., rose tree) argument.\footnote{Predicate liftings such as
  $\mathsf{List^\land}$ and $\mathsf{Rose^\land}$ can either be
  supplied as primitives or generated automatically from their
  associated data type definitions as described in
  Section~\ref{sec:ADTs-and-nesteds} below. The predicate lifting for
  a container type like $\mathsf{List\,t}$ or $\mathsf{Rose\,t}$
  simply traverses containers of that type and applies its predicate
  argument pointwise to the constituent data of type $\mathsf{t}$.}
Deep induction was also shown in~\cite{jp20} to
% be the missing piece making it possible to
%be the theretofore missing component needed to
% be the essential ingredient making it possible to
deliver the first-ever induction rules --- structural or otherwise ---
for the $\mathsf{Bush}$ data type~\cite{bm98} and other truly nested
types. Deep induction for ADTs and nested types is reviewed in
Section~\ref{sec:ADTs-and-nesteds} below.

This paper shows how to extend deep induction to proper GADTs, i.e.,
to GADTs that are not simply nested types (and thus are not ADTs).  A
constructor for such a GADT $\mathsf{G}$ may, like a constructor for a
nested type, take as arguments data whose types involve instances of
$\mathsf{G}$ other than the one being defined --- including instances
that involve $\mathsf{G}$ itself. But if $\mathsf{G}$ is a proper GADT
then at least one of its constructors will also have such a structured
instance of $\mathsf{G}$ --- albeit one not involving $\mathsf{G}$
itself --- as its codomain.
%But while the types of the arguments to a GADT's constructor can
%include instances of $\mathsf{G}$ that themselves involve
%$\mathsf{G}$, its return type cannot.
For example, the constructor $\mathsf{pair}$ for the GADT {\color{red}
  Perhaps also show non-inhabitation?}

\vspace*{-0.1in}

\begin{equation}\label{eq:seq}
\begin{array}{l}
\mathsf{data\, Seq\,(a : Set) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;const :\, a \to Seq\,a}\\
\mathsf{\;\;\;\;\;\;\;\;\;\;\,pair :\, Seq \,a \to Seq\,b \to
  Seq\,(a \times b)}
\end{array}
\end{equation}
\noindent
of sequences only constructs sequences of pairs, rather than sequences
of arbitrary type, as does $\mathsf{const}$. If all of the
constructors for a GADT $\mathsf{G}$ return structured instances of
$\mathsf{G}$, then some of $\mathsf{G}$'s instances might not be
inhabited. GADTs therefore have two distinct, but equally natural,
semantics: a functorial semantics interpreting them as left Kan
extensions~\cite{mac71}, and a parametric semantics interpreting them
as their Church encodings~\cite{atk12,vw10}. As explained
in~\cite{jgj21}, a key difference in the two semantics is that the
former views GADTs as their {\em functorial completions}~\cite{jp19},
and thus as containing more data than just those expressible in
syntax. By contrast, the latter views them as what might be called
{\em syntax-only} GADTs. Happily, these two views of GADTs coincide
for those that are ADTs or other nested types. However, both they and
their attendant properties differ greatly for proper GADTs. In fact,
the views deriving from the functorial and parametric semantics for
proper GADTs are sufficiently distinct that, by contrast with the
situation for ADTs and other nested
types~\cite{bfss90,gjfor15,jgj21f}, it is not actually possible to
define a functorial parametric semantics for them~\cite{jgj21}.

This observation seems, at first, to be a death knell for the prospect
of extending deep induction to GADTs. Indeed, since induction can be
seen as unary parametricity, we quickly realize that GADTs viewed as
their functorial completions cannot possibly support induction rules.
This makes sense intuitively: induction is a syntactic proof
technique, so of course it cannot be used to prove properties of those
elements of a GADT's functorial completion that are not expressible in
syntax. All is not lost, however. As we show below, the Church
encoding interpretation's syntax-only view does support induction
rules --- including deep induction rules --- for GADTs. Perhaps
surprisingly, ours are the first-ever induction rules --- deep or
otherwise --- for a general class of proper GADTs. But this paper
actually delivers more: it gives a general framework for deriving deep
induction rules for a general class of deep GADTs directly from their
syntax. This framework can serve as a basis for extending modern proof
assistants' automatic generation of structural induction rules for
ADTs to automatic generation of deep induction rules for GADTs. As for
ADTs and other nested types, the structural induction rule for any
GADT can be recovered from its deep induction rule simply by taking
the custom predicates in its deep induction rule to be constantly
$\mathsf{True}$-valued predicates.

Deep induction rules for GADTs cannot, however, be derived by somehow
extending the techniques of~\cite{jp20} to syntax-only GADTs. Indeed,
the derivation of induction rules given there makes crucial use of the
functoriality of data types' interpretations from~\cite{jp19}, and
that is precisely what the interpretation of GADTs as their Church
encodings fails to deliver. Instead, we first give a predicate lifting
styled after those of~\cite{jp20}, together with a (deep) induction
rule, and for the simplest --- and arguably most important --- GADT,
namely the equality GADT. (See Section~\ref{sec:ind-equal}.) We can
then derive the deep induction rule for any other GADT $\mathsf{G}$ by
{\em i}) using the equality GADT to represent $\mathsf{G}$ as its
so-called {\em Henry Ford
  encoding}~\cite{ch03,hin03,mcb99,sjsv09,sp04}, and {\em ii}) using
the predicate liftings for the equality GADT and any other GADTs
appearing in the definition of $\mathsf{G}$ to appropriately thread
the custom predicates for the primitive types appearing in
$\mathsf{G}$ through its structure. This two-step process delivers
deep induction rules for a broad class of deep GADTs. In
Section~\ref{sec:GADTs} we introduce a series of increasingly complex
GADTs as running examples, and in Section~\ref{sec:deep-ind-GADTs} we
derive a deep induction rule for each of them. In particular, we
derive the deep induction rule for $\mathsf{Seq}$ in
Section~\ref{sec:ind-seq}. We present our general framework for
deriving (deep) induction rules for (deep) GADTs in
Section~\ref{sec:framework}, and observe that the derivations in
Section~\ref{sec:deep-ind-GADTs} are all instances of it.  In
Section~\ref{sec:GADT-nested} we show that, by contrast with truly
nested types, which do have a functorial semantics, syntax-only GADTs'
lack of functoriality means that it is not possible to extend
induction --- deep or otherwise --- to truly nested GADTs. This does
not appear to be much of a restriction, however, since GADTs defined
over themselves do not, to our knowledge, appear in applications or
the literature.

All of the deep induction rules appearing in this paper have been
derived using our general framework. Our Agda code implementing them
is available at~\cite{web-page}.

%\subsection{Related Work}

\vspace*{0.05in}

{\bf Related Work\/}
Various techniques for deriving induction rules for data types that go
beyond ADTs have been studied. For example, Fu and
Selinger~\cite{fs18} show, via examples, how to derive induction rules
for arbitrary nested types. Unfortunately, however, their technique is
rather {\em ad hoc}, so is unclear how to generalize it to nested
types other than the specific ones in the examples. Moreover, it
actually derives induction rules for data types {\em related} to the
original nested types rather than for the original nested types
themselves, and it is unclear whether or not the derived rules are
sufficiently expressive to prove all results about the original nested
types that we would expect to be provable by induction. This latter
point echoes the issue with Coq-derived induction rule for rose trees
raised in Section~\ref{sec:intro}, which has the unfortunate effect of
forcing users to manually write induction (and other) rules for such
types for use in that system. Tassi~\cite{tas19} has done exactly
that, deriving induction rules for data type definitions in Coq using
unary parametricity. Tassi's technique seems to be essentially
equivalent to that of~\cite{jp19} for nested types, although he does
not permit true nesting. More recently, Ulrich~\cite{ull20} has
implemented a plugin in MetaCoq to generate induction rules for nested
types. This plugin is also based on unary parametricity and, again,
true nesting is not permitted.  {\color{red} As far as we know, no
  attempt has (yet) been made to extend either implementation to
  GADTs.} In fact, we know of no work other than that reported here
that specifically addresses induction rules for (deep) GADTs.

\section{Deep induction for ADTs and nested types}\label{sec:ADTs-and-nesteds}

A structural induction rule for a data type allows us to prove that if
a predicate holds for every element inductively produced by the data
type's constructors then it holds for every element of the data type.
In this paper, we are interested in induction rules for proof-relevant
predicates.  A proof-relevant predicate on a type $\mathsf{A : Set}$
is a function $\mathsf{P\,:\,A \to Set}$ mapping each $\mathsf{a : A}$
to the set of proofs that $\mathsf{P\,a}$ holds.  For example, the
induction rule for the standard list type

\vspace*{-0.1in}

\begin{equation*}\label{eq:list}
\begin{array}{l}
\mathsf{data\ List : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;nil\,\,\,\,\,\, :\, List\,A}\\
\mathsf{\;\;\;\;\;\;\;cons\, :\, A \to List\,A \to List\,A} 
\end{array}
\end{equation*}

\vspace*{-0.1in}

\noindent
is

\vspace*{-0.2in}

\begin{equation*}
\mathsf{
\forall (A : Set) (P : List\,A \to Set)
\to P\,nil
\to \big( \forall (a : A) (as: List\,A)
\to P\,as
\to P\,(cons\,a\,as)\big)
\to \forall (as : List\,A)
\to P\, as
}
\end{equation*}

\vspace*{-0.1in}

\noindent
As in Coq's induction rule for rose trees, the data inside a
structure of type $\mathsf{List}$ is treated monolithically (i.e.,
ignored) by this structural induction rule.
By contrast, the deep induction rule for lists is parameterized over a
custom predicate $\mathsf{Q}$ on $\mathsf{A}$ as described in the
introduction. For $\mathsf{List^\wedge}$ as described in the
introduction it is

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (A : Set) (P : List\, A \to Set) (Q : A \to Set)
\to P\,Nil \to \big( \forall (a : A) (as: List\,A) \to Q\,a \to P\,as
\to P\,(Cons\,a\,as)\big)} \\ 
\quad\mathsf{\to \forall (as : List\,A) \to List^{\wedge}\,A\,Q\,as
  \to P\,as } 
\end{array}\]

\vspace*{-0.1in}

Structural induction can be extended to nested types, such as the
following type of perfect trees~\cite{bm98}:

\pagebreak

\begin{equation*}\label{eq:ptree}
\begin{array}{l}
\mathsf{data\ PTree : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;pleaf\,\,\, :\, A \to PTree\,A}\\
\mathsf{\;\;\;\;\;\;\;pnode\, :\, PTree\,(A \times A) \to PTree\,A} 
\end{array}
\end{equation*}
Perfect trees can be thought of as lists constrained to have lengths
that are powers of 2. In the above code, the constructor
$\mathsf{pnode}$ uses data of type $\mathsf{PTree\,(A \times A)}$ to
construct data of type $\mathsf{PTree\,A}$. Thus, it is clear that the
instances of $\mathsf{PTree}$ at various indices cannot be defined
independently, and that the entire inductive family of types must
therefore be defined at once. This intertwinedness of the instances of
nested types is reflected in their structural induction rules, which,
as explained in~\cite{jp20}, must necessarily involve polymorphic
predicates rather than the monomorphic predicates appearing in
structural induction rules for ADTs. The structural induction rule for
perfect trees, for example, is

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to PTree\, A \to Set)
\to \big( \forall (A : Set) (a : A) \to P\,A\,(pleaf\, a) \big)} \\
\quad\mathsf{\to \big( \forall (A : Set) (tt : PTree\,(A \times A))
  \to P\,(A \times A)\,tt \to P\,a\,(pnode\,tt)\big) \to \forall (A :
  Set) (t : PTree\,A) \to P\,A\,t } 
\end{array}\]

\vspace*{-0.1in}

\noindent
The deep induction rule for perfect trees similarly uses polymorphic
predicates but otherwise follows the now-familiar pattern:

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to PTree\,A
  \to Set) \to \big( \forall (A : Set) (Q : A \to Set) (a : A) \to
  Q\,a \to P\,A\,Q\,(Pleaf\, a) \big)} \\ \quad \mathsf{\to \big(
  \forall (A : Set) (Q : A \to Set) (tt : PTree\,(A \times A)) \to
  P\,(A \times A)\,(Pair^{\wedge}\,A\,A\,Q\,Q)\,tt \to
  P\,A\,Q\,(Pnode\,tt)\big)} \\ \quad \mathsf{\to \forall (A : Set) (Q
  : A \to Set) (t : PTree\,A) \to PTree^{\wedge}\,A\,Q\,t \to
  P\,A\,Q\,t }
\end{array}\]

\vspace*{-0.1in}

\noindent
Here, $\mathsf{Pair^{\wedge} : \forall (A\; B: Set) \to (A \to Set)
  \to (B \to Set) \to A \times B \to Set}$ lifts predicates
$\mathsf{Q_A}$ on data of type $\mathsf{A}$ and $\mathsf{Q_B}$ on data
of type $\mathsf{B}$ to a predicate on pairs of type $\mathsf{A \times
  B}$ in such a way that $\mathsf{Pair^{\wedge}\,A\,B\,Q_A\,Q_B\,(a,b)
 = Q_A\,a \times Q_B\,b}$. Similarly, $\mathsf{PTree^{\wedge} :
  \forall (A : Set) \to (A \to Set) \to PTree\,A \to Set}$ lifts a
predicate $\mathsf{Q}$ on data of type $\mathsf{A}$ to a predicate on
data of type $\mathsf{PTree\,A}$ asserting that $\mathsf{Q}$ holds for
every element of type $\mathsf{A}$ contained in its perfect tree
argument.

It is not possible to extend structural induction to {\em truly}
nested types, i.e., to nested types whose recursive occurrences appear
below themselves. The quintessential example of such a type is that of
bushes~\cite{bm98}:
\begin{equation*}\label{eq:bush}
\begin{array}{l}
\mathsf{data\ Bush : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;bnil\,\,\,\,\,\, :\, Bush\,A}\\
\mathsf{\;\;\;\;\;\;bcons\, :\, A \to Bush\,(Bush\,A) \to Bush\,A} 
\end{array}
\end{equation*}

\noindent
Even defining a structural induction rule for bushes requires that we
be able to lift the rule's polymorphic predicate argument to
$\mathsf{Bush}$ itself. The more general observation that an induction
rule for any truly nested type must therefore necessarily be a deep
induction rule was, in fact, the original motivation for the
development of deep induction in~\cite{jp20}. The deep induction rule
for bushes is
\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Bush\, A \to Set)
\to \big( \forall (A : Set) \to P\,A\,bnil \big)} \\ 
\quad\mathsf{\to \big( \forall (A : Set) (Q : A \to Set) (a : A) (bb :
  Bush\,(Bush\,A)) \to Q\,a \to
  P\,(Bush\,A)\,(Bush^{\wedge}\,A\,Q)\,bb \to P\,A\,Q\,(bcons\,a\,bb)
  \big)} \\ 
\quad\mathsf{\to \forall (A : Set) (Q : A \to Set) (b : Bush\,A) \to
  Bush^{\wedge}\,A\,Q\,b \to P\,A\,Q\,b } 
\end{array}\]
Here, $\mathsf{Bush^{\wedge} : \forall (A : Set) \to (A \to Set) \to
  Bush\,A \to Set}$ is the following lifting of a predicate
$\mathsf{Q}$ on data of type $\mathsf{A}$ to a predicate on data of
type $\mathsf{Bush\,A}$ asserting that $\mathsf{Q}$ holds for every
element of type $\mathsf{A}$ contained in its argument bush:
\begin{align*}
\mathsf{Bush^{\wedge}\,A\,Q\,bnil} &= \mathsf{\top} \\
\mathsf{Bush^{\wedge}\,A\,Q\,(bcons\,a\,bb)} &= \mathsf{Q\,a \times
  Bush^{\wedge}\,(Bush\,A)\,(Bush^{\wedge}\,A\,Q)\,bb}
\end{align*}

Although a truly nested type admits only a single induction rule, it
is worth noting that for those nested types that do admit distinct
structural induction and deep induction rules, the latter generalizes
the former. Indeed, the structural induction rule for such a nested
type is recoverable from its deep induction rule by taking the custom
predicates on its data of primitive types to be constantly
$\mathsf{True}$-valued predicates. This instantiation ensures that the
resulting induction rule only inspects the top-level structure of its
argument, rather than the contents of that structure, which exactly
coindices with what structural induction should do.
%A concrete example of such a derivation is given in
%Section~\ref{sec:ind-equal}.

\section{(Deep) GADTs}\label{sec:GADTs}

While a data constructor for a nested type can take {\em as arguments}
data whose types involve instances of that type at indices other than
the one being defined, its return type must still be at the (variable)
type instance being defined. For example, every data constructor for
$\mathsf{PTree\,A}$ must return an element of type
$\mathsf{PTree\,A}$, regardless of the instances of $\mathsf{PTree}$
appearing in the types of its arguments. GADTs relax this restriction,
allowing their data constructors both to take as arguments \emph{and
  return as results} data whose types involve instances of them other
than the one being defined. And as with the return type of
$\mathsf{pair}$ in~\eqref{eq:seq}, these instances can be structured.
%Applications of GADTs include generic programming, modeling
%programming languages via higher-order abstract syntax, maintaining
%invariants in data structures, and expressing constraints in embedded
%domain-specific languages. GADTs have also been used, e.g., to
%implement tagless interpreters~\cite{pl04,pr06,pvww06}, to improve
%memory performance~\cite{min15}, and to design APIs~\cite{pen20}.

GADTs are used in precisely those situations in which different
behaviors at different instances of data types are desired. This is
achieved by allowing the programmer to give the type signatures of the
GADT's data constructors independently, and then taking advantage of
pattern matching to force the desired type refinement. For example,
the {\em equality} GADT
\begin{equation}\label{eq:equal}
\begin{array}{l}
\mathsf{data\ Equal : Set \to Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;\;\;\;\,refl :\, Equal\,A\,A}
\end{array}
\end{equation}
is parametrized by two type indices, but it is only possible to
construct data elements of type $\mathsf{Equal\,a\,b}$ if $\mathsf{a}$
and $\mathsf{b}$ are instantiated at the same type. If the types
$\mathsf{a}$ and $\mathsf{b}$ are syntactically identical then the
type $\mathsf{Equal\,a\,b}$ contains the single data element
$\mathsf{refl}$. It contains no data elements otherwise.

The importance of the equality GADT lies in the fact that we can
understand other GADTs in terms of it. For example, the GADT
$\mathsf{Seq}$ from~\eqref{eq:seq} comprises constant sequences of
data of any type $\mathsf{A}$ and sequences obtained by pairing the
data in two already existing sequences. This GADT can be rewritten as
its Henry Ford encoding, which makes critical use of the equality
GADT, as follows:
\begin{equation}\label{eq:eq_seq}
\begin{array}{l}
\mathsf{data\ Seq : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;const :\, A \to Seq\,A}\\ 
\mathsf{\;\;\;\;\;\;\;\;pair\,\, :\, \forall (B\,C : Set) \to
  Equal\,A\,(B \times C) \to Seq\,B \to Seq\,C \to Seq\,A}\\ 
\end{array}
\end{equation}
Here, the requirement that $\mathsf{pair}$ produce data at an instance
of $\mathsf{Seq}$ that is a product type is replaced with the
requirement that $\mathsf{pair}$ produce data at an instance of
$\mathsf{Seq}$ that is \emph{equal} to a product type. As we will see
in Section~\ref{sec:deep-ind-GADTs}, the presence of the equality GADT is
key to deriving deep induction rules for GADTs.

Neither $\mathsf{Equal}$ nor $\mathsf{Seq}$ is a deep GADT, but the
following GADT $\mathsf{LTerm}$, which encodes terms of a simply typed
lambda calculus, is. More robust variations on $\mathsf{LTerm}$ are,
of course, possible. But since this variation is rich enough to
illustrate all essential aspects of deep GADTs --- and later, in
Section~\ref{sec:ind-lam}, their deep induction rules --- while still
being small enough to ensure clarity of exposition, we keep it to a
minimum.

Types are either booleans, arrow types, or list types. They are
represented by the Henry Ford GADT
\begin{equation}\label{eq:eq_ltype}
\begin{array}{l}
\mathsf{data\ LType : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;bool :\, \forall (B : Set) \to Equal\,A\,Bool
  \to LType\,A}\\ 
\mathsf{\;\;\;\;\;\;\;\;arr\,\,\, :\, \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to LType\,B \to LType\,C \to LType\,A}\\ 
  \mathsf{\;\;\;\;\;\;\;\;list\,\, :\, \forall (B : Set) \to
    Equal\,A\,(\List\,B) \to LType\,B \to LType\,A} 
\end{array}
\end{equation}
Terms are either variables, abstractions, applications, or lists of
terms. They are represented by
\begin{equation}\label{eq:eq_lterm}
\begin{array}{l}
\mathsf{data\ LTerm : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;var\,\,\,\,\,\,\,:\, String \to LType\,A \to
  LTerm\,A} \\  
\mathsf{\;\;\;\;\;\;\;\;abs\,\,\,\,\,\, :\, \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to String \to LType\,B \to LTerm\,C \to
  LTerm\,A}\\ 
  \mathsf{\;\;\;\;\;\;\;\;app\,\,\,\,\, :\, \forall (B : Set) \to
    LTerm (B \to A) \to LTerm\,B \to LTerm\,A} \\ 
  \mathsf{\;\;\;\;\;\;\;\;list\,\,\, :\, \forall (B : Set) \to
    Equal\,A\,(List\,B) \to List\,(LTerm\,B) \to LTerm\,A} 
\end{array}
\end{equation}
The type parameter for $\mathsf{LTerm}$ tracks the types of simply
typed lambda calculus terms. For example, $\mathsf{LTerm\,A}$ is the
type of simply typed lambda terms of type $\mathsf{A}$. Variables are
tagged with their types by the data constructors $\mathsf{var}$ and
$\mathsf{abs}$, whose $\mathsf{LType}$ arguments ensure that their
type tags are legal types. This ensures that all lambda terms
produced by $\mathsf{var}$, $\mathsf{abs}$, $\mathsf{app}$, and
$\mathsf{list}$ are well-typed.  We will revisit these GADTs in
Sections~\ref{sec:deep-ind-GADTs} and~\ref{sec:app}.

\section{(Deep) induction for GADTs}\label{sec:deep-ind-GADTs}

The equality constraints engendered by GADTs' data constructors makes
deriving (deep) induction rules for then more involved than for ADTs
and other nested types. Nevertheless, we show in this section how to
do so. We first illustrate the key components of our approach by
deriving deep induction rules for the three specific GADTs introduced
in Section~\ref{sec:GADTs}. Then, in Section~\ref{sec:framework}, we
abstract these to a general framework that can be applied to any deep
GADT that is not truly nested. As hinted above, the predicate lifting
for the equality GADT plays a central role in deriving both s
tructural and deep induction rules for more general GADTs.

\subsection{(Deep) induction for $\mathsf{Equal}$}\label{sec:ind-equal}

To define the (deep) induction rule for any (deep) GADT $\mathsf{G}$
we first need to define a predicate lifting that maps a predicate on a
type $\mathsf{A}$ and to a predicate on $\mathsf{G\,A}$. Such a
predicate lifting
$\mathsf{Equal^{\wedge} : \forall (A\,B : Set) \to (A \to Set) \to (B
  \to Set) \to Equal\,A\,B \to Set}$
for $\mathsf{Equal}$ is defined by
\[\mathsf{Equal^{\wedge}\,A\,A\,Q\,Q'\,refl = \forall (a : A) \to
  Equal\,(Q\,a)(Q'\,a)}\]
which takes two predicates on the same type as input and tests them
for extensional equality.

Next, we need to associate with each data constructor $\mathsf{c}$ of
$\mathsf{G}$ an {\em induction hypothesis} asserting that, if the
custom predicate arguments to a predicate $\mathsf{P}$ on $\mathsf{G}$
can be lifted to $\mathsf{G}$ itself, then $\mathsf{c}$ {\em respects}
$\mathsf{P}$, i.e., $\mathsf{c}$ constructs data satisfying the
instance of $\mathsf{P}$ at those custom predicates. The following
induction hypothesis $\mathsf{dIndRefl}$ is thus associated with the
$\mathsf{refl}$ constructor for $\mathsf{Equal}$:
\begin{equation*}\label{eq:ind-refl}
\begin{array}{l}
\mathsf{\lambda (P : \forall (A\,B : Set) \to (A \to Set) \to (B \to
  Set) \to Equal\,A\,B \to Set)} \\ 
\quad\mathsf{\to \forall (C : Set) (Q\, Q' : C \to Set) \to
  Equal^{\wedge}\,C\,C\,Q\,Q'\,refl \to P\,C\,C\,Q\,Q'\,refl} 
\end{array}
\end{equation*}

The deep induction rule for $\mathsf{G}$ now states that, if all of
$\mathsf{G}$'s data constructors respect a predicate $\mathsf{P}$,
then $\mathsf{P}$ is satisfied by every element of $\mathsf{G}$ to
which the custom predicate arguments to $\mathsf{P}$ can be
successfully lifted.  The deep induction rule for $\mathsf{Equal}$ is
thus
\begin{equation}\label{eq:ind-equal}
\begin{array}{l}
\mathsf{\forall (P : \forall (A\,B : Set) \to (A \to Set) \to (B \to
  Set) \to Equal\,A\,B \to Set) \to dIndRefl\,P \to}\\ \quad 
  \mathsf{\forall (A\,B : Set) (Q_A : A \to Set) (Q_B : B \to Set) (e:
  Equal\,A\,B) \to Equal^{\wedge}\,A\,B\,Q_A\,Q_B\,e \to
  P\,A\,B\,Q_A\,Q_B\,e}
\end{array}
\end{equation}
To prove that this rule is sound we must provide a witness
$\mathsf{dIndEqual}$ inhabiting the type in~\eqref{eq:ind-equal}.  By
pattern matching, we need only consider the case where $\mathsf{A} =
\mathsf{B}$ and $\mathsf{e} = \mathsf{refl}$, so we can define
$\mathsf{dIndEqual}$ by
$\mathsf{dIndEqual\;P\;crefl\;A\;A\;Q_A\;Q_A'\;refl\;liftE =
  crefl\;A\;Q_A\;Q_A'\,liftE}$. We can recover the structural
induction rule
\begin{equation}\label{eq:sind-equal}
\mathsf{\forall (Q : \forall (A\,B : Set)
  \to Equal\,A\,B \to Set) \to \big( \forall (C : Set) \to
  P\,C\,C\,refl \big) \to \forall (A\,B : Set) (e: Equal\,A\,B) \to
  P\,A\,B\,e}
\end{equation}
for $\mathsf{Equal}$ by defining a term $\mathsf{indEqual}$ of the
type in~\eqref{eq:sind-equal} by $\mathsf{indEqual\;Q\;srefl\;A\;B\;e
  = dIndEqual\;P\;srefl\;A\;B\;K^A_\top\;K^B_\top\;e\,liftE}$. Here,
$\mathsf{e : Equal\,A\,B}$, $\mathsf{P : \forall (A\,B : Set) \to (A
  \to Set) \to (B \to Set) \to Equal\,A\,B \to Set}$ is defined by
$\mathsf{P\;A\;B\;Q_A\;Q_B\;e =}$ $\mathsf{Q\;A\;B\;e}$,
$\mathsf{K^A_\top}$ and $\mathsf{K^B_\top}$ are the constantly
{\color{red} $\mathsf{\top}$}-valued predicates on $\mathsf{A}$ and
$\mathsf{B}$, respectively, and $\mathsf{liftE :}$
$\mathsf{Equal^{\wedge}\;A\;B\;K^A_\top\;K^B_\top\;e}$ is
%fully
defined
%(by pattern matching)
by $\mathsf{liftE\,a = refl : Equal\,A\,A}$ for every $\mathsf{a :
  A}$.  The structural induction rule for any GADT $\mathsf{G}$ that
is not truly nested can similarly be recovered from its deep induction
rule by instantiating every custom predicate by the appropriate
constantly {\color{red}$\mathsf{\top}$}-valued predicate.

\subsection{(Deep) induction for $\mathsf{Seq}$}\label{sec:ind-seq}

To derive the deep induction rule for the GADT $\mathsf{Seq}$ we use
its Henry Ford encoding from~\eqref{eq:eq_seq}.  We first define its
predicate lifting $\mathsf{Seq^\wedge : \forall (A : Set) \to (A \to
  Set) \to Seq\,A \to Set}$ by
\[\begin{array}{lll}
\mathsf{Seq^{\wedge}\,A\,Q_A\,(const\,a)} & = & \mathsf{Q_A\,a}\\
\mathsf{Seq^{\wedge}\,A\,Q_A\,(sPair\,B\,C\,e\,s_B\,s_C)}
&=&\mathsf{\exists [Q_B] \exists [Q_C]\, Equal^{\wedge}\,A\, (B
  \times C)\, Q_A\, (Q_B \times Q_C) \, e \times
  Seq^{\wedge}\,B\,Q_B\,s_B \times Seq^{\wedge}\,C\,Q_C\,s_C}
\end{array}\]
Here, $\mathsf{a : A}$, $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C : C
  \to Set}$, $\mathsf{e : Equal\,A\,(B \times C)}$, $\mathsf{s_B :
  Seq\,B}$, $\mathsf{s_C : Seq\,C}$, and $\mathsf{\exists [x]\, F
  \,x}$ is syntactic sugar for the type of dependent pairs
$\mathsf{(x,b)}$ where $\mathsf{x : A}$ and $\mathsf{b : F\, x}$ and
$\mathsf{F : A \to Set}$.

Next, let $\mathsf{dIndConst}$ be the induction hypothesis
\[\mathsf{
\lambda (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to Set) \to
\forall (A : Set) (Q_A : A \to Set) (a : A) \to Q_A\,x \to
P\,A\,Q_A\,(const\,a)}\] associated with the constructor
$\mathsf{const}$, and let $\mathsf{dIndPair}$ be the induction
hypothesis
\[\begin{array}{l}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \to \\ 
\quad \mathsf{\forall (A\,B\,C : Set) (Q_A : A \to Set) (Q_B : B
  \to Set) (Q_C : C \to Set)(s_B : Seq\,B) (s_C : Seq\,C) (e :
  Equal\,A\,(B \times C)) \to} \\ 
\quad \mathsf{Equal^{\wedge} A\, (B \times C)\, Q_A\,
  (Pair^{\wedge}\,B\,C\,Q_B\,Q_C)\, e \to P\,B\,Q_B\,s_B \to
  P\,C\,Q_C\,s_C \to P A Q_A ( pair\,B\,C\,e\,s_B\,s_C )}
\end{array}\]
associated with the constructor $\mathsf{pair}$. Then the deep
induction rule for $\mathsf{Seq}$ is
\begin{equation}\label{eq:ind-seq}
\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \mathsf{\to dIndConst\,P \to dIndPair\,P \to} \\ \quad
\mathsf{\forall (A : Set)(Q_A : A \to Set)(s_A : Seq\,A) \to
  Seq^{\wedge}\,A\,Q_A\,s_A \to P\,A\,Q_A\,s_A}
\end{array}
\end{equation}

To prove that this rule is sound we provide a witness
$\mathsf{dIndSeq}$ inhabiting the type in~\eqref{eq:ind-seq} as
follows:
\[\begin{array}{lll}
\mathsf{dIndSeq\;P\;cconst\;cpair\;A\;Q_A\;(const\,a)\;liftA}  &
    = &  \mathsf{cconst\;A\;Q_A\;a\;liftA}\\
\mathsf{dIndSeq\,P\,cconst\,cpair\,A\,Q_A\,(sPair\,B\,C\,e\,s_B\,s_C)\,
    (Q_B,Q_C, liftE, liftB, liftC)}  & = &
\mathsf{cpair\,A\,B\,C\,Q_A\,Q_B\,Q_C\,s_B\,s_C\,e\,liftE\,p_B\,p_C}
\end{array}\]
In the first clause ahove, $\mathsf{a : A}$, $\mathsf{Q_A : A \to
  Set}$, $\mathsf{liftA : Seq^{\wedge}\,A\,Q_A\,(const\,a) =
  Q_A\,a}$. In the second, $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C :
  C \to Set}$, $\mathsf{e : Equal\,A\,(B \times C)}$, $\mathsf{s_B :
  Seq\,B}$, $\mathsf{s_C : Seq\,C}$, $\mathsf{liftE :
  Equal^{\wedge}\,A\, (B \times C)\, Q_A\, (Q_B \times Q_C) \, e}$,
$\mathsf{liftB : Seq^{\wedge}\,B\,Q_B\,s_B}$, and $\mathsf{liftC :}$
$\mathsf{Seq^{\wedge}\,C\,Q_C\,s_C}$ --- which together ensure that
$\mathsf{(Q_B, Q_C, liftE, liftB, liftC) :
  Seq^{\wedge}\,A\,Q\,(sPair\,B\,C\,e\,s_B\,s_C)}$ --- and
$\mathsf{p_B} =\mathsf{dIndSeq\,P\,cconst\,cpair\,B\,Q_B\,s_B\,liftB :
  P\,B\,Q_B\,s_B}$ and $\mathsf{p_C}
=\mathsf{dIndSeq\,P\,cconst\,cpair\,C\,Q_C\,s_C\,liftC :
  P\,C\,Q_C\,s_C}$.

\subsection{(Deep) induction for $\mathsf{LTerm}$}\label{sec:ind-lam}

\begin{figure*}[t]

  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{LType^{\wedge}\,A\,Q_A\,(bool\,B\,e)} & = &\mathsf{\exists
  [Q_B]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(arr\, B\, C\, e\, T_B\, T_C)}
&=&\mathsf{\exists [Q_B] \,\exists [Q_c]\, Equal^{\wedge}\,A\,
  (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e \times
  \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(list\, B\, e\, T_B)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge}\,A\, (List\, B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times LType^{\wedge}\,B\,Q_B\,T_B}\\[1ex]
\mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A)} & = &
\mathsf{LType^{\wedge}\, A\, Q_A\, T_A}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} &
= & \mathsf{\exists [Q_B]\,\exists [Q_C]\, Equal^{\wedge} \, A\, (B \to
  C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e \times \,
  LType^{\wedge}\, B\, Q_B\, T_B \times \, LTerm^{\wedge}\, C\, Q_C\,
  t_C }\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)} & = &
\mathsf{\exists [Q_B]\, LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \,
  B\, A\, Q_B \, Q_A)\, t_{BA} \times LTerm^{\wedge}\, B\, Q_B\,
  t_B}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (list\, B\, e\, ts)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge} \, A\, (List\,B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times List^{\wedge}\, (LTerm\,B) \,
  (LTerm^{\wedge} \, B\, Q_B) \, ts}
\end{array}   \]}

\vspace*{-0.1in}

\caption{Predicate liftings for $\mathsf{LType}$ and
  $\mathsf{LTerm}$}\label{fig:liftings} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

To derive the deep induction rule for the GADT $\mathsf{LTerm}$ we use
its Henry Ford encoding from~\eqref{eq:eq_ltype}
and~\eqref{eq:eq_lterm}. We first define predicate lifting
$\mathsf{Arr^{\wedge} : \forall (A\, B : Set) \to (A \to Set) \to (B
  \to Set) \to (A \to B) \to Set}$ for arrow types, since arrow types
appear in $\mathsf{LType}$ and $\mathsf{LTerm}$. It is given by
$\mathsf{Arr^{\wedge}\, A\, B\, Q_A\, Q_B\, f = \forall (a : A) \to
  Q_A\,a \to Q_B\, (f\,a)}$. The predicate liftings
$\mathsf{LType^{\wedge} : \forall (A : Set) \to (A \to Set) \to
  LType\,A \to Set}$ for $\mathsf{LType}$ and $\mathsf{LTerm^{\wedge}
  :}$ $\mathsf{\forall (A : Set) \to (A \to Set) \to LTerm\,A \to
  Set}$ for $\mathsf{LTerm}$ are defined in Figure~\ref{fig:liftings}.
There, $\mathsf{s : String}$, $\mathsf{Q_A : A \to Set}$, $\mathsf{Q_B
  : B \to Set}$, $\mathsf{Q_C : C \to Set}$,
$\mathsf{K^{Bool}_{\top}}$ is the constantly {\color{red}
  $\mathsf{\top}$}-valued predicate on $\mathsf{Bool}$, $\mathsf{T_A :
  LType\, A}$, $\mathsf{T_B : LType \,B}$, $\mathsf{T_C : LType \,C}$,
$\mathsf{t_B : LTerm \, B}$, $\mathsf{t_C : LTerm \, C}$, and
$\mathsf{t_{BA} : LTerm \, (B \to A)}$.  Moreover, $\mathsf{e :
  Equal\,A\,Bool}$ in the first clause, $\mathsf{e : Equal\, A\, (B
  \to C)}$ in the second, $\mathsf{e : Equal\, A\, (List\, B)}$ in the
third, $\mathsf{e : Equal \, A \, (B \to C)}$ in the fifth, and
$\mathsf{e : Equal\, A\, (List \,B)}$, $\mathsf{ts : List\, (LTerm
  B)}$, and $\mathsf{List^\wedge}$ is the predicate lifting for lists
from~\eqref{eq:rose} in the seventh.

With these liftings in hand we can define the induction hypotheses
$\mathsf{dIndVar}$, $\mathsf{dIndAbs}$, $\mathsf{dIndApp}$, and
$\mathsf{dIndList}$ associated with $\mathsf{LTerms}$'s data
constructors. These are, respectively,
\[\begin{array}{l}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A  \to Set) \to}\\
  \quad\mathsf{\forall (A : Set) (Q_A : A \to Set) (s : String) (T_A :
  LType\, A) \to LType^{\wedge} \, A\, Q_A\, T_A \to P \, A\, Q_A\,
  (var \; s\, T_A)}\\[1ex]
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A \to Set)} \\ 
  \quad\mathsf{\to 
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C
  \to Set) (e : Equal\, A\, (B \to C)) (s : String) } \\ 
  \quad\mathsf{ \to (T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e } \\
  \quad\mathsf{
  \to LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)
  }\\[1ex]
  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set)} \\ 
  \quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B)} \\
  \quad \mathsf{
  \to P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }\\[1ex]
  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set)} \\ 
  \quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B))} \\ 
  \quad \mathsf{
    \to Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (list \,B \,e \, ts) }
\end{array}\]
The deep induction rule for $\mathsf{LTerm}$ is thus
\begin{equation}\label{eq:ind-lam}
\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
  \to Set) \to dIndVar\,P \to dIndAbs\,P \to dIndApp\,P \to
  dIndList\,P \to} \\ \quad \mathsf{\forall (A : Set)(Q_A : A \to
  Set)(t_A : LTerm\,A) \to LTerm^{\wedge}\,A\,Q_A\,t_A \to
  P\,A\,Q_A\,t_A}
\end{array}
\end{equation}

\begin{figure*}[t]

  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (var\;s\,T_A) \, liftA} & = & \mathsf{cvar \, A\, Q_A\, s\, T_A\,
  liftA}\\ 
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (abs \,B \,C \,e \,s \,T_B \, t_C) \, (Q_B , Q_C , liftE,
  lift_{T_B}, lift_{t_C})} & = & \mathsf{cabs\,A\,B\,C\, Q_A\,
  Q_B\, Q_C\, e\, s\, T_B\, t_C\, liftE\, lift_{T_B}\, p_C}\\
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (app \,B \,\,t_{BA} \, t_B)\, (Q_B , list_{t_{BA}}, list_{t_B})} &
= & \mathsf{capp\,A\,B\,Q_A\, Q_B\, t_{BA}\, t_B\, p_{BA} \, p_B}\\
  \mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clistc \, A\, Q_A\,
    (list \,B \,e \, ts) \, (Q_B , liftE', lift_{List})} & = & 
  = \mathsf{clistc \,A\,B\,Q_A\, Q_B\, e\, ts\, liftE'\, p_{List} }
\end{array}\]}

\vspace*{-0.1in}

\caption{$\mathsf{dIndLTerm}$}\label{fig:dindlterm} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

\noindent
To prove that this rule is sound we define a witness
$\mathsf{dIndLTerm}$ inhabiting the type in~\eqref{eq:ind-lam} as in
Figure~\ref{fig:dindlterm}. There, $\mathsf{s : String}$, $\mathsf{Q_A
  : A \to Set}$, $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C : C \to
  Set}$, $\mathsf{T_A : LType\,A}$, $\mathsf{T_B : LType\,B}$,
$\mathsf{t_B : LTerm\,B}$, $\mathsf{t_C : LTerm\,C}$, $\mathsf{t_{BA}
  : LTerm\,{\color{red}??}}$, $\mathsf{liftA : LTerm^{\wedge}\, A\,
  Q_A\, (var\;s\,T_A) = LType^{\wedge}\,A\,Q_A\,T_A}$, $\mathsf{liftE
  : Equal^{\wedge} \, A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e}$, $\mathsf{lift_{T_B}: LType^{\wedge} \, B\, Q_B\,
  T_B}$,~$\mathsf{lift_{t_C}: LTerm^{\wedge} \, C\, Q_C\,
  T_C}$,~$\mathsf{lift_{t_{BA}}: LTerm^{\wedge} \, (B \to A)\,
  (Arr^{\wedge} \, B\, A\, Q_B \, Q_A)\,
  t_{BA}}$,~$\mathsf{lift_{t_B}: LTerm^{\wedge} \, B\, Q_B\,
  t_B}$,\\ $\mathsf{liftE' : Equal^{\wedge}\, A\, (List\,B)\, Q_A\,
  (List^{\wedge}\, B\, Q_B)\, e}$, and $\mathsf{lift_{List}:
  List^{\wedge} \, (LTerm\, B) \, (LTerm^{\wedge}\, B\, Q_B) \, ts}$.
Moreover,
\[\begin{array}{lll}
\mathsf{p_C} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  C\, Q_C\, t_C\, lift_{t_C} : P \, C\, Q_C \, t_C }\\
\mathsf{p_B} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  B\, Q_B\, t_B\, lift_{t_B} : P \, B\, Q_B \, t_B }\\
\mathsf{p_{BA}} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp
  \,clist\, (B \to A)\,(Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \,
  t_{BA}\, lift_{t_{BA}} : P \, (B \to A)\, (Arr^{\wedge} \, B\, A\,
  Q_B \, Q_A) \, t_{BA}}\\ 
\mathsf{p_{List}} & = &\mathsf{liftListMap \, (LTerm\, B) \,
  (LTerm^{\wedge} \, B \, Q_B)\, (P\,B\,Q_B)\, p_{ts} \, ts\,
  lift_{List} : List^{\wedge}\, (LTerm\,B) \, (P\,B\,Q_B) \, ts}\\
\mathsf{p_{ts}} & = & \mathsf{dIndLTerm\, P\, cvar\, cabs\, capp\,
  clist\, B\, Q_B : PredMap\,(LTerm\,B) \,(LTerm^{\wedge}\, B\, Q_B)
  \, (P\,B\,Q_B)}
\end{array}\]
where, in the final clause, $\mathsf{PredMap : \forall\, (A : Set) \to
  (A \to Set) \to (A \to Set) \to Set }$ is the type constructor
producing the type of morphisms between predicates defined by
$\mathsf{PredMap \,A\, Q\,Q'\, = \forall\, (a : A) \to Q\,a \to
  Q'\,a}$ and $\mathsf{liftListMap : \forall\, (A : Set) \to (Q \, Q'
  : A \to Set) \to PredMap\,A\,Q\,Q' \to PredMap\,(List\,A)
  \,(List^{\wedge}\, A\, Q)\, (List^{\wedge}\, A\, Q')}$, which takes
a morphism $\mathsf{f}$ of predicates and produces a morphism of
lifted predicates, is defined by $\mathsf{liftListMap\, A\, Q\, Q'\,
  m\, nil\, tt = tt}$ (since $\mathsf{x : List^{\wedge}\, A\, Q\, nil
  = \top}$ must necessarily be $\mathsf{tt}$), and by
$\mathsf{liftListMap\, A\, Q\, Q'\, m\, (cons\, a\, l')\, (y, x') =
  (m\,a\,y, liftListMap\, A\, Q\, Q'\, m\, l'\, x')}$ (since $\mathsf{x
  : List^{\wedge}\, A\, Q\, (cons\, a\, l') = \top}$ must be of the
form $\mathsf{x = (y, x')}$ where $\mathsf{y : Q\,a}$ and $\mathsf{x'
  : List^{\wedge}\, A\, Q\, l'}$). {\color{red} Double-check!!}

\section{The general framework}\label{sec:framework}

{\color{red} HERE!!!!}

%Since, as seen in Section~\ref{sec:GADT-nested},
%GADTs featuring nesting do not admit an induction rule,
%we restrict our scope to GADTs not featuring nesting.
Finally, we generalize the approach taken in the previous examples and
provide a general framework to derive deep induction rules for
arbitrary GADTs.  For that, we need to give a grammar for the types we
will be considering.  A generic GADT
\begin{equation*}
\begin{array}{l}
\mathsf{data\ G : Set^\alpha \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;c_i\, :\, F_i\,G\,\ol{B} \to G (\ol{K_i\,\ol{B}})}
\end{array}
\end{equation*}
is defined by a finite number of constructors $\mathsf{c_i}$.
In the definition above, $\mathsf{F_i}$ is a type constructor with signature $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K_i}$ is a type constructor with signature $\mathsf{Set^{\beta} \to Set}$
(i.e. a type constructor of arity $\mathsf{\beta}$).
The overline notation denotes a finite list:
%$\mathsf{\ol{a}}$ is a list of types of length $\mathsf{\alpha}$,
%so that it can be applied to the type constructor $\mathsf{G}$ of arity $\mathsf{\alpha}$.
%and the vertical bars notation denotes the length of a list:
%$\mathsf{G}$ depends on a finite list of types $\mathsf{\ol{a}}$ of length $\mathsf{\alpha}$,
each of the $\mathsf{\alpha}$-many $\mathsf{K_i}$ is a type constructor of arity $\mathsf{\beta}$
so that it can be applied to the list of types $\mathsf{\ol{B}}$ of length $\mathsf{\beta}$.
%depends on a finite list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$,
Moreover, notice that the arity of $\mathsf{G}$ matches the number of type constructors $\mathsf{K_i}$.
We allow each $\mathsf{F_i}$ to be inductively built in the following ways (and with the following restrictions):
\begin{itemize}
\item $\mathsf{F_i = F_i' \times F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same signature as $\mathsf{F_i}$ and are built recursively from the same induction rules.
\item $\mathsf{F_i = F_i' + F_i''}$ where $\mathsf{F_i'}$ and $\mathsf{F_i''}$ have the same signature as $\mathsf{F_i}$ and are built recursively from the same induction rules.
\item $\mathsf{F_i = F_i' \to F_i''}$
where $\mathsf{F_i'}$ does not contain the recursive variable,
i.e., $\mathsf{F_i' : Set^\beta \to Set}$ is a type constructor of arity $\mathsf{\beta}$,
and $\mathsf{F_i''}$ has the same signature as $\mathsf{F_i}$ and is built recursively from the same induction rules.
\item $\mathsf{F_i\,G\,\ol{B} = G(\ol{F_j\,\ol{B}})}$
where none of the $\mathsf{\alpha}$-many $\mathsf{F_j}$ contains the recursive variable,
i.e., $\mathsf{F_j : Set^\beta \to Set}$
is a type constructor of arity $\mathsf{\beta}$
for each $\mathsf{j = 1, \dots, \alpha}$.
Such restriction is necessary to prevent nesting,
as that would break the induction rule as discussed in Section~\ref{sec:GADT-nested}.
\item $\mathsf{F_i\,G\,\ol{B} = H\,\ol{B}}$
where $\mathsf{H}$ is a type constructor of arity $\mathsf{\beta}$ not containing the recursive variable,
i.e., $\mathsf{H : Set^\beta \to Set}$.
Notice that this covers the case in which $\mathsf{F_i}$ is a closed type,
so, in particular, the unit and empty types, $\mathsf{1}$ and $\mathsf{0}$,
and the case in which $\mathsf{F_i}$ is a variable.
%The only requirement on $\mathsf{H}$ is that it needs to have a predicate-lifting function
%(and any type made of constant types, sums, products, arrow types and other GADTs satisfies this condition).
\item $\mathsf{F_i\,G\,\ol{B} = H (\ol{F_k\,G\,\ol{B}})}$
where $\mathsf{H}$ is a $\mathsf{\gamma}$-ary type constructor not containing the recursive variable,
i.e., $\mathsf{H : Set^\gamma \to Set}$,
and $\mathsf{F_k}$ has the same signature as $\mathsf{F_i}$  and is built recursively from the same induction rules, for every $\mathsf{k = 1 \dots \gamma}$.
Moreover, we require that $\mathsf{H}$ is not a GADT itself (but we allow it to be an ADT or even a nested type).
This way we know that $\mathsf{H}$ admits functorial semantics~\cite{jp20}, and thus there is a map function for $\mathsf{H^{\wedge}}$,
\[
\mathsf{HLMap : \forall (\ol{C : Set}) (\ol{Q_C\;Q_C' : C \to Set})
	\to \ol{PredMap\,C\,Q_C\,Q_C'}
	\to PredMap\,(H\,\ol{C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C'})}
\]
%where $\mathsf{PredMap : \forall (C : Set) \to (C \to Set) \to (C \to Set) \to Set}$
%is defined as
%\[
%\mathsf{PredMap\,C\,Q_c\,Q_c' = \forall (x : c) \to Q_c\,x \to Q_c'\,x}
%\]
%and represents the type of morphisms between predicates.
A concrete way to define $\mathsf{HLMap}$ is to proceed by induction on the structure of the type $\mathsf{H}$,
and give an inductive definition when $\mathsf{H}$ is an ADT or a nested type.
Such details are not essential to the present discussion, and thus we omit them.
\end{itemize}
We can summarize the above inductive definition with the following grammar
(but beware that the above restrictions and requirements still apply):
\[
\mathsf{F_i\,G\,\ol{B} :=
F_i'\,G\,\ol{B} \times F_i''\,G\,\ol{B}
\ \vert\ F_i'\,G\,\ol{B} + F_i''\,G\,\ol{B}
\ \vert\ F_i'\,\ol{B} \to F_i''\,G\,\ol{B}
\ \vert\ G(\ol{F_j\,\ol{B}})
\ \vert\ H\,\ol{B}
\ \vert\ H (\ol{F_k\,G\,\ol{B}})
}
\]
A further requirement that applies to all of the types appearing above,
including the types $\mathsf{K_i}$,
is that every type needs to have a predicate-lifting function.
This is not an overly restrictive condition, though:
all types made by sums, products, arrow types and type application do, and so do GADTs as defined above.
A concrete way to define the predicate-lifting function for a type is to proceed by induction on the structure of the type,
and we have seen in the previous sections examples of how to do so for products and type application.
We do not give here the general definition of lifting,
as that would require to first present a full type calculus,
and that is beyond the scope of the paper.

%Before proceeding with the derivation of the induction rule for a generic GADT,
%we need to define the type
%\[
%\mathsf{PredMap\,a\,Q_a\,Q_a' = \forall (x : a) \to Q_a\,x \to Q_a'\,x}
%\]
%where $\mathsf{Q_a}$ and  $\mathsf{Q_a'}$ are predicates on  $\mathsf{a : Set}$,
%which represents the type of morphisms between predicates $\mathsf{Q_a}$ and  $\mathsf{Q_a'}$.
%Moreover, if $\mathsf{H}$ is a type constructor having an associated predicate-lifting function $\mathsf{H^{\wedge}}$,
%we define a map function $\mathsf{HLMap}$ for the lifting $\mathsf{H^{\wedge}}$ as
%\[
%\mathsf{HLMap : \forall (a : Set) (Q_a\;Q_a' : a \to Set) \to PredMap\,a\,Q_a\,Q_a' \to PredMap\,(G\,a)\,(G^{\wedge}\,a\,Q_a)\,(G^{\wedge}\,a\,Q_a')}
%\]

Consider a generic GADT as defined above,
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\ G : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;c\, :\, F\,G\,\ol{B} \to G (K\,\ol{B})}
\end{array}
\end{equation}
which, for ease of notation,
we assume to be a unary type constructor (i.e. it depends on a single type parameter $\mathsf{A}$)
and to have only one constructor $\mathsf{c}$.
Extending the argument to GADTs of arbitrary arity and with multiple constructors presents no difficulty
other than heavier notation.
In the definition above, $\mathsf{F}$ has signature $\mathsf{(Set \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K}$ has signature $\mathsf{Set^{\beta} \to Set}$.
The constructor $\mathsf{c}$ can be rewritten using the $\mathsf{Equal}$ type as
\[
\mathsf{c : \forall (\ol{B : Set}) \to Equal\,A\,(K\,\ol{B}) \to F\,G\,\ol{B} \to G\,A}
\]
which is the form we shall use from now on.

In order to state the induction rule for $\mathsf{G}$,
we first need to define $\mathsf{G}$'s associated predicate-lifting function 
\[
\mathsf{G^{\wedge} : \forall (A : Set) \to (A \to Set) \to G\,A \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,A\,Q_A\,(c\,\ol{B}\,e\,x)
= \sum [\ol{Q_B : B \to Set}]\,
Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
\times F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}
\]
where $\mathsf{Q_A : A \to Set}$
and $\mathsf{c\,\ol{B}\,e\,x : G\,A}$,
i.e., $\mathsf{e : Equal\,A\,(K\,\ol{B})}$ and $\mathsf{x : F\,G\,\ol{B}}$.
As already mentioned before, we also assume to have liftings for $\mathsf{F}$,
\[
\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{B : Set})
\to (\forall (A : Set) \to (A \to Set) \to G\,A \to Set)
\to (\ol{B \to Set})
\to F\,G\,\ol{B} \to Set}
\]
and for  $\mathsf{K}$,
\[
\mathsf{K^{\wedge} : \forall (\ol{B : Set}) \to (\ol{B \to Set}) \to K\,\ol{B} \to Set}
\]
Finally, associate the function
\[
\begin{array}{l}
\mathsf{dIndC = \lambda (P : \forall (A : Set) \to (A \to Set) \to G\,A \to Set)} \\
\quad \mathsf{\to \forall (A : Set) (\ol{B : Set}) (Q_A : A \to Set) (\ol{Q_B : B \to Set}) (e : Equal\,A\,(K\,\ol{B})) (x : F\,G\,\ol{B})} \\
\quad \mathsf{\to Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
	\to F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x
	\to P\,A\,Q_A\,(c\,\ol{B}\,e\,x)}
\end{array}
\]
with the constructor $\mathsf{c}$.

%We can now define the induction rule for $\mathsf{G}$
The induction rule for $\mathsf{G}$ is
\[
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to G\,A \to Set) \to dIndC\,P
%\mathsf{\to \Big(
%\forall (\ol{a : Set}) (\ol{b : Set}) (\ol{Q_a : a \to Set}) (\ol{Q_b : b \to Set}) (\ol{e_a : Equal\,a\,(K\,\ol{b})}) (x : F\,G\,\ol{b})} \\
%\mathsf{\to \ol{Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e_a}
%	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
%	\to P\,\ol{a}\,\ol{Q_a}\,(C\,\ol{b}\,\ol{e_a}\,x)
%\Big)} \\
\to \forall (A : Set)(Q_A : A \to Set)(y : G\,A) \to G^{\wedge}\,A\,Q_A\,y \to P\,A\,Q_A\,y}
\]
As we already did in the previous examples,
we validate the induction rule by providing a term $\mathsf{dIndG}$ for the type above.
Define
\[
\mathsf{dIndG\,P\,cc\,A\,Q_A\,(c\,\ol{B}\,e\,x)\,(\ol{Q_B}, liftE, liftF)
= cc\,A\,\ol{B}\,Q_A\,\ol{Q_B}\,e\,x\,liftE\,(p\,x\,liftF)}
\]
where $\mathsf{cc : dIndC\,P}$ and
\begin{itemize}
\item $\mathsf{c\,\ol{B}\,e\,x : G\,A}$,
i.e., $\mathsf{e : Equal\,A\,(K\,\ol{B})}$,
and $\mathsf{x : F\,G\,\ol{B}}$;
\item $\mathsf{(\ol{Q_B}, liftE, liftF) : G^{\wedge}\,A\,Q_A (c\,\ol{B}\,e\,x)}$,
i.e.,
$\mathsf{Q_B : B \to Set}$ for each $\mathsf{B}$,
$\mathsf{liftE : Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e}$,
and $\mathsf{liftF : F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}$.
\end{itemize}
Finally, the morphism of predicates
\[
\mathsf{p : PredMap\,(F\,G\,\ol{B})\,(F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}) (F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}
\]
is defined by structural induction on $\mathsf{F}$ as follows:
\begin{itemize}
\item Case $\mathsf{F = F_1 \times F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same signature as $\mathsf{F}$.
We have that
\[
\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}
= Pair^{\wedge}(F_1\,G\,\ol{B})(F_2\,G\,\ol{B})(F_1^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}) (F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}
\]
By inductive hypothesis, there exist morphisms of predicates
\begin{gather*}
\mathsf{p_1 : PredMap\,(F_1\,G\,\ol{B})\,((F_1)^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}) ((F_1)^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})} \\
\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{B})\,((F_2)^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}) ((F_2)^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}
\end{gather*}
Thus, we define $\mathsf{p (x_1, x_2) (liftF_1, liftF_2) = (p_1\,x_1\,liftF_1, p_2\,x_2\,liftF_2)}$ 
for 
$\mathsf{x_1 : F_1\,G\,\ol{B}}$,
$\mathsf{liftF_1 : F_1^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x_1}$,
$\mathsf{x_2 : F_2\,G\,\ol{B}}$
and $\mathsf{liftF_2 : F_2^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x_2}$.
%%%
\item Case $\mathsf{F = F_1 + F_2}$ where $\mathsf{F_1}$ and $\mathsf{F_2}$ have the same signature as $\mathsf{F}$.
Analogous to case $\mathsf{F = F_1 \times F_2}$.
%%%
\item Case $\mathsf{F = F_1 \to F_2}$
where $\mathsf{F_1}$ does not contain the recursive variable,
i.e., $\mathsf{F_1 : Set^\beta \to Set}$,
and $\mathsf{F_2}$ has the same signature as $\mathsf{F}$.
We have that
\[
\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x = \forall (z : F_1\,\ol{B}) \to F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z \to F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,(x\,z)}
\]
where $\mathsf{x : F\,G\,\ol{B} = F_1\,\ol{B} \to F_2\,G\,\ol{B}}$.
By inductive hypothesis, there exist a morphism of predicates
\[
\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{B}) (F_2^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}) (F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}
\]
Thus, we define $\mathsf{p\,x\,liftF : F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x}$
for $\mathsf{liftF : F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x}$
as $\mathsf{p\,x\,liftF\,z\,liftF_1 = p_2 (x\,z) (liftF\,z\,liftF_1)}$
for $\mathsf{z : F_1\,\ol{B}}$
and $\mathsf{liftF_1 : F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z}$.
Notice that $\mathsf{F_1}$ not containing the recursive variable
is a necessary restriction, as the proof relies on
$\mathsf{F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x}$
and $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x}$
having the same domain $\mathsf{F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z}$.
%%%
\item Case $\mathsf{F\,G\,\ol{B} = G(F'\,\ol{B})}$
where $\mathsf{F'}$ does not contain the recursive variable,
i.e., $\mathsf{F' : Set^\beta \to Set}$.
Thus, $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} = P (F'\,\ol{B}) (F'^{\wedge}\,\ol{B}\,\ol{Q_B})}$.
So, $\mathsf{p}$ is defined as
\[
\mathsf{p = dIndG\,P\,cc\,(F'\,\ol{B})\,(F'^{\wedge}\,\ol{B}\,\ol{Q_B})}
\]
%for $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,\ol{Q_b}\,x}$.
%Note that the fact that the type constructor $\mathsf{F'}$ does not contain the recursive variable
%means that $\mathsf{G}$ does not feature nesting.
%As explained and exemplified in Section~\ref{sec:GADT-nested},
%that assumption is required to make induction work.
%%%
\item Case $\mathsf{F\,G\,\ol{B} = H\,\ol{B}}$
where $\mathsf{H}$ is a $\mathsf{\beta}$-ary type constructor not containing the recursive variable,
i.e., $\mathsf{H : Set^\beta \to Set}$.
In such case, $\mathsf{p : PredMap (H\,\ol{B}) (H^{\wedge}\,\ol{B}\,\ol{Q_B}) (H^{\wedge}\,\ol{B}\,\ol{Q_B})}$
is just the identity morphism of predicates.
%%%
\item Case $\mathsf{F\,G\,\ol{B} = H (\ol{F_k\,G\,\ol{B}})}$
where $\mathsf{H}$ is a $\mathsf{\gamma}$-ary type constructor not containing the recursive variable,
i.e., $\mathsf{H : Set^\gamma \to Set}$,
and $\mathsf{F_k}$ has the same signature as $\mathsf{F}$, for every $\mathsf{k = 1 \dots \gamma}$.
Moreover, we assume that $\mathsf{H}$ has an associated predicate-lifting function,
\[
\mathsf{H^{\wedge} : \forall (\ol{C : Set}) \to (\ol{C \to Set}) \to H\,\ol{C} \to Set}
\]
and that this predicate-lifting function has a map function $\mathsf{HLMap}$ of type
\[
\mathsf{\forall (\ol{C : Set}) (\ol{Q_C\;Q_C' : C \to Set}) \to PredMap\,\ol{C}\,\ol{Q_C}\,\ol{Q_C'}
\to PredMap\,(H\,\ol{C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C'})}
\]
That means that $\mathsf{H}$ cannot be a GADT,
as GADTs have no functorial semantics~\cite{jgj21}.
and incur in the issue exposed in Section~\ref{sec:GADT-nested},
but it can be an ADT or even a nested type as those types have functorial semantics~\cite{jp19,jgj21}.
Thus,
\[
\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} = H^{\wedge} (\ol{F_k\,G\,\ol{B}}) (\ol{F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}})}
\]
By induction hypothesis, there is a morphism of predicates
\[
\mathsf{p_k : PredMap\,(F_k\,G\,\ol{B}) (F_k^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}) (F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}
\]
for every $\mathsf{k = 1 \dots \gamma}$.
So, $\mathsf{p}$ is defined as
\[
\mathsf{p
= HLMap\,(\ol{F_k\,G\,\ol{B}}) (\ol{F_k^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}}) (\ol{F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}})\,\ol{p_k}}
\]
\end{itemize}
%Notice that, by proceeding by structural induction as above,
%we have implicitly defined a grammar for GADTs admitting an induction rule.
%In particular, we allow $\mathsf{F}$ to feature sums, products, arrow types and (limited) type application.

\section{{\color{red}Induction for GADTs with nesting}}\label{sec:GADT-nested}

%%% Change title: something about the inherent limitations of the method
%%% we might want to extend previous result
%%% but we cannot so we already have the best possible result

In the previous sections,
we derive induction rules for examples of GADTs that do not feature nesting,
in the sense that their constructors contain no nested calls of the recursive variable,
as truly nested types (such as $\mathsf{Bush}$, Equation~\ref{eq:bush}) do.
Since both nested types
and GADTs without nesting admit induction rules,
as seen in the previous sections,
it is just natural to expect that GADTs with nesting would as well.
Surprisingly, that is not the case:
indeed, the induction rule generally relies on (unary) parametricity of the semantic interpretation,
and in the case of nested types it also relies on functorial semantics~\cite{jp20},
but GADTs cannot admit both functorial and parametric semantics at the same time~\cite{jgj21}.
In this section we show how induction for GADTs featuring nesting goes wrong
by analyzing the following concrete example of such a type.
\begin{equation}\label{gadt-nested}
\begin{array}{l}
\mathsf{data\, G\,(a : Set) : Set\,where}\\
\mathsf{\;\;\;\;\;\;C :\, G(G\,a) \to G(a \times a)}
\end{array}
\end{equation}
The constructor $\mathsf{C}$ can be rewritten as
\[
\mathsf{C : \exists (b : Set) \to Equal\,a\,(b \times b) \to G(G\,b) \to G\,a}
\]
which is the form we shall use from now on.
The predicate-lifting function of $\mathsf{G}$,
\[
\mathsf{G^{\wedge} : \forall (a : Set) \to (a \to Set) \to G\,\ol{a} \to Set}
\]
is defined as
\[
\mathsf{G^{\wedge}\,a\,Q_a\,(C\,b\,e\,x)
= \exists (Q_b : b \to Set)
\to Equal^{\wedge}\,a\,(b \times b)\,Q_a\,(Pair^{\wedge}\,b\,b\,Q_b\,Q_b)\,e
\times G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b)\,x}
\]
where $\mathsf{Q_a : a \to Set}$, $\mathsf{e : Equal\,a\,(b \times b)}$ and $\mathsf{x : G(G\,b)}$.
Finally, let $\mathsf{CC}$ be the function
\begin{multline*}
\mathsf{\lambda (P : \forall (a : Set) \to (a \to Set) \to G\,a \to Set)} \\
\mathsf{\to \forall (a\;b : Set) (Q_a : a \to Set) (Q_b : b \to Set) (e : Equal\,a\,(b \times b)) (x : G\,(G\,b))} \\
\mathsf{\to Equal^{\wedge}\,a\,(b \times b)\,Q_a\,(Pair^{\wedge}\,b\,b\,Q_b\,Q_b)\,e
	\to P\,(G\,b)\,(P\,b\,Q_b)\,x
	\to P\,a\,Q_a\,(C\,b\,e\,x)}
\end{multline*}
associated with the $\mathsf{C}$ constructor.

The induction rule for $\mathsf{G}$ is
\begin{multline*}
\mathsf{\forall (P : \forall (a : Set) \to (a \to Set) \to G\,a \to Set) \to CC\,P} \\
\mathsf{\to \forall (a : Set)(Q_a : a \to Set)(y : G\,a) \to G^{\wedge}\,a\,Q_a\,y \to P\,a\,Q_a\,y}
\end{multline*}
Consistently with the previous examples,
to validate the induction rule
we try to define a term of the above type, $\mathsf{DIG}$, as
\[
\mathsf{DIG\,P\,cc\,a\,Q_a\,(C\,b\,e\,x)\,(Q_b, L_E, L_G)
= cc\,a\,b\,Q_a\,Q_b\,e\,x\,L_E\,p}
\]
where $\mathsf{cc : CC\,P}$ and
\begin{itemize}
\item $\mathsf{C\,b\,e\,x : G\,a}$,
i.e., $\mathsf{e : Equal\,a\,(b \times b)}$ and $\mathsf{x : G(G\,b)}$;
\item $\mathsf{(Q_b, L_E, L_G) : G^{\wedge}\,a\,Q_a\,(C\,b\,e\,x)}$,
i.e., $\mathsf{Q_b : b \to Set}$, $\mathsf{L_E : Equal^{\wedge}\,a\,(b \times b)\,Q_a\,(Pair^{\wedge}\,b\,b\,Q_b\,Q_b)\,e}$,
and $\mathsf{L_G : G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b)\,x}$.
\end{itemize}
We still need to define $\mathsf{p : P\,(G\,b)\,(P\,b\,Q_b)\,x}$.
We do so by using the induction rule and letting
\[
\mathsf{p = DIG\,P\,cc\,(G\,b)\,(P\,b\,Q_b)\,x\,q}
\]
where we still need to provide $\mathsf{q : G^{\wedge}\,(G\,b)\,(P\,b\,Q_b)\,x}$.
If we had the map function of $\mathsf{G^{\wedge}}$,
\[
\mathsf{GLMap : \forall (a : Set) (Q_a\;Q_a' : a \to Set) \to PredMap\,a\,Q_a\,Q_a' \to PredMap\,(G\,a)\,(G^{\wedge}\,a\,Q_a)\,(G^{\wedge}\,a\,Q_a')}
\]
then we would be able to define
\[
\mathsf{q = GLMap\,(G\,b)\,(G^{\wedge}\,b\,Q_b)\,(P\,b\,Q_b)\,(DIG\,P\,cc\,b\,Q_b)\,x\,L_G}
\]
Unfortunately, we cannot define such a $\mathsf{GLMap}$.
Indeed, its definition would have to be
\[
\mathsf{GLMap\,a\,Q_a\,Q_a'\,M\,(C\,b\,e\,x)\,(Q_b, L_E, L_G) = (Q_b', L_E', L_G')}
\]
where $\mathsf{Q_a : a \to Set}$, $\mathsf{Q_a' : a \to Set}$,
$\mathsf{M : PredMap\,a\,Q_a\,Q_a'}$,
$\mathsf{C\,b\,e\,x : G\,a}$, i.e.,
\begin{itemize}
\item $\mathsf{e : Equal\,a\,(b \times b)}$;
\item $\mathsf{x : G\,(G\,b)}$;
\end{itemize}
$\mathsf{(Q_b, L_E, L_G) :G^{\wedge}\,a\,Q_a\,(C\,b\,e\,x)}$, i.e.,
\begin{itemize}
\item $\mathsf{Q_b : b \to Set}$;
\item $\mathsf{L_E : Equal^{\wedge}\,a\,(b \times b)\,Q_a\,(Pair^{\wedge}\,b\,b\,Q_b\,Q_b)\,e}$;
\item $\mathsf{L_G : G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b)\,x}$;
\end{itemize}
and $\mathsf{(Q_b', L_E', L_G') : G^{\wedge}\,a\,Q_a'\,(C\,b\,e\,x)}$, i.e.,
\begin{itemize}
\item $\mathsf{Q_b' : b \to Set}$;
\item $\mathsf{L_E' : Equal^{\wedge}\,a\,(b \times b)\,Q_a'\,(Pair^{\wedge}\,b\,b\,Q'_b\,Q'_b)\,e}$;
\item $\mathsf{L_G' : G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b')\,x}$;
\end{itemize}
%\begin{gather*}
%\mathsf{Q_a\;Q_a' : a \to Set} \\
%\mathsf{Q_b\;Q_b' : b \to Set} \\
%%\mathsf{M : PredMap\,b\,Q_a\,Q_a'} \\
%\mathsf{L_E : Equal^{\wedge}\,a\,(b \times b)\,Q_a\,(Pair^{\wedge}\,b\,b\,Q_b\,Q_b)\,e} \\
%\mathsf{L_G\ : G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b)\,x} \\
%\mathsf{L_E' : Equal^{\wedge}\,a\,(b \times b)\,Q_a'\,(Pair^{\wedge}\,b\,b\,Q'_b\,Q'_b)\,e} \\
%\mathsf{L_G' : G^{\wedge}\,(G\,b)\,(G^{\wedge}\,b\,Q_b')\,x}
%\end{gather*}
In other words, we have a proof $\mathsf{L_E}$ of the (extensional) equality
of the predicates $\mathsf{Q_a}$ and $\mathsf{Pair^{\wedge}\,b\,b\,Q_b\,Q_b}$
and a morphism of predicates $\mathsf{M}$ from $\mathsf{Q_a}$ to $\mathsf{Q_a'}$,
and we need to use those to deduce a proof of the (extensional) equality
of the predicates $\mathsf{Q_a'}$ and $\mathsf{Pair^{\wedge}\,b\,b\,Q_b'\,Q_b'}$,
for some  for some predicate $\mathsf{Q_b'}$ on $\mathsf{b}$.
But that is not generally possible:
the facts that $\mathsf{Q_a}$ is equal to $\mathsf{Pair^{\wedge}\,b\,b\,Q_b\,Q_b}$
and that there is a morphism of predicates $\mathsf{M}$ from $\mathsf{Q_a}$ to $\mathsf{Q_a'}$
do not guarantee that $\mathsf{Q_a'}$ is equal to $\mathsf{Pair^{\wedge}\,bb\,Q_b'\,Q_b'}$ for some $\mathsf{Q_b'}$.

At a deeper level, the fundamental issue is that
the $\mathsf{Equal}$ type does not have functorial semantics,
so that having morphisms $\mathsf{A \to A'}$ and $\mathsf{B \to B'}$
and a proof that $\mathsf{A}$ is equal to $\mathsf{A'}$
does not provide a proof that $\mathsf{B}$ is equal to $\mathsf{B'}$.
This is because GADTs can either have a syntax-only semantics or a functorial-completion semantics.
Since we are interested in induction rules, we considered the syntax-only semantics,
which is parametric but not functorial.
Had we considered the functorial-completion semantics, which is functorial,
we would have forfeited parametricity instead.
In both cases, thus, we cannot derive an induction rule for GADTs featuring nesting.
Unlike nestes types, indeed, GADTs do not admit a semantic interpretation that is both parametric and functorial~\cite{jgj21}.

\section{Case Study: Well-typed lambda terms}\label{sec:app}

{\color{red} Can get rid of Maybe using non-empty lists and postulates}

In this section we use deep induction for the $\mathsf{LTerm}$ GADT to 
extract the type from a lambda term. 
We have a predicate 
\begin{align*}
  &\mathsf{getType : \forall \, (a : Set) \to (t : LTerm\,a) \to Set} \\
  &\mathsf{getType \,a \,t = Maybe \, (LType \, a)}
\end{align*}
that takes a lambda term and produces its type (using $\mathsf{Maybe}$ to represent
potential failure).
We want to show this predicate is satisfied for every element of $\mathsf{LTerm\,a}$.
Because of the $\mathsf{ListC}$ constructor, this cannot be achieved without 
deep induction. In particular, deep induction is required to apply the induction to the 
individual terms in a list of terms.

So, using deep induction, we want to prove: 
\[
  \mathsf{getTypeProof : \forall \, (a : Set) \to (t : LTerm\,a) \to getType \,a \,t} \\
\]
which we prove by 
\[
  \mathsf{getTypeProof \,a \,t = 
    DILTerm\, (\lambda \, b\, Q_b\, t\, \to getType \,b\, t) \,
    gtVar\,  gtABs\,  gtApp\,  gtListC\,  a\,\, K1\,\, t\, (LTerm^{\wedge}K1\, a\, t)
    }
\]
where $\mathsf{K1 : a \to Set}$ is the constantly true predicate:
\[
  \mathsf{K1 \, x = \top}
\]
and $\mathsf{LTerm^{\wedge}K1\, a\, t : LTerm^{\wedge}\, a\, K1\, t}$.
Notice that there is no space in $\mathsf{LTerm^{\wedge}K1}$, because
\[
  \mathsf{LTerm^{\wedge}K1 : \forall (a : Set) (t : LTerm A) \to LTerm^{\wedge}\, a\, K1\,t}
\]
is a function that we will define. In addition to defining $\mathsf{LTerm^{\wedge}K1}$,
we also have to give a proof for each constructor $\mathsf{Var, Abs, App, ListC}$:
\begin{multline*}
  \mathsf{gtVar : \forall (a : Set) (Q_a : a \to Set) (s : String) (T_a : LType\, a) 
      \to LType^{\wedge} \, a\, Q_a\, T_a\, \to Maybe\, (LType\, a)}
\end{multline*}

% abs 
\begin{multline*}
    \mathsf{gtAbs : \forall (a \, b\, c: Set) (Q_a : a \to Set) (Q_b : b \to Set) (Q_c : c \to Set) 
      (e : Equal\, a\, (b \to c)) (s : String)} \\ 
    %
    \mathsf{ (T_b : LType\, b) (t_c : LTerm \, c)
      \to Equal^{\wedge}\, a\, (b \to c)\, Q_a\, (Arr^{\wedge}\, b\, c\, Q_b\, Q_c)\, e
      \to LType^{\wedge} \, b\, Q_b\, T_b} \\ 
    %
    \mathsf{
      \to Maybe\, (LType\,c) \to Maybe\, (LType\, a)}
\end{multline*}

% app 
\begin{multline*}
    \mathsf{gtApp : \forall (a \, b : Set) (Q_a : a \to Set) (Q_b : b \to Set) 
      (t_{ba} : LTerm\, (b \to a)) (t_b : LTerm\, b)} \\ 
    \mathsf{
      \to Maybe\, (LType\, (b \to a))
      \to Maybe\, (LType\, b)
      \to Maybe\, (LType\, a)
    }
\end{multline*}

% list 
\begin{multline*}
    % abs
    \mathsf{gtListC : \forall (a \, b : Set) (Q_a : a \to Set) (Q_b : b \to Set) 
      (e : Equal\, a\, (List\, b)) (ts : List\, (LTerm\, b))} \\ 
    \mathsf{
      \to Equal^{\wedge}\, a\, (List\,b) \, Q_a\, (List^{\wedge}\, b\, Q_b)\, e
      \to List^{\wedge}\, (LTerm\, b)\, (getType\, b)\, ts
      \to Maybe\, (LType\, a)
    }
\end{multline*}

For variables we simply return the type $\mathsf{T_a}$, 
and the cases for abstraction and application are similar.
The interesting case is $\mathsf{gtListC}$, in which we have to use 
the results of ($\mathsf{List^{\wedge}\, (LTerm\, b)\, (getType\, b)\, ts}$)
in order to extract the type of one of the terms in the list. 
To define $\mathsf{gtListC}$ we pattern-match on the list of terms $\mathsf{ts}$.

If $\mathsf{ts}$ is the empty list (denoted by $\mathsf{[]}$), we cannot extract a type, 
so we return $\mathsf{nothing}$. \\
{\color{red} Maybe handle this case differently, by using non-empty lists, for example}
\[
  \mathsf{gtListC\, a\, b\, Q_a\, Q_b\, e\, [] \, L_e L_{ts} = nothing}
\]
If $\mathsf{ts}$ is a non-empty list, we pattern match on $\mathsf{L_{ts}}$ and use the result
to construct the type we need:
\begin{align*}
  &\mathsf{gtListC\, a\, b\, Q_a\, Q_b\, e\, (t :: ts) \, L_e (nothing , L_{ts}) = nothing} \\
  &\mathsf{gtListC\, a\, b\, Q_a\, Q_b\, e\, (t :: ts) \, L_e (just\, T_b , L_{ts}) = just \, (TList\, b\, e\, T_b)}
\end{align*}
where $\mathsf{e : Equal\, a\, (List\,b)}$ and $\mathsf{T_b : LType\, b}$.

\subsection{Defining $\mathsf{LTerm^{\wedge}K1}$}

{\color{red} Maybe this section can be deleted by assuming $\mathsf{LTerm^{\wedge}\, a\, K1\,t = K1}$.
Maybe we can say this derives from parametricity.
Currently we say that $\mathsf{K1 \times K1 = K1}$ based on the fact that products are a built-in type and so this seems to be obviously true.}

The last piece of infrastructure we need to define $\mathsf{getTypeProof}$ is a function 
\[
  \mathsf{LTerm^{\wedge}K1 : \forall (a : Set) \to (t : LTerm A) \to LTerm^{\wedge}\, a\, K1\,t}
\]
that provides a 
proof of $\mathsf{LTerm^{\wedge}\,a\,K1\,t}$ for any term $\mathsf{t : LTerm\, a}$. 
Because $\mathsf{LTerm^{\wedge}}$ is defined in terms of $\mathsf{LType^{\wedge}}$, $\mathsf{Arr^{\wedge}}$, 
and $\mathsf{List^{\wedge}}$, we will need analogous functions for these liftings as well. 
We only give the definition of $\mathsf{LTerm^{\wedge}K1}$, but the definitions for 
$\mathsf{LType^{\wedge}}$, $\mathsf{Arr^{\wedge}}$, and $\mathsf{List^{\wedge}}$ are analogous.

$\mathsf{LTerm^{\wedge}K1}$ is defined by pattern matching on the lambda term $\mathsf{t}$. 
For the $\mathsf{Var}$ case, let $\mathsf{t = (Var\, s\, T_a)}$ and define
\[
  \mathsf{LTerm^{\wedge}K1\,a\,(Var\,s\,T_a) = LType^{\wedge}K1\,a\,T_a}
\]
For the $\mathsf{Abs}$ case, let $\mathsf{t = (Abs \,b \,c \, e \,s \,T_b \, t_c)} $ and recall the 
definition of $\mathsf{LTerm^{\wedge}}$ for the $\mathsf{Abs}$ constructor, 
instantiating the predicate $\mathsf{Q_a}$ to $\mathsf{K1}$: 
% \,a\,Q_a\, (Abs \,b \,c \,e \,s \,T_b \,t_c)}$, :
\begin{multline*}
  \mathsf{LTerm^{\wedge}\,a\,K1\,
  (Abs \,b \,c \,e \,s \,T_b \,t_c)} \\ 
\mathsf{
  = \exists (Q_b : b \to Set) (Q_c : c \to Set) \to
      Equal^{\wedge} \, a\, (b \to c)\, K1\, (Arr^{\wedge} \, b\, c\, Q_b \, Q_c)\, e} \\ 
\mathsf{
      \times \, LType^{\wedge}\, b\, Q_b\, T_b
      \times \, LTerm^{\wedge}\, c\, Q_c\, t_c }
\end{multline*}
so to define the $\mathsf{Abs}$ case of $\mathsf{LTerm^{\wedge}K1}$, we need a proof of 
\[
  \mathsf{Equal^{\wedge} \, a\, (b \to c)\, K1\, (Arr^{\wedge} \, b\, c\, Q_b \, Q_c)\, e}
\]
i.e., that $\mathsf{K1}$ is (extensionally) equal to the lifting $\mathsf{(Arr^{\wedge} \, b\, c\, Q_b \, Q_c)}$
for some predicates $\mathsf{Q_b, Q_c}$. The only reasonable choice for $\mathsf{Q_b}$ and $\mathsf{Q_c}$ 
is to let both be $\mathsf{K1}$, which means we need a proof of:
\[
  \mathsf{Equal^{\wedge} \, a\, (b \to c)\, K1\, (Arr^{\wedge} \, b\, c\, K1 \, K1)\, e}
\]
Since we are working with proof-relevant predicates (i.e., functions into $\mathsf{Set}$ rather than 
functions into $\mathsf{Bool}$), 
the lifting $\mathsf{(Arr^{\wedge} \, b\, c\, K1 \, K1)}$ of $\mathsf{K1}$ to arrow types
is not identical to $\mathsf{K1}$ on arrow types, but the predicates are (extensionally) isomorphic. 
We discuss this issue in more detail at the end of the section. 
For now, we assume a proof 
\[
  \mathsf{Equal^{\wedge}ArrK1 : Equal^{\wedge} \, a\, (b \to c)\, K1\, (Arr^{\wedge} \, b\, c\, K1 \, K1)\, e}
\]
and define the $\mathsf{Abs}$ case of $\mathsf{LTerm^{\wedge}K1}$ as 
\[
  \mathsf{LTerm^{\wedge}K1\,a\, (Abs \,b \,c \, e \,s \,T_b \, t_c) = 
    (K1 , K1 , Equal^{\wedge}ArrK1 , LType^{\wedge}K1 \,b\,T_b\, , LTerm^{\wedge}K1\,c\,t_c)
  }
\]
For the $\mathsf{App}$ case, let $\mathsf{t = (App \,b \,\,t_{ba} \, t_b)}$ and just as we did for the $\mathsf{Abs}$ case, 
recall the definition of $\mathsf{LTerm^{\wedge}\,a\,(App \,b \,\,t_{ba} \, t_b)}$ with all of the predicates
instantiated with $\mathsf{K1}$: 
\[
\mathsf{
      LTerm^{\wedge}\, (b \to a)\, (Arr^{\wedge} \, b\, a\, K1 \, K1)\, t_{ba}
      \times LTerm^{\wedge}\, b\, K1\, t_b }
\]
The second component can be given using $\mathsf{LTerm^{\wedge}K1}$, and we can 
define the first component using a proof of
\[
\mathsf{LTerm^{\wedge}\, (b \to a)\, K1\,\, t_{ba}}
\]
and a map-like function 
\begin{multline*}
  \mathsf{LTerm^{\wedge}EqualMap : \forall\, \{a : Set\} \to (Q_a\,Q'_a : a \to Set) 
    \to (Equal^{\wedge}\,a\,a\,Q_a\,Q'_a\,rfl)} \\
    \mathsf{
    \to PredMap\, (LTerm^{\wedge}\,a\,Q_a)\,(LTerm^{\wedge}\,a\,Q'_a)}
\end{multline*}
that takes two (extensionally) equal predicates with the same carrier 
and produces a morphism of predicates between their liftings. The definition is 
straightforward enough, so we omit the details. 
But it is worth noting that while a true $\mathsf{HLMap}$ function, 
which takes a \emph{morphism} of predicates instead of a proof of equality, cannot be defined 
for GADTs in general, an analogue of $\mathsf{LTerm^{\wedge}EqualMap}$ should be definable 
for every GADT. These analogues of $\mathsf{LTerm^{\wedge}EqualMap}$ will be required to define 
$\mathsf{G^{\wedge}K1}$ whenever $\mathsf{G}$ has a constructor of the form
  ($\mathsf{c : G \, (F\, b) \to G\,(K\,b)}$).

Using $\mathsf{LTerm^{\wedge}EqualMap}$, we can define the $\mathsf{App}$ case of $\mathsf{LTerm^{\wedge}K1}$ as 
\[
  \mathsf{LTerm^{\wedge}K1\,a\, (App \,b \,\,t_{ba} \, t_b) = 
    (K1 , L_{Arr^{\wedge}\,K1} , LTerm^{\wedge}K1\,b\,t_b)}
\]
where $\mathsf{ L_{Arr^{\wedge}\,K1} : LTerm^{\wedge}\, (b \to a)\, (Arr^{\wedge} \, b\, a\, K1 \, K1)\, t_{ba}}$ is defined as
\[
  \mathsf{L_{Arr^{\wedge}\,K1} =  LTerm^{\wedge}EqualMap\,K1\,(Arr^{\wedge}\,b\,a\,K1\,K1)\,\, Equal^{\wedge}ArrK1\, \,\,t_{ba}\, \,L_{K1}}
\]
where $\mathsf{L_{K1} = LTerm^{\wedge}K1\, (b \to a)\, t_{ba} : LTerm^{\wedge}\, (b \to a) \, K1\, t_{ba}}$.

Finally, we define the $\mathsf{ListC}$ case for $\mathsf{LTerm^{\wedge}K1}$. 
Let $\mathsf{t = (ListC\, b\, e\, ts)}$ and recall the definition of $\mathsf{LTerm^{\wedge}\,a\,(ListC\, b\, e\, ts)}$ with
all of the predicates instantiated to $\mathsf{K1}$: 
\[
\mathsf{ Equal^{\wedge} \, a\, (List\,b)\, K1\, (List^{\wedge} \, b\, K1) \, e
      \times List^{\wedge}\, (LTerm\,b) \, (LTerm^{\wedge} \, b\, K1) \, ts }
\]
We can give the first component by assuming a proof \,
$\mathsf{Equal^{\wedge}ListK1 : Equal^{\wedge} \, a\, (List\,b)\, K1\, (List^{\wedge} \, b\, K1) \, e}$, 
but for the second component we again have multiple liftings nested together. In this case, we can 
get a proof of
\[
\mathsf{ List^{\wedge}\, (LTerm\,b) \, (LTerm^{\wedge} \, b\, K1) \, ts }
\]
using 
\[\mathsf{List^{\wedge}map : \forall\, (a : Set) \to (Q_a \, Q'_a : a \to Set) 
  \to PredMap\,Q_a\,Q'_a \to PredMap \,(List^{\wedge}\, a\, Q_a)\, (List^{\wedge}\, a\, Q'_a)}\]
to map a morphism of predicates 
\[
  \mathsf{PredMap \, (K1)\, (LTerm^{\wedge}\,b\,K1)}
\]
to a morphism of lifted predicates
\[
  \mathsf{PredMap \,(List^{\wedge}\,(LTerm\,b)\,K1) \, (List^{\wedge}\,(LTerm\,b)\, (LTerm^{\wedge}\,b\,K1))}
\]

We define the $\mathsf{ListC}$ case of $\mathsf{LTerm^{\wedge}K1}$ as 
\[
  \mathsf{LTerm^{\wedge}K1\,a\, (ListC\, b\, e\, ts) =
    (K1 , \, Equal^{\wedge}ListK1 , \,  L_{List^{\wedge}LTerm^{\wedge}K1} ) }
\]
where
 $\mathsf{L_{List^{\wedge}LTerm^{\wedge}K1} : List^{\wedge}\, (LTerm\,b) \, (LTerm^{\wedge} \, b\, K1) \, ts}$
\[
  \mathsf{ 
 L_{List^{\wedge}LTerm^{\wedge}K1} =
  List^{\wedge}map \, (LTerm\,b) \, K1 \, (LTerm^{\wedge}\, b\, K1) \, m_{K1} \, ts \, (List^{\wedge}K1\, (LTerm\, b)\, ts)
    }
\]
and 
 $\mathsf{m_{K1} : PredMap \, (K1)\, (LTerm^{\wedge}\,b\,K1)}$
\[
  \mathsf{m_{K1} \, t\, \ast\, = LTerm^{\wedge}K1\, b\, t}
\]
where $\mathsf{\ast}$ is the single element of $\mathsf{(K1\, t)}$. 
The use of $\mathsf{List^{\wedge}map}$ is required in the $\mathsf{ListC}$ case because 
$\mathsf{ListC}$ takes an argument of type $\mathsf{List\, (LTerm\, b)}$. 
The same technique can be used to define $\mathsf{G^{\wedge}K1}$ whenever $\mathsf{G}$ has 
a constructor of the form ($\mathsf{c : F \, (G\, a) \to G\, (K\,b)}$). 
We only allow constructors of this form when 
$\mathsf{F}$ is a nested type or ADT, so we are guaranteed to have a $\mathsf{F^{\wedge}map}$ function.

\subsection{Liftings of $\mathsf{K1}$}
To provide a proof of $\mathsf{G^{\wedge}\, a\, K1 \, t}$ for every term $\mathsf{t : G\, a}$, 
we need to know that the lifting of $\mathsf{K1}$ by a type $\mathsf{H}$ is extensionally equal to $\mathsf{K1}$ on $\mathsf{H}$. 
For example, we might need a proof that $\mathsf{Pair^{\wedge}\,a\,b\,K1\,K1}$ is equal to the predicate $\mathsf{K1}$ on pairs. 
Given a pair $\mathsf{(x , y) : a \times b}$, we have

\begin{align*}
  &\mathsf{Pair^{\wedge}\,a\,b\,K1\,K1 (x, y)}  \\
  = \, \, &\mathsf{K1 \, x \times K1\, y} \\ 
  = \, \, &\mathsf{\top \times \top}
\end{align*}
while 
\begin{align*}
 \mathsf{K1\, (x, y) = \top}
\end{align*}
and while these types are not equal they are clearly isomorphic. So for simplicity of presentation, we 
assume ($\mathsf{F^{\wedge}\,a\,K1}$) is equal to $\mathsf{K1}$ for every nested type and ADT $\mathsf{F}$.

%\section{Primitive representation for GADTs}

\section{Conclusion}\label{sec:conclusion}

\section{TODO}

\begin{itemize}
\item find correct entcsmacro file (current one is for 2018). Maybe ask Ana Sokolova (anas@cs.uni-salzburg.at).
\item reference (correctly) Haskell Symposium paper
\item reference inspiration for STLC GADT : https://www.seas.upenn.edu/~cis194/spring15/lectures/11-stlc.html
\item Data type vs. data structure
\item Weird to code in Agda if we're talking about induction rules for
  Coq? 
\item Agda style conventions
\item spacing in data type declarations
\item Mention Agda flags that need to be toggled to handle true nesting
\end{itemize}

\begin{thebibliography}{10}\label{bibliography}

\bibitem{atk12} Atkey, R.  {\em Relational parametricity for higher
  kinds}.  Computer Science Logic, pp.~46-61, 2012.

\bibitem{bfss90} Bainbridge, E. S., Freyd, P. Scedrov, A., and Scott,
  P. J. {\em Functorial polymorphism}. Theoretical Computer Science
  70(1), pp. 35-64, 1990.

\bibitem{bm98} Bird, R. and Meertens, L. {\em Nested
  datatypes}. Proceedings, Mathematics of Program Construction,
  pp. 52–67, 1998.

\bibitem{ch03} Cheney, J. and Hinze, R. {\em First-class phantom
 types}. CUCIS TR2003-1901, Cornell University, 2003.

\bibitem{ch88} Coquand, T. and Huet, G. {\em The calculus of
  constructions}. Information and Computation 76(2/3), 1988.

\bibitem{chl} Chlipala, A. {\em Library Inductive
  Types}. $\mathtt{http://adam.chlipala.net/cpdt/html/InductiveTypes.html}$

\bibitem{coq20} The Coq Development Team. {\em The Coq Proof
  Assistant}, version 8.11.0, January 2020.
  $\mathtt{https://doi.org/10.5281/zenodo.3744225}$

\bibitem{fs18} Fu, P. and Selinger, P.  Dependently typed folds for
  nested data types, 2018. $\mathtt{https://arxiv.org/abs/1806.05230}$

\bibitem{gjfor15} Ghani, N., Johann, P., Nordvall Forsberg, F.,
  Orsanigo, F., and Revell, T. {\em Bifibrational functorial semantics
    for parametric polymorphism}. Proceedings, Mathematical
  Foundations of Program Semantics, pp. 165-181, 2015.

\bibitem{hin03} Hinze, R. {\em Fun with phantom types}. Proceedings,
 The Fun of Programming, pp. 245–262, 2003.

\bibitem{web-page} Johann, P., Ghiorzi, E., and Jeffries,
  D. Accompanying Agda code for this paper.
    $\mathtt{https://cs.appstate.edu/~johannp/FoSSaCS21Code.html}$

\bibitem{jgj21f} Johann, P., Ghiorzi, E., and Jeffries, D. {\em
  Parametricity for primitive nested types}. Proceedings, Foundations
  of Software Science and Computation Structures, pp. 324-343, 2021.

\bibitem{jgj21} Johann, P., Ghiorzi, E., and Jeffries, D. {\em
  Parametricity in the presence of GADTs}. Submitted, 2021.

\bibitem{jp19} Johann, P. and Polonsky, A. {\em Higher-kinded data
  types: Syntax and semantics} Proceedings, Logic in Computer Science
  2019. {\color{red} PAGES?}

\bibitem{jp20} Johann, P. and Polonsky, A. {\em Deep induction:
  Induction rules for (truly) nested types}.  Proceedings, Foundations
  of Software Science and Computation Structures, pp. 339-358, 2020.

\bibitem{mac71} MacLane, S. {\em Catgories for the Working
  Mathematician}. Springer, 1971.

\bibitem{mcb99} McBride, C. {\em Dependently Typed Programs and their
  Proofs}. PhD thesis, University of Edinburgh, 1999.

\bibitem{min15} Minsky, Y.  {\em Why {GADT}s matter for performance}.
  $\mathtt{https://blog.janestreet.com/why-gadts-matter-for-performance/}$,
  2015.

\bibitem{pl04} Pasalic, E., and Linger, N.  {\em Meta-programming with
  typed object-language representations}.  Generic Programming and
  Component Engineering, pp.~136-167, 2004.

\bibitem{pen20} Penner, C.  {\em Simpler and safer {API} design using
  {GADT}s}.  $\mathtt{https://chrispenner.ca/posts/gadt-design}$,
  2020.

\bibitem{pvww06} Peyton Jones, S., Vytiniotis, D., Weirich, S., and
  Washburn, G. {\em Simple unification-based type inference for
    GADTs}. Proceedings, International Conference on Functional
  Programming, 2006. {\color{red} PAGES?}

\bibitem{pr06} Pottier, F., and R{\'e}gis-Gianas, Y.  {\em Stratified
  type inference for generalized algebraic data types}.  Principles of
  Programming Languages, pp.~232-244, 2006.

\bibitem{sjsv09} Schrijvers, T, Peyton Jones, S. L., Sulzmann, M., and
  Vytiniotis, D. {\em Complete and decidable type inference for
    GADTs}. Proceedings, International Conference on Functional
  Programming, pp. 341– 352, 2009.

\bibitem{sp04} Sheard, T., and Pasalic, E. {\em Meta-programming with
  built-in type equality}. Proceedings, Workshop on Logical Frameworks
  and Meta-languaegs, 2004. {\color{red} PAGES?}

\bibitem{tas19} Tassi, E.: Deriving proved equality tests in Coq-elpi:
  Stronger induction principles for containers in Coq. Proceedings,
  Interactive Theorem Proving, pp. 1-18, 2019.

\bibitem{vw10} Vytiniotis, D., and Weirich, S.  {\em Parametricity,
  type equality, and higher-order polymorphism}.  Journal of
  Functional Programming 20(2), pp.~175--210, 2010.

\bibitem{xcc03} Xi, H., Chen, C. and Chen, G. {\em Guarded recursive
  datatype constructors}. Proceedings, Principles of Programming
  Languages, pp. 224–235, 2003.

\bibitem{ull20} Ullrich, M. {\em Generating Induction Principles for
Nested Induction Types in MetaCoq}. PhD thesis, Saarland University,
  2020.

\end{thebibliography}

\end{document}