### 2021-05-17 induction-paper/ENTCS-GADT-induction.tex
final MFPS

\documentclass[9pt]{entcs}
\usepackage{entcsmacro}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{ccicons}
\usepackage{verbatim}
\usepackage{amscd}
\usepackage{MnSymbol}
\usepackage{xcolor}
\usepackage{bbold}
\usepackage{url}
\usepackage{upgreek}
\usepackage{stmaryrd}
\usepackage{lipsum}
\usepackage{tikz-cd}
\usetikzlibrary{cd}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usepackage{bussproofs}
\EnableBpAbbreviations

\sloppy
% The following is enclosed to allow easy detection of differences in
% ascii coding.
% Upper-case    A B C D E F G H I J K L M N O P Q R S T U V W X Y Z
% Lower-case    a b c d e f g h i j k l m n o p q r s t u v w x y z
% Digits        0 1 2 3 4 5 6 7 8 9
% Exclamation   !           Double quote "          Hash (number) #
% Dollar        $           Percent      %          Ampersand     &
% Acute accent  '           Left paren   (          Right paren   )
% Asterisk      *           Plus         +          Comma         ,
% Minus         -           Point        .          Solidus       /
% Colon         :           Semicolon    ;          Less than     <
% Equals        =3D           Greater than >          Question mark ?
% At            @           Left bracket [          Backslash     \
% Right bracket ]           Circumflex   ^          Underscore    _
% Grave accent  `           Left brace   {          Vertical bar  |
% Right brace   }           Tilde        ~

\bibliographystyle{entcs}

\def\lastname{Johann, Ghiorzi, and Jeffries}

\input{macros}

\begin{document}
\begin{frontmatter}
  \vspace*{-0.1in}
  \title{(Deep) Induction Rules for GADTs\vspace*{-0.1in}}
  \author{Patricia Johann~~~~}%\thanksref{JohannEmail}}
  \author{Enrico Ghiorzi~~~~}%\thanksref{GhiorziEmail}}
  \author{Daniel Jeffries}%\thanksref{JeffriesEmail}}
%  \thanks[JohannEmail]{Email: \href{mailto:johannp@appstate.edu} {\texttt{\normalshape johannp@appstate.edu}}}
%  \thanks[GhiorziEmail]{Email: \href{mailto:ghiorzie@appstate.edu} {\texttt{\normalshape ghiorzie@appstate.edu}}}
%  \thanks[JeffriesEmail]{Email: \href{mailto:jeffriesd@appstate.edu} {\texttt{\normalshape jeffriesd@appstate.edu}}}
  \address{$\mathtt{\{johannp,ghiorzie,jeffriesd\}@appstate.edu}$\\
    Appalachian State University}

\vspace*{-0.1in}

\begin{abstract} 
  Deep data types are those that are defined in terms of other such
  data types, including, possibly, themselves.  In that case, they are
  said to be truly nested.  Deep induction is an extension of
  structural induction that traverses {\em all} of the structure in a
  deep data type, propagating predicates on its primitive data
  throughout the entire structure.  Deep induction can be used to prove
  properties of nested types, including truly nested types, that
  cannot be proved via structural induction.  In this paper we show
  how to extend deep induction to deep GADTs that are not truly
  nested. We also show that deep induction cannot be extended to truly
  nested GADTs.\looseness=-1
\end{abstract}

\end{frontmatter}

\vspace*{-0.1in}

\section{Introduction}\label{sec:intro}

Induction is one of the most important techniques available for
working with advanced data types, so it is both inevitable and
unsurprising that it plays an essential role in modern proof
assistants. In the proof assistant Coq~\cite{coq20}, for example,
functions and predicates over advanced types are defined inductively,
and almost all non-trivial proofs of their properties are either
proved by induction outright or rely on lemmas that are.  Every time a
new inductive type is declared in Coq an induction rule is
automatically generated for it.\looseness=-1

The inductive data types handled by Coq include (possibly mutually
inductive) polynomial algebraic data types (ADTs), and the induction
rules Coq generates for them are the expected ones for standard
structural induction. It has long been understood, however, that these
rules are too weak to be genuinely useful for deep ADTs~\cite{jp20},
i.e., ADTs that are (possibly mutually inductively) defined in terms
of (other) such ADTs.\footnote{A data type is {\em deep} if it is
  (possibly mutually inductively) defined in terms of other such data
  types (including, possibly, itself).}  The following data type of
rose trees, here coded in Agda and defined in terms of the standard
type $\mathsf{List}$ of lists (see
Section~\ref{sec:ADTs-and-nesteds}), is a deep ADT:

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{data\, Rose\, : Set \to Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;empty\, :\,\forall \{A : Set\} \to Rose\,A}\\
\mathsf{\;\;\;\;\;\;\;\;\,node\,\,\,\, :\, \forall \{A : Set\} \to A \to List\,(Rose\,A) \to Rose\,A} 
\end{array}\]
\noindent
The induction rule Coq automatically generates for (the analogous Coq
defintion of) rose trees is

\vspace*{-0.075in}

\[\begin{array}{l}
\mathsf{\forall\, (A : Set)\,(P : Rose\,A \to Set) \to P\,empty \to}\\
\mathsf{\hspace*{0.15in}
 (\forall\, (a : A)\,(ts :
  List\,(Rose\,A)) \to P\,(node\,a\,ts)) \to 
  \forall \,(x : Rose\,A) \to P\,x}
\end{array}\]
\noindent
Unfortunately, this is neither the induction rule we intuitively
expect, nor is it expressive enough to prove even basic properties of
rose trees that ought to be amenable to inductive proof. What is
needed here is an enhanced notion of induction that, when specialized
to rose trees, will propagate the predicate $\mathsf{P}$ through the
outer list structure and to the rose trees sitting inside
$\mathsf{node}$'s list argument. More generally, this enhanced notion
of induction should traverse {\em all} of the layers present in a data
structure, propagating suitable predicates to {\em all} of the data it
contains. With data types becoming ever more advanced, and with deeply
structured such types becoming increasingly ubiquitous in
formalizations, it is critically important that proof assistants
%{\color{red} (e.g., Coq, Agda, Lean)}
are able to automatically generate genuinely useful induction rules
for data types that go well beyond traditional ADTs. Such data types
include not just deep ADTs, but also (truly\footnote{A truly nested
  type is a nested type that is defined over itself. The data type
  $\mathsf{Bush}$ in Section~\ref{sec:ADTs-and-nesteds} provides a
  concrete example.}) nested types~\cite{bm98},
%, such as $\mathsf{Rose}$ above, as well as truly nested types
generalized algebraic data types
(GADTs)~\cite{ch03,pvww06,sp04,xcc03}, more richly indexed
families~\cite{ch88}, and deep variants of all of these.

\pagebreak

{\em Deep induction}~\cite{jp20} is a generalization of structural
induction that fits this bill exactly. Whereas structural induction
rules induct over only the top-level structure of data, leaving any
data internal to the top-level structure untouched, deep induction
rules induct over {\em all} of the structured data present. The key
idea is to parameterize induction rules not just over a predicate over
the top-level data type being considered, but also over additional
custom predicates on the types of primitive data they contain. These
custom predicates are then lifted to predicates on any internal
structures containing these data, and the resulting predicates on
these internal structures are lifted to predicates on any internal
structures containing structures at the previous level, and so on,
until the internal structures at all levels of the data type
definition, including the top level, have been so
processed. Satisfaction of a predicate by the data at one level of a
structure is then conditioned upon satisfaction of the appropriate
predicates by {\em all} of the data at the preceding level.

Deep induction was shown in~\cite{jp20} to deliver induction rules
appropriate to deep nested types, including deep ADTs.  For the deep
ADT of rose trees, for example, it gives the following genuinely
useful induction rule:

\vspace*{-0.075in}

\begin{equation}\label{eq:rose}
\begin{array}{l}
\mathsf{\forall \,(A : Set)\,(P : Rose\,A \to Set)\,(Q : A \to Set)
  \to P\,empty \to}\\ 
\mathsf{\hspace*{0.15in}(\forall \,(a : A)\, (ts :
  List\,(Rose\,A))\to Q\,a \to List^\land\,(Rose\,A)\, P\,ts \to P\,(node\,a\,ts)) \to}\\
\mathsf{\hspace*{0.15in}\forall \,(x :
  Rose\,A) \to Rose^\land\,A\,Q\, x \to P\,x} 
\end{array}
\end{equation}

\noindent
%Here, $\mathsf{List^\land}$ lifts its predicate argument $\mathsf{P}$
%on data of type $\mathsf{Rose\,a}$ to a predicate on data of type
%$\mathsf{List\,(Rose\,a)}$ asserting that $\mathsf{P}$ holds for every
%element of its argument list, and similarly for the lifting
%$\mathsf{Rose^\land}$ and the predicate $\mathsf{Q}$.
Here, $\mathsf{List^\land}$ (resp., $\mathsf{Rose^\land}$) lifts its
predicate argument $\mathsf{P}$ (resp., $\mathsf{Q}$) on data of type
$\mathsf{Rose\,A}$ (resp., $\mathsf{A}$) to a predicate on data of
type $\mathsf{List\,(Rose\,A)}$ (resp., $\mathsf{Rose\,A}$) asserting
that $\mathsf{P}$ (resp., $\mathsf{Q}$) holds for every element of its
list (resp., rose tree) argument.\footnote{Predicate liftings such as
  $\mathsf{List^\land}$ and $\mathsf{Rose^\land}$ can either be
  supplied as primitives or generated automatically from their
  associated data type definitions as described in
  Section~\ref{sec:ADTs-and-nesteds} below. The predicate lifting for
  a container type like $\mathsf{List\,t}$ or $\mathsf{Rose\,t}$
  simply traverses containers of that type and applies its predicate
  argument pointwise to the constituent data of type $\mathsf{t}$.}
Deep induction was also shown in~\cite{jp20} to
% be the missing piece making it possible to
%be the theretofore missing component needed to
% be the essential ingredient making it possible to
deliver the first-ever induction rules --- structural or otherwise ---
for the $\mathsf{Bush}$ data type~\cite{bm98} and other truly nested
types. Deep induction for ADTs and nested types is reviewed in
Section~\ref{sec:ADTs-and-nesteds} below.

This paper shows how to extend deep induction to proper GADTs, i.e.,
to GADTs that are not simply nested types, and thus are not ADTs.  A
constructor for such a GADT $\mathsf{G}$ may, like a constructor for a
nested type, take as arguments data whose types involve instances of
$\mathsf{G}$ other than the one being defined --- including instances
that involve $\mathsf{G}$ itself. But if $\mathsf{G}$ is a proper GADT
then at least one of its constructors will also have such a structured
instance of $\mathsf{G}$ --- albeit one not involving $\mathsf{G}$
itself --- as its codomain.
%But while the types of the arguments to a GADT's constructor can
%include instances of $\mathsf{G}$ that themselves involve
%$\mathsf{G}$, its return type cannot.
For example, the constructor $\mathsf{pair}$ for the GADT

\vspace*{-0.1in}

\begin{equation}\label{eq:seq}
\begin{array}{l}
\mathsf{data\, Seq\,: Set \to Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;const :\, \forall \{A : Set\} \to A \to Seq\,A}\\
\mathsf{\;\;\;\;\;\;\;\;pair\;\;\, :\,\forall \{A\,B : Set\} \to Seq \,A \to Seq\,B \to
  Seq\,(A \times B)}
\end{array}
\end{equation}
\noindent
of sequences\footnote{The type of $\mathsf{Seq}$ is actually
  $\mathsf{Set \to \Set_1}$, but to aid readability we elide the
  explicit tracking of Agda universe levels in this paper.} only
constructs sequences of pairs, rather than sequences of arbitrary
type, as does $\mathsf{const}$. If one or more of the constructors for
a GADT $\mathsf{G}$ return structured instances of $\mathsf{G}$, then
the GADT will have two distinct, but equally natural, semantics: a
functorial semantics interpreting it as a left Kan
extension~\cite{mac71}, and a parametric semantics interpreting it as
the interpretation of its Church encoding~\cite{atk12,vw10}. As
detailed in~\cite{jgj21}, a key difference in the two semantics is
that the former views GADTs as their {\em functorial
  completions}~\cite{jp19}, and thus as containing more data than just
those expressible in syntax. By contrast, the latter views them as
what might be called {\em syntax-only} GADTs. Happily, these two views
of GADTs coincide for those GADTs that are ADTs or other nested
types. However, both they and their attendant properties differ
greatly for proper GADTs. In fact, the functorial and parametric
semantics for proper GADTs are sufficiently disparate that, by
contrast with the semantics customarily given for ADTs and other
nested types~\cite{bfss90,gjfor15,jgj21f}, it is not actually possible
to define a functorial parametric semantics for them~\cite{jgj21}.

This observation seems, at first, to be a death knell for the prospect
of extending deep induction to GADTs. Indeed, induction can be seen as
unary parametricity, so we quickly realize that GADTs viewed as their
functorial completions cannot possibly support induction rules.  This
makes sense intuitively: induction is a syntactic proof technique, so
of course it cannot be used to prove properties of those elements of a
GADT's functorial completion that are not expressible in syntax. All
is not lost, however. As we show below, the syntax-only view of GADTs
determined by their Church encodings {\em does} support induction
rules --- including deep induction rules --- for GADTs.  Indeed, this
paper gives the first-ever induction rules --- deep or otherwise ---
for proper GADTs. But it actually delivers far more: it gives a
general framework for deriving deep induction rules for deep GADTs
directly from their syntax. This framework can serve as a basis for
extending modern proof assistants' automatic generation of structural
induction rules for ADTs to automatic generation of deep induction
rules for GADTs. In addition, as for ADTs and other nested types, the
structural induction rule for any GADT can be recovered from its deep
induction rule simply by taking the custom predicates in its deep
induction rule to be constantly $\mathsf{True}$-valued (i.e.,
constantly $\mathsf{\top}$-valued) predicates.

Significantly, deep induction rules for GADTs cannot be derived by
somehow extending the approach of~\cite{jp20} to syntax-only
GADTs. Indeed, the approach taken there makes crucial use of the
functoriality of data types' interpretations from~\cite{jp19}, and
functoriality is precisely what interpreting GADTs as the
interpretations of their Church encodings fails to deliver. Our
approach is to instead first give a predicate lifting styled after
those of~\cite{jp20}, together with a (deep) induction rule, for the
simplest --- and arguably most important --- GADT, namely the equality
GADT. (See Section~\ref{sec:ind-equal}.)  We then derive the deep
induction rule for a more complex GADT $\mathsf{G}$ by {\em i}) using
the equality GADT to represent $\mathsf{G}$ as its so-called {\em
  Henry Ford encoding}~\cite{ch03,hin03,mcb99,sjsv09,sp04}, and {\em
  ii}) using the predicate liftings for the equality GADT, and any
other GADTs appearing in the definition of $\mathsf{G}$, to
appropriately thread the custom predicates for the primitive types
appearing in $\mathsf{G}$ throughout $\mathsf{G}$'s structure. This
two-step process delivers deep induction rules for a very general
class of deep GADTs. In Section~\ref{sec:GADTs} we introduce a series
of increasingly complex GADTs as running examples, and in
Section~\ref{sec:deep-ind-GADTs} we derive a deep induction rule for
each of them. In particular, we derive the deep induction rule for the
$\mathsf{Seq}$ data type in~\eqref{eq:seq} in
Section~\ref{sec:ind-seq}. We present our general framework for
deriving (deep) induction rules for (deep) GADTs in
Section~\ref{sec:framework}, and observe that the derivations in
Section~\ref{sec:deep-ind-GADTs} are all instances of it.  In
Section~\ref{sec:GADT-nested} we show that, by contrast with truly
nested types, which do have a functorial semantics, syntax-only GADTs'
lack of functoriality means that it is not possible to extend
induction --- deep or otherwise --- to truly nested GADTs. This does
not appear to be much of a restriction, however, since GADTs defined
over themselves do not, to our knowledge, appear in practice or in the
literature. Section~\ref{sec:app} comprises a case study in using deep
induction. All of the deep induction rules appearing in this paper
have been derived by instantiating our general framework. Our Agda
code implementing them is available at~\cite{web-page}.

%\subsection{Related Work}

\vspace*{0.05in}

{\bf Additional Related Work\/} Various techniques for deriving
induction rules for data types that go beyond ADTs have been
studied. For example, Fu and Selinger~\cite{fs18} show, via examples,
how to derive induction rules for arbitrary nested
types. Unfortunately, however, their technique is rather {\em ad hoc},
so is unclear how to generalize it to nested types other than the
specific ones studied there. Moreover,~\cite{fs18} actually derives
induction rules for data types {\em related} to the original nested
types rather than for the original nested types themselves, and it is
unclear whether or not the derived rules are sufficiently expressive
to prove all results about the original nested types that we would
expect to be provable by induction. This latter point echoes the issue
with Coq-derived induction rule for rose trees raised in
Section~\ref{sec:intro}, which has the unfortunate effect of forcing
users to manually write induction (and other) rules for such types for
use in that system. Tassi~\cite{tas19} derives induction rules for
data type definitions in Coq using unary parametricity. His technique
seems to be essentially equivalent to that of~\cite{jp19} for nested
types, although he does not permit true nesting. More recently,
Ullrich~\cite{ull20} has implemented a plugin in MetaCoq to generate
induction rules for nested types. This plugin is also based on unary
parametricity and true nesting still is not permitted.  As far as we
know, no attempts have been made to extend either implementation to
truly nested types or proper GADTs or their deep variants. In fact, we
know of no work other than that reported here that specifically
addresses induction rules for such data types.

\section{Deep induction for ADTs and nested types}\label{sec:ADTs-and-nesteds}

A structural induction rule for a data type allows us to prove that if
a predicate holds for every element inductively produced by the data
type's constructors then it holds for every element of the data type.
In this paper, we are interested in induction rules for proof-relevant
predicates.  A proof-relevant predicate on $\mathsf{A : Set}$ is a
function $\mathsf{P\,:\,A \to Set}$ mapping each $\mathsf{a : A}$ to
the set of proofs that $\mathsf{P\,a}$ holds.  For example, the
structural induction rule for the list type

\vspace*{-0.2in}

\begin{equation*}\label{eq:list}
\begin{array}{l}
\mathsf{data\ List : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;nil\,\,\,\,\,\; :\, \forall \{A : Set\} \to List\,A}\\
\mathsf{\;\;\;\;\;\;\;cons\, :\, \forall \{A : Set\} \to A \to List\,A \to List\,A} 
\end{array}
\end{equation*}

\vspace*{-0.1in}

\noindent
is

\vspace*{-0.3in}

\begin{equation*}
\mathsf{
\forall (A : Set) (P : List\,A \to Set)
\to P\,nil
\to \big( \forall (a : A) (as: List\,A)
\to P\,as
\to P\,(cons\,a\,as)\big)
\to \forall (as : List\,A)
\to P\, as
}
\end{equation*}

\vspace*{-0.1in}

\noindent
As in Coq's induction rule for rose trees, the data inside a structure
of type $\mathsf{List}$ is treated monolithically (i.e., is ignored)
by this structural induction rule.  By contrast, the deep induction
rule for lists is parameterized over a custom predicate $\mathsf{Q}$
on $\mathsf{A}$. For $\mathsf{List^\wedge}$ as described in the
introduction the deep induction rule for lists is

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (A : Set) (P : List\, A \to Set) (Q : A \to Set)
\to P\,nil \to \big( \forall (a : A) (as: List\,A) \to Q\,a \to P\,as
\to P\,(cons\,a\,as)\big) \to} \\ 
\quad\mathsf{\forall (as : List\,A) \to List^{\wedge}\,A\,Q\,as
  \to P\,as } 
\end{array}\]

\vspace*{-0.1in}

Structural induction can be extended to nested types, such as the
following type of perfect trees~\cite{bm98}:

\begin{equation*}\label{eq:ptree}
\begin{array}{l}
\mathsf{data\ PTree : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;pleaf\,\,\; :\,\forall \{A : Set\} \to A \to PTree\,A}\\
\mathsf{\;\;\;\;\;\;\;pnode\, :\,\forall \{A : Set\} \to PTree\,(A \times A) \to PTree\,A} 
\end{array}
\end{equation*}
Perfect trees can be thought of as lists constrained to have lengths
that are powers of 2. In the above code, the constructor
$\mathsf{pnode}$ uses data of type $\mathsf{PTree\,(A \times A)}$ to
construct data of type $\mathsf{PTree\,A}$. Thus, it is clear that the
instances of $\mathsf{PTree}$ at various indices cannot be defined
independently, and that the entire inductive family of types must
therefore be defined at once. This intertwinedness of the instances of
nested types is reflected in their structural induction rules, which,
as explained in~\cite{jp20}, must necessarily involve polymorphic
predicates rather than the monomorphic predicates appearing in
structural induction rules for ADTs. The structural induction rule for
perfect trees, for example, is

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to PTree\, A \to Set)
\to \big( \forall (A : Set) (a : A) \to P\,A\,(pleaf\, a) \big) \to} \\
\quad\mathsf{\big( \forall (A : Set) (pp : PTree\,(A \times A)) \to
  P\,(A \times A)\,pp \to P\,A\,(pnode\,pp)\big) \to \forall (A :
  Set) (p : PTree\,A) \to P\,A\,p } 
\end{array}\]

\vspace*{-0.1in}

\noindent
The deep induction rule for perfect trees similarly uses polymorphic
predicates but otherwise follows the now-familiar pattern:

\vspace*{-0.2in}

\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to PTree\,A
  \to Set) \to \big( \forall (A : Set) (Q : A \to Set) (a : A) \to
  Q\,a \to P\,A\,Q\,(pleaf\, a) \big) \to} \\ \quad \mathsf{ \big(
  \forall (A : Set) (Q : A \to Set) (pp : PTree\,(A \times A)) \to
  P\,(A \times A)\,(Pair^{\wedge}\,A\,A\,Q\,Q)\,pp \to
  P\,A\,Q\,(pnode\,pp) \big) \to} \\ \quad \mathsf{\forall (A : Set) (Q
  : A \to Set) (p : PTree\,A) \to PTree^{\wedge}\,A\,Q\,p \to
  P\,A\,Q\,p }
\end{array}\]

\vspace*{-0.075in}

\noindent
Here, $\mathsf{Pair^{\wedge} : \forall (A\; B: Set) \to (A \to Set)
  \to (B \to Set) \to A \times B \to Set}$ lifts predicates
$\mathsf{Q_A}$ on data of type $\mathsf{A}$ and $\mathsf{Q_B}$ on data
of type $\mathsf{B}$ to a predicate on pairs of type $\mathsf{A \times
  B}$ in such a way that $\mathsf{Pair^{\wedge}\,A\,B\,Q_A\,Q_B\,(a,b)
 = Q_A\,a \times Q_B\,b}$. Similarly, $\mathsf{PTree^{\wedge} :
  \forall (A : Set) \to (A \to Set) \to PTree\,A \to Set}$ lifts a
predicate $\mathsf{Q}$ on data of type $\mathsf{A}$ to a predicate on
data of type $\mathsf{PTree\,A}$ asserting that $\mathsf{Q}$ holds for
every element of type $\mathsf{A}$ contained in its perfect tree
argument.

It is not possible to extend structural induction to {\em truly}
nested types, i.e., to nested types whose recursive occurrences appear
below themselves. The quintessential example of such a type is that of
bushes\footnote{To define truly nested types in Agda we must use the
  $\mathsf{NO\_POSITIVITY\_CHECK}$ flag. A similar flag is required in
  Coq.}\cite{bm98}:

\vspace*{-0.1in}

\begin{equation*}\label{eq:bush}
\begin{array}{l}
\mathsf{data\ Bush : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;bnil\,\,\,\,\,\; :\, \forall \{A : Set\} \to Bush\,A}\\
\mathsf{\;\;\;\;\;\;bcons\, :\, \forall \{A : Set\} \to A \to Bush\,(Bush\,A) \to Bush\,A} 
\end{array}
\end{equation*}

\vspace*{0.025in}

\noindent
Even defining a structural induction rule for bushes requires that we
be able to lift the rule's polymorphic predicate argument to
$\mathsf{Bush}$ itself. The more general observation that an induction
rule for any truly nested type must therefore necessarily be a deep
induction rule was, in fact, the original motivation for the
development of deep induction in~\cite{jp20}. The deep induction rule
for bushes is

\vspace*{-0.075in}

\[\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Bush\, A \to Set)
\to \big( \forall (A : Set)\,(Q : A \to Set) \to P\,A\,Q\,bnil \big) \to} \\ 
\quad\mathsf{\big( \forall (A : Set) (Q : A \to Set) (a : A) (bb :
  Bush\,(Bush\,A)) \to Q\,a \to
  P\,(Bush\,A)\,(P\,A\,Q)\,bb \to P\,A\,Q\,(bcons\,a\,bb)
  \big) \to} \\ 
\quad\mathsf{\forall (A : Set) (Q : A \to Set) (b : Bush\,A) \to
  Bush^{\wedge}\,A\,Q\,b \to P\,A\,Q\,b } 
\end{array}\]

\vspace*{-0.05in}

\noindent
Here, $\mathsf{Bush^{\wedge} : \forall (A : Set) \to (A \to Set) \to
  Bush\,A \to Set}$ is the following lifting of a predicate
$\mathsf{Q}$ on data of type $\mathsf{A}$ to a predicate on data of
type $\mathsf{Bush\,A}$ asserting that $\mathsf{Q}$ holds for every
element of type $\mathsf{A}$ contained in its argument bush:

\vspace*{-0.15in}

\begin{align*}
\mathsf{Bush^{\wedge}\,A\,Q\,bnil} &= \mathsf{\top} \\
\mathsf{Bush^{\wedge}\,A\,Q\,(bcons\,a\,bb)} &= \mathsf{Q\,a \times
  Bush^{\wedge}\,(Bush\,A)\,(Bush^{\wedge}\,A\,Q)\,bb}
\end{align*}

\vspace*{-0.05in}

Although a truly nested type admits only a single induction rule, it
is worth noting that for those nested types that do admit distinct
structural induction and deep induction rules, the latter always
generalize the former. Indeed, the structural induction rule for each
such nested type is recoverable from its deep induction rule by taking
the custom predicates on its data of primitive types to be constantly
$\mathsf{\top}$-valued predicates. This instantiation ensures that the
resulting induction rule only inspects the top-level structure of its
argument, rather than the contents of that structure, which is exactly
what structural induction should do.
%A concrete example of such a derivation is given in
%Section~\ref{sec:ind-equal}.

\section{(Deep) GADTs}\label{sec:GADTs}

While a data constructor for a nested type can take {\em as arguments}
data whose types involve instances of that type at indices other than
the one being defined, its return type must still be at the (variable)
type instance being defined. For example, each of $\mathsf{pleaf}$ and
$\mathsf{pnode}$ returns an element of type $\mathsf{PTree\,A}$
regardless of the instances of $\mathsf{PTree}$ appearing in the types
of its arguments. GADTs relax this restriction, allowing their data
constructors both to take as arguments \emph{and return as results}
data whose types involve instances other than the one being
defined. That is, GADTs' constructors' return type instances can, like
that of $\mathsf{pair}$ in~\eqref{eq:seq}, be structured.\looseness=-1
%these instances can be structured.
%
%Applications of GADTs include generic programming, modeling
%programming languages via higher-order abstract syntax, maintaining
%invariants in data structures, and expressing constraints in embedded
%domain-specific languages. GADTs have also been used, e.g., to
%implement tagless interpreters~\cite{pl04,pr06,pvww06}, to improve
%memory performance~\cite{min15}, and to design APIs~\cite{pen20}.

GADTs are used in precisely those situations in which different
behaviors at different instances of data types are desired. This is
achieved by allowing the programmer to give the type signatures of the
GADT's data constructors independently, and then taking advantage of
pattern matching to force the desired type refinement. For example,
the {\em equality} GADT
\begin{equation}\label{eq:equal}
\begin{array}{l}
\mathsf{data\ Equal : Set \to Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;\;\;\;\,refl :\, \forall \{A : Set\} \to Equal\,A\,A}
\end{array}
\end{equation}
is parameterized by two type indices, but it is only possible to
construct data elements of type $\mathsf{Equal\,A\,B}$ if $\mathsf{A}$
and $\mathsf{B}$ are instantiated at the same type. If the types
$\mathsf{A}$ and $\mathsf{B}$ are syntactically identical then the
type $\mathsf{Equal\,A\,B}$ contains the single data element
$\mathsf{refl}$. It contains no data elements otherwise.

The importance of the equality GADT lies in the fact that we can
understand other GADTs in terms of it. For example, the GADT
$\mathsf{Seq}$ from~\eqref{eq:seq} comprises constant sequences of
data of any type $\mathsf{A}$ and sequences obtained by pairing the
data in two already existing sequences. This GADT can be rewritten as
its Henry Ford encoding~\cite{ch03,hin03,mcb99,sjsv09,sp04}, which
makes critical use of the equality GADT, as follows:
\begin{equation}\label{eq:eq_seq}
\begin{array}{l}
\mathsf{data\ Seq : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;const :\, \forall \{A : Set\} \to A \to Seq\,A}\\ 
\mathsf{\;\;\;\;\;\;\;pair\,\;\; :\, \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \times C) \to Seq\,B \to Seq\,C \to Seq\,A}\\ 
\end{array}
\end{equation}
Here, the requirement that $\mathsf{pair}$ produce data at an instance
of $\mathsf{Seq}$ that is a product type is replaced with the
requirement that $\mathsf{pair}$ produce data at an instance of
$\mathsf{Seq}$ that is \emph{equal} to a product type. As we will see
in Section~\ref{sec:deep-ind-GADTs}, this encoding in terms of the
equality GADT is key to deriving deep induction rules for GADTs.

Neither $\mathsf{Equal}$ nor $\mathsf{Seq}$ is a deep GADT, but the
following GADT $\mathsf{LTerm}$, inspired by~\cite{cis194}, is.  It
encodes terms of a simply typed lambda calculus. More robust
variations on $\mathsf{LTerm}$ are, of course, possible, but this
variation is rich enough to illustrate all essential aspects of deep
GADTs --- and later, in Section~\ref{sec:ind-lam}, their deep
induction rules --- while still being small enough to ensure clarity
of exposition.

Types are either booleans, arrow types, or list types. They are
represented by the Henry Ford GADT
\begin{equation}\label{eq:eq_ltype}
\begin{array}{l}
\mathsf{data\ LType : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;bool :\, \forall \{A : Set\} \to \forall (B : Set) \to Equal\,A\,Bool
  \to LType\,A}\\ 
\mathsf{\;\;\;\;\;\;\;\;arr\;\;\; :\, \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to LType\,B \to LType\,C \to LType\,A}\\ 
  \mathsf{\;\;\;\;\;\;\;\;list\,\;\; :\,  \forall \{A : Set\} \to\forall (B : Set) \to
    Equal\,A\,(\List\,B) \to LType\,B \to LType\,A} 
\end{array}
\end{equation}
Terms are either variables, abstractions, applications, or lists of
terms. They are similarly represented by
\begin{equation}\label{eq:eq_lterm}
\begin{array}{l}
\mathsf{data\ LTerm : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;var\,\,\,:\,  \forall \{A : Set\} \to String \to LType\,A \to
  LTerm\,A} \\  
\mathsf{\;\;\;\;\;\;\;\;abs\,\, :\,  \forall \{A : Set\} \to \forall (B\,C : Set) \to
  Equal\,A\,(B \to C) \to String \to LType\,B \to LTerm\,C \to
  LTerm\,A}\\ 
  \mathsf{\;\;\;\;\;\;\;\;app\; :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    LTerm (B \to A) \to LTerm\,B \to LTerm\,A} \\ 
  \mathsf{\;\;\;\;\;\;\;\;list\;\,\, :\,  \forall \{A : Set\} \to \forall (B : Set) \to
    Equal\,A\,(List\,B) \to List\,(LTerm\,B) \to LTerm\,A} 
\end{array}
\end{equation}
The type parameter for $\mathsf{LTerm}$ tracks the types of simply
typed lambda calculus terms. For example, $\mathsf{LTerm\,A}$ is the
type of simply typed lambda terms of type $\mathsf{A}$. Variables are
tagged with their types by the data constructors $\mathsf{var}$ and
$\mathsf{abs}$, whose $\mathsf{LType}$ arguments ensure that their
type tags are legal types. This ensures that all lambda terms
produced by $\mathsf{var}$, $\mathsf{abs}$, $\mathsf{app}$, and
$\mathsf{list}$ are well-typed.  We will revisit these GADTs in
Sections~\ref{sec:deep-ind-GADTs} and~\ref{sec:app}.

\section{(Deep) induction for GADTs}\label{sec:deep-ind-GADTs}

The equality constraints engendered by GADTs' data constructors makes
deriving (deep) induction rules for them more involved than for ADTs
and other nested types. Nevertheless, we show in this section how to
do so. We first illustrate the key components of our approach by
deriving deep induction rules for the three specific GADTs introduced
in Section~\ref{sec:GADTs}. Then, in Section~\ref{sec:framework}, we
abstract these to a general framework that can be applied to any deep
GADT that is not truly nested. As hinted above, the predicate lifting
for the equality GADT plays a central role in deriving both structural
and deep induction rules for more general GADTs.

\subsection{(Deep) induction for $\mathsf{Equal}$}\label{sec:ind-equal}

To define the (deep) induction rule for any (deep) GADT $\mathsf{G}$
we first need to define a predicate lifting that maps a predicate on
a type $\mathsf{A}$ to a predicate on $\mathsf{G\,A}$. Such a predicate
lifting $\mathsf{Equal^{\wedge} : \forall (A\,B : Set) \to (A \to Set)
  \to (B \to Set) \to Equal\,A\,B \to Set}$ for $\mathsf{Equal}$ is
defined by $\mathsf{Equal^{\wedge}\,A\,A\,Q\,Q'\,refl = \forall (a :
  A) \to Equal\,(Q\,a)(Q'\,a)}$.  It takes two predicates on the same
type as input and tests them for extensional equality.
%
Next, we need to associate with each data constructor $\mathsf{c}$ of
$\mathsf{G}$ an {\em induction hypothesis} asserting that, if the
custom predicate arguments to a predicate $\mathsf{P}$ on $\mathsf{G}$
can be lifted to $\mathsf{G}$ itself, then $\mathsf{c}$ {\em respects}
$\mathsf{P}$, i.e., $\mathsf{c}$ constructs data satisfying the
instance of $\mathsf{P}$ at those custom predicates. The following
induction hypothesis $\mathsf{dIndRefl}$ is thus associated with the
$\mathsf{refl}$ constructor for $\mathsf{Equal}$:
\begin{equation*}\label{eq:ind-refl}
\begin{array}{l}
\mathsf{\lambda (P : \forall (A\,B : Set) \to (A \to Set) \to (B \to
  Set) \to Equal\,A\,B \to Set) \to} \\ 
\quad\mathsf{\forall (C : Set) (Q\, Q' : C \to Set) \to
  Equal^{\wedge}\,C\,C\,Q\,Q'\,refl \to P\,C\,C\,Q\,Q'\,refl} 
\end{array}
\end{equation*}

The deep induction rule for $\mathsf{G}$ now states that, if all of
$\mathsf{G}$'s data constructors respect a predicate $\mathsf{P}$,
then $\mathsf{P}$ is satisfied by every element of $\mathsf{G}$ to
which the custom predicate arguments to $\mathsf{P}$ can be
successfully lifted.  The deep induction rule for $\mathsf{Equal}$ is
thus
\begin{equation}\label{eq:ind-equal}
\begin{array}{l}
\mathsf{\forall (P : \forall (A\,B : Set) \to (A \to Set) \to (B \to
  Set) \to Equal\,A\,B \to Set) \to dIndRefl\,P \to}\\ \quad 
  \mathsf{\forall (A\,B : Set) (Q_A : A \to Set) (Q_B : B \to Set) (e:
  Equal\,A\,B) \to Equal^{\wedge}\,A\,B\,Q_A\,Q_B\,e \to
  P\,A\,B\,Q_A\,Q_B\,e}
\end{array}
\end{equation}

To prove that this rule is sound we must provide a witness
$\mathsf{dIndEqual}$ inhabiting the type in~\eqref{eq:ind-equal}.  By
pattern matching, we need only consider the case where $\mathsf{A} =
\mathsf{B}$ and $\mathsf{e} = \mathsf{refl}$, so we can define
$\mathsf{dIndEqual}$ by
$\mathsf{dIndEqual\;P\;crefl\;A\;A\;Q_A\;Q_A'\;refl\;liftE =
  crefl\;A\;Q_A\;Q_A'\,liftE}$. To recover $\mathsf{Equal}$'s structural
induction rule
\begin{equation}\label{eq:sind-equal}
\mathsf{\forall (Q : \forall (A\,B : Set)
  \to Equal\,A\,B \to Set) \to \big( \forall (C : Set) \to
  Q\,C\,C\,refl \big) \to \forall (A\,B : Set) (e: Equal\,A\,B) \to
  Q\,A\,B\,e}
\end{equation}
we define a term $\mathsf{indEqual}$ of the type
in~\eqref{eq:sind-equal} by $\mathsf{indEqual\;Q\;srefl\;A\;B\;refl =
  dIndEqual\;P\;srefl'\;A\;B\;
  K^A_\top\;K^B_\top\;refl\,sliftE}$. Here, $\mathsf{P : \forall (A\,B
  : Set) \to (A \to Set) \to (B \to Set) \to Equal\,A\,B \to Set}$ is
defined by $\mathsf{P\;A\;B\;Q_A\;Q_B\;e =}$ $\mathsf{Q\;A\;B\;e}$,
$\mathsf{K^A_\top}$ and $\mathsf{K^B_\top}$ are the constantly
$\mathsf{\top}$-valued predicates on $\mathsf{A}$ and $\mathsf{B}$,
respectively, $\mathsf{sliftE :}$ $\mathsf{Equal^{\wedge}\;A\;B\;
  K^A_\top\;K^B_\top\;refl}$ is defined by $\mathsf{sliftE\,a = refl :
  Equal\, \top\,\top}$ for every $\mathsf{a : A}$, and $\mathsf{srefl'
  : \forall (C\,: Set) (Q_c\,Q'_c\, : C \to Set) \to
  Equal^{\wedge}\,C\,C\,Q_c\,Q'_c\,refl \to \,Q\,C\,C\,refl}$ is
defined by $\mathsf{srefl'\,C\,Q_c\,Q'_c\,liftE'\,=srefl\, C}$.  The
structural induction rule for any GADT $\mathsf{G}$ that is not truly
nested can similarly be recovered from its deep induction rule by
instantiating every custom predicate by the appropriate constantly
$\mathsf{\top}$-valued predicate.

\subsection{(Deep) induction for $\mathsf{Seq}$}\label{sec:ind-seq}

To derive the deep induction rule for the GADT $\mathsf{Seq}$ we use
its Henry Ford encoding from~\eqref{eq:eq_seq}.  We first define its
predicate lifting $\mathsf{Seq^\wedge : \forall (A : Set) \to (A \to
  Set) \to Seq\,A \to Set}$ by

\vspace*{-0.2in}

\[\begin{array}{lll}
\mathsf{Seq^{\wedge}\,A\,Q_A\,(const\,a)} & = & \mathsf{Q_A\,a}\\
\mathsf{Seq^{\wedge}\,A\,Q_A\,(pair\,B\,C\,e\,s_B\,s_C)}
&=&\mathsf{\exists [Q_B] \exists [Q_C]\, Equal^{\wedge}\,A\, (B
  \times C)\, Q_A\, (Pair^\wedge\,B\,C\, Q_B \, Q_C) \, e \times
  Seq^{\wedge}\,B\,Q_B\,s_B \times Seq^{\wedge}\,C\,Q_C\,s_C}
\end{array}\]
Here, $\mathsf{a : A}$, $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C : C
  \to Set}$, $\mathsf{e : Equal\,A\,(B \times C)}$, $\mathsf{s_B :
  Seq\,B}$, $\mathsf{s_C : Seq\,C}$, and $\mathsf{\exists [x]\, F
  \,x}$ is syntactic sugar for the type of dependent pairs
$\mathsf{(x,b)}$, where $\mathsf{x : A}$, $\mathsf{b : F\, x}$, and
$\mathsf{F : A \to Set}$.

Next, let $\mathsf{dIndConst}$ be the induction hypothesis
\[\mathsf{
\lambda (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to Set) \to
\forall (A : Set) (Q_A : A \to Set) (a : A) \to Q_A\,a \to
P\,A\,Q_A\,(const\,a)}\] associated with the constructor
$\mathsf{const}$, and let $\mathsf{dIndPair}$ be the induction
hypothesis

\pagebreak

\[\begin{array}{l}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \to \\ 
\quad \mathsf{\forall (A\,B\,C : Set) (Q_A : A \to Set) (Q_B : B
  \to Set) (Q_C : C \to Set)(s_B : Seq\,B) (s_C : Seq\,C) (e :
  Equal\,A\,(B \times C)) \to} \\ 
\quad \mathsf{Equal^{\wedge} A\, (B \times C)\, Q_A\,
  (Pair^{\wedge}\,B\,C\,Q_B\,Q_C)\, e \to P\,B\,Q_B\,s_B \to
  P\,C\,Q_C\,s_C \to P\, A\, Q_A\, ( pair\,B\,C\,e\,s_B\,s_C )}
\end{array}\]
associated with the constructor $\mathsf{pair}$. Then the deep
induction rule for $\mathsf{Seq}$ is
\begin{equation}\label{eq:ind-seq}
\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to Seq\,A \to
  Set)} \mathsf{\to dIndConst\,P \to dIndPair\,P \to} \\ \quad
\mathsf{\forall (A : Set)(Q_A : A \to Set)(s_A : Seq\,A) \to
  Seq^{\wedge}\,A\,Q_A\,s_A \to P\,A\,Q_A\,s_A}
\end{array}
\end{equation}

To prove that this rule is sound we provide a witness
$\mathsf{dIndSeq}$ inhabiting the type in~\eqref{eq:ind-seq} as
follows:
\[\begin{array}{lll}
\mathsf{dIndSeq\;P\;cconst\;cpair\;A\;Q_A\;(const\,a)\;liftA}  &
    = &  \mathsf{cconst\;A\;Q_A\;a\;liftA}\\
\mathsf{dIndSeq\,P\,cconst\,cpair\,A\,Q_A\,(pair\,B\,C\,e\,s_B\,s_C)\,
    (Q_B,Q_C, liftE, liftB, liftC)}  & = &
\mathsf{cpair\,A\,B\,C\,Q_A\,Q_B\,Q_C\,s_B\,s_C\,e\,liftE\,p_B\,p_C}
\end{array}\]
In the first clause, $\mathsf{a : A}$, $\mathsf{Q_A : A \to Set}$, and
$\mathsf{liftA : Seq^{\wedge}\,A\,Q_A\,(const\,a) = Q_A\,a}$. In the
second clause we also have $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C : C \to
  Set}$, $\mathsf{e : Equal\,A\,(B \times C)}$, $\mathsf{s_B :
  Seq\,B}$, $\mathsf{s_C : Seq\,C}$, $\mathsf{liftE :
  Equal^{\wedge}\,A\, (B \times C)\, Q_A\, (Pair^\wedge \,B\,C\,Q_B \,
  Q_C) \, e}$, $\mathsf{liftB : Seq^{\wedge}\,B\,Q_B\,s_B}$, and
$\mathsf{liftC :}$ $\mathsf{Seq^{\wedge}\,C\,Q_C\,s_C}$ --- which
together ensure that $\mathsf{(Q_B, Q_C, liftE, liftB, liftC) :
  Seq^{\wedge}\,A\,Q\,(pair\,B\,C\,e\,s_B\,s_C)}$ --- and
$\mathsf{p_B} =\mathsf{dIndSeq\,P\,cconst\,cpair\,B\,Q_B\,s_B\,liftB :
  P\,B\,Q_B\,s_B}$ and $\mathsf{p_C}
=\mathsf{dIndSeq\,P\,cconst\,cpair\,C\,Q_C\,s_C\,liftC :
  P\,C\,Q_C\,s_C}$.

\subsection{(Deep) induction for $\mathsf{LTerm}$}\label{sec:ind-lam}

\begin{figure*}[t]

  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{LType^{\wedge}\,A\,Q_A\,(bool\,B\,e)} & = &\mathsf{\exists
  [Q_B]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(arr\, B\, C\, e\, T_B\, T_C)}
&=&\mathsf{\exists [Q_B] \,\exists [Q_c]\, Equal^{\wedge}\,A\,
  (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e \times
  \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C}\\
\mathsf{LType^{\wedge}\,A\,Q_A\,(list\, B\, e\, T_B)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge}\,A\, (List\, B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times LType^{\wedge}\,B\,Q_B\,T_B}\\[1ex]
\mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A)} & = &
\mathsf{LType^{\wedge}\, A\, Q_A\, T_A}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} &
= & \mathsf{\exists [Q_B]\,\exists [Q_C]\, Equal^{\wedge} \, A\, (B \to
  C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e \times \,
  LType^{\wedge}\, B\, Q_B\, T_B \times \, LTerm^{\wedge}\, C\, Q_C\,
  t_C }\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)} & = &
\mathsf{\exists [Q_B]\, LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \,
  B\, A\, Q_B \, Q_A)\, t_{BA} \times LTerm^{\wedge}\, B\, Q_B\,
  t_B}\\
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (list\, B\, e\, ts)} & = &
\mathsf{\exists [Q_B]\, Equal^{\wedge} \, A\, (List\,B)\, Q_A\,
  (List^{\wedge} \, B\, Q_B) \, e \times List^{\wedge}\, (LTerm\,B) \,
  (LTerm^{\wedge} \, B\, Q_B) \, ts}
\end{array}   \]}

\vspace*{-0.1in}

\caption{Predicate liftings for $\mathsf{LType}$ and
  $\mathsf{LTerm}$}\label{fig:liftings} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

To derive the deep induction rule for the GADT $\mathsf{LTerm}$ we use
its Henry Ford encoding from~\eqref{eq:eq_ltype}
and~\eqref{eq:eq_lterm}. We first define the predicate lifting
$\mathsf{Arr^{\wedge} : \forall (A\, B : Set) \to (A \to Set) \to (B
  \to Set) \to (A \to B) \to Set}$ for arrow types, since arrow types
appear in $\mathsf{LType}$ and $\mathsf{LTerm}$.  It is given by
$\mathsf{Arr^{\wedge}\, A\, B\, Q_A\, Q_B\, f = \forall (a : A) \to
  Q_A\,a \to Q_B\, (f\,a)}$. The predicate liftings
$\mathsf{LType^{\wedge} : \forall (A : Set) \to (A \to Set) \to
  LType\,A \to Set}$ for $\mathsf{LType}$ and $\mathsf{LTerm^{\wedge}
  :}$ $\mathsf{\forall (A : Set) \to (A \to Set) \to LTerm\,A \to
  Set}$ for $\mathsf{LTerm}$ are defined in Figure~\ref{fig:liftings}.
There, $\mathsf{s : String}$, $\mathsf{Q_A : A \to Set}$, $\mathsf{Q_B
  : B \to Set}$, $\mathsf{Q_C : C \to Set}$,
$\mathsf{K^{Bool}_{\top}}$ is the constantly $\mathsf{\top}$-valued
predicate on $\mathsf{Bool}$, $\mathsf{T_A : LType\, A}$, $\mathsf{T_B
  : LType \,B}$, $\mathsf{T_C : LType \,C}$, $\mathsf{t_B : LTerm \,
  B}$, $\mathsf{t_C : LTerm \, C}$, and $\mathsf{t_{BA} : LTerm \, (B
  \to A)}$.  Moreover, $\mathsf{e : Equal\,A\,Bool}$ in the first
clause, $\mathsf{e : Equal\, A\, (B \to C)}$ in the second, $\mathsf{e
  : Equal\, A\, (List\, B)}$ in the third, $\mathsf{e : Equal \, A \,
  (B \to C)}$ in the fifth, and $\mathsf{e : Equal\, A\, (List \,B)}$,
$\mathsf{ts : List\, (LTerm B)}$, and $\mathsf{List^\wedge}$ is the
predicate lifting for lists from~\eqref{eq:rose} in the seventh.

With these liftings in hand we can define the induction hypotheses
$\mathsf{dIndVar}$, $\mathsf{dIndAbs}$, $\mathsf{dIndApp}$, and
$\mathsf{dIndList}$ associated with $\mathsf{LTerms}$'s data
constructors. These are, respectively,
\[\begin{array}{l}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A  \to Set) \to}\\
  \quad\mathsf{\forall (A : Set) (Q_A : A \to Set) (s : String) (T_A :
  LType\, A) \to LType^{\wedge} \, A\, Q_A\, T_A \to P \, A\, Q_A\,
  (var \; s\, T_A)}\\[1ex]
  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to
  LTerm\,A \to Set) \to} \\ 
  \quad\mathsf{
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C
  \to Set) (e : Equal\, A\, (B \to C)) (s : String) \to } \\ 
  \quad\mathsf{(T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e \to  } \\
  \quad\mathsf{
  LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)
  }\\[1ex]
\end{array}\]
\[\begin{array}{l}
  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set)\to} \\ 
  \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B) \to} \\
  \quad \mathsf{
  P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }\\[1ex]
  \mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
    \to Set) \to} \\ 
  \quad \mathsf{
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B)) \to} \\ 
  \quad \mathsf{
    Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (list \,B \,e \, ts) }
\end{array}\]

\noindent
The deep induction rule for $\mathsf{LTerm}$ is thus

\vspace*{-0.225in}

\begin{equation}\label{eq:ind-lam}
\begin{array}{l}
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A
  \to Set) \to dIndVar\,P \to dIndAbs\,P \to dIndApp\,P \to
  dIndList\,P \to} \\ \quad \mathsf{\forall (A : Set)(Q_A : A \to
  Set)(t_A : LTerm\,A) \to LTerm^{\wedge}\,A\,Q_A\,t_A \to
  P\,A\,Q_A\,t_A}
\end{array}
\end{equation}

\begin{figure*}[t]

  \begin{adjustbox}{varwidth=6.8in, max width=6.6in, margin=-0.1in 0in
      -0.2in 0in, fbox, center} 
{\small
\[\begin{array}{lll}
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (var\;s\,T_A) \, liftA} & = & \mathsf{cvar \, A\, Q_A\, s\, T_A\,
  liftA}\\ 
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
  (abs \,B \,C \,e \,s \,T_B \, t_C) \, (Q_B , Q_C , liftE,
  lift_{T_B}, lift_{t_C})} & = & \mathsf{cabs\,A\,B\,C\, Q_A\,
  Q_B\, Q_C\, e\, s\, T_B\, t_C\, liftE\, lift_{T_B}\, p_C}\\
\mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (app \,B \,\,t_{BA} \, t_B)\, (Q_B , lift_{t_{BA}}, lift_{t_B})} &
= & \mathsf{capp\,A\,B\,Q_A\, Q_B\, t_{BA}\, t_B\, p_{BA} \, p_B}\\
  \mathsf{dIndLTerm \, P\, cvar \, cabs\, capp\, clist \, A\, Q_A\,
    (list \,B \,e \, ts) \, (Q_B , liftE', lift_{List})} & = & 
  \mathsf{clist \,A\,B\,Q_A\, Q_B\, e\, ts\, liftE'\, p_{List} }
\end{array}\]}

\vspace*{-0.1in}

\caption{$\mathsf{dIndLTerm}$}\label{fig:dindlterm} \vspace*{0.1in} 
\end{adjustbox}
\end{figure*}

\vspace*{-0.125in}

To prove this rule is sound we define a witness
$\mathsf{dIndLTerm}$ inhabiting the type in~\eqref{eq:ind-lam} as in
Figure~\ref{fig:dindlterm}. There, $\mathsf{s : String}$, $\mathsf{Q_A
  : A \to Set}$, $\mathsf{Q_B : B \to Set}$, $\mathsf{Q_C : C \to
  Set}$, $\mathsf{T_A : LType\,A}$, $\mathsf{T_B : LType\,B}$,
$\mathsf{t_B : LTerm\,B}$, $\mathsf{t_C : LTerm\,C}$, $\mathsf{t_{BA}
  : LTerm\,(B \to A)}$, $\mathsf{liftA : LTerm^{\wedge}\, A\,
  Q_A\, (var\;s\,T_A) = LType^{\wedge}\,A\,Q_A\,T_A}$, $\mathsf{liftE
  : Equal^{\wedge} \, A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\,
  Q_B \, Q_C) \, e}$, $\mathsf{lift_{T_B}: LType^{\wedge} \, B\, Q_B\,
  T_B}$, $\mathsf{lift_{t_C}: LTerm^{\wedge} \, C\, Q_C\,
  T_C}$, $\mathsf{lift_{t_{BA}}: LTerm^{\wedge} \, (B \to A)\,
  (Arr^{\wedge} \, B\, A\, Q_B \, Q_A)\,
  t_{BA}}$, $\mathsf{lift_{t_B}: LTerm^{\wedge} \, B\, Q_B\,
  t_B}$, $\mathsf{liftE' : Equal^{\wedge}\, A\, (List\,B)\, Q_A\,
  (List^{\wedge}\, B\, Q_B)\, e}$, and $\mathsf{lift_{List}:
  List^{\wedge} \, (LTerm\, B) \, (LTerm^{\wedge}\, B\, Q_B) \, ts}$.
Moreover,

\vspace*{-0.225in}

\[\begin{array}{lll}
\mathsf{p_C} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  C\, Q_C\, t_C\, lift_{t_C} : P \, C\, Q_C \, t_C }\\
\mathsf{p_B} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp \,clist\,
  B\, Q_B\, t_B\, lift_{t_B} : P \, B\, Q_B \, t_B }\\
\mathsf{p_{BA}} & = & \mathsf{dIndLTerm\,P\,cvar\,cabs \,capp
  \,clist\, (B \to A)\,(Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \,
  t_{BA}\, lift_{t_{BA}} : P \, (B \to A)\, (Arr^{\wedge} \, B\, A\,
  Q_B \, Q_A) \, t_{BA}}\\ 
\mathsf{p_{List}} & = &\mathsf{liftListMap \, (LTerm\, B) \,
  (LTerm^{\wedge} \, B \, Q_B)\, (P\,B\,Q_B)\, p_{ts} \, ts\,
  lift_{List} : List^{\wedge}\, (LTerm\,B) \, (P\,B\,Q_B) \, ts}\\
\mathsf{p_{ts}} & = & \mathsf{dIndLTerm\, P\, cvar\, cabs\, capp\,
  clist\, B\, Q_B : PredMap\,(LTerm\,B) \,(LTerm^{\wedge}\, B\, Q_B)
  \, (P\,B\,Q_B)}
\end{array}\]

\vspace*{-0.1in}

\noindent
where, in the final clause, $\mathsf{PredMap : \forall\, (A : Set) \to
  (A \to Set) \to (A \to Set) \to Set }$ is the type constructor
producing the type of morphisms between predicates defined by
$\mathsf{PredMap \,A\, Q\,Q'\, = \forall\, (a : A) \to Q\,a \to
  Q'\,a}$ and $\mathsf{liftListMap : \forall\, (A : Set) \to (Q \, Q'
  : A \to Set) \to PredMap\,A\,Q\,Q' \to PredMap\,(List\,A)
  \,(List^{\wedge}\, A\, Q)\, (List^{\wedge}\, A\, Q')}$, which takes
a morphism $\mathsf{f}$ of predicates and produces a morphism of
lifted predicates, is defined by $\mathsf{liftListMap\, A\, Q\, Q'\,
  m\, nil\, tt = tt}$ (since $\mathsf{x : List^{\wedge}\, A\, Q\,
  nil}$ must necessarily be the sole inhabitant $\mathsf{tt}$ of
$\mathsf{\top}$), and by $\mathsf{liftListMap\, A\, Q\, Q'\, m\,
  (cons\, a\, l')\, (y, x') = (m\,a\,y, liftListMap\, A\, Q\, Q'\, m\,
  l'\, x')}$ (since $\mathsf{x : List^{\wedge}\, A\, Q\, (cons\, a\,
  l')}$ must be of the form $\mathsf{x = (y, x')}$ where
$\mathsf{y : Q\,a}$ and $\mathsf{x' : List^{\wedge}\, A\, Q\, l'}$).

\vspace*{-0.05in}

\section{The general framework}\label{sec:framework}

We can generalize the approach taken in
Section~\ref{sec:deep-ind-GADTs} to a general framework for deriving
deep induction rules for deep GADTs. We will treat deep GADTs of the
form

\vspace*{-0.05in}

\begin{equation}\label{eq:gadts}
\begin{array}{l}
  \mathsf{data\ G : Set^\alpha
    \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;c\, :\, \forall \{\ol{B : Set}\} \to F\,G\,\ol{B} \to G (\ol{K\,\ol{B}})}
\end{array}
\end{equation}

\noindent
For brevity and clarity we indicate only one constructor $\mathsf{c}$
in~\eqref{eq:gadts}, even though a GADT can have any finite number of
them, each with a type of the same form as
$\mathsf{c}$'s. In~\eqref{eq:gadts}, $\mathsf{F}$ and each
$\mathsf{K}$ in $\ol{\mathsf{K}}$ are type constructors with
signatures $\mathsf{(Set^{\alpha} \to Set) \to Set^{\beta} \to Set}$
and $\mathsf{Set^{\beta} \to Set}$, respectively. If $\mathsf{T}$ is a
type constructor with signature $\mathsf{Set^{\gamma} \to Set}$ then
$\mathsf{T}$ has {\em arity} $\mathsf{\gamma}$.  The
overline notation denotes a finite list whose length is exactly the
arity of the type constructor being applied to it. The number of type
constructors in $\ol{\mathsf{K}}$ (resp., $\ol{\mathsf{B}}$) is thus
$\alpha$ (resp., $\beta$). In addition, the type constructor
$\mathsf{F}$ must be constructed inductively according to the
following grammar:
\[\mathsf{F\,G\,\ol{B}\; :=\;
F_1\,G\,\ol{B} \times F_2\,G\,\ol{B} \ \vert\ F_1\,G\,\ol{B} +
F_2\,G\,\ol{B} \ \vert\ F_1\,\ol{B} \to F_2\,G\,\ol{B}
\ \vert\ G\,(\ol{F_1\,\ol{B}}) \ \vert\ H\,\ol{B} \ \vert\ H\,
(\ol{F_1\,G\,\ol{B}}) }\]

\noindent
This grammar is subject to the following restrictions. In the third
clause the type constructor $\mathsf{F_1}$ does not contain
$\mathsf{G}$. In the fourth clause, none of the $\mathsf{\alpha}$-many
type constructors in $\mathsf{\ol{F_1}}$ contains $\mathsf{G}$.  This
prevents nesting, which would make it impossible to give an induction
rule for $\mathsf{G}$; see Section~\ref{sec:GADT-nested} below. In the
fifth clause, $\mathsf{H}$ does not contain $\mathsf{G}$. This clause
therefore subsumes the cases in which $\mathsf{F\,G\,\ol{B}}$ is a
closed type or one of the $\mathsf{B_i}$. In the sixth clause,
$\mathsf{H : Set^\gamma \to Set}$ does not contain $\mathsf{G}$
(although $\ol{\mathsf{F_1\,G\,\ol{B}}}$ can). Moreover, although
$\mathsf{H}$ can construct any (truly) nested type, it must not
construct a GADT. This ensures that $\mathsf{H}$ admits functorial
semantics~\cite{jp20}, and thus has an associated map function. From
the map function for $\mathsf{H}$ we can also construct a map function

\vspace*{-0.05in}

\begin{equation}\label{eq:hliftmap}
\mathsf{H^\wedge Map : \forall (\ol{A : Set}) (\ol{Q\;Q' : A \to Set}) 
\to \ol{PredMap\,A\,Q\,Q'} \to
PredMap\,(H\,\ol{A})\,(H^{\wedge}\,\ol{A}\,\ol{Q})\, 
(H^{\wedge}\,\ol{A}\,\ol{Q'})}
\end{equation}

\vspace*{0.05in}

\noindent
for $\mathsf{H^{\wedge}}$. A concrete way to define $\mathsf{H^\wedge
  Map}$ is by induction on the structure of the type $\mathsf{H}$, but
we omit such details since they are not essential to the present
discussion. A further requirement that applies to all of the type
constructors appearing in the right-hand side of the above grammar,
including those in $\ol{\mathsf{K}}$, is that they must all admit
predicate liftings. This is not an overly restrictive condition,
though: all types constructed from sums, products, arrow types and
type application admit predicate liftings, and so do GADTs constructed
from the grammar. In fact, we have seen such liftings for products and
type application in Section~\ref{sec:deep-ind-GADTs}. A concrete way
to define more general predicate liftings is, again, by induction on
the structure of the types. We do not give a general definition of
predicate liftings here, though, since that would require us to first
design a full type calculus, which is beyond the scope of the present
paper.

We assume in the development below that $\mathsf{G}$ is a unary type
constructor, i.e., that $\alpha = 1$ in~\eqref{eq:gadts}. Extending
the argument to GADTs of arbitrary arity presents no difficulty other
than heavier notation. In this case the type of $\mathsf{G}$'s single
data constructor $\mathsf{c}$ can be rewritten as $\mathsf{c : \forall
  (\ol{B : Set}) \to Equal\,A\,(K\,\ol{B}) \to F\,G\,\ol{B} \to
  G\,A}$. The predicate lifting $\mathsf{G^{\wedge} : \forall (A :
  Set) \to (A \to Set) \to G\,A \to Set}$ for $\mathsf{G}$ is
therefore

\vspace*{-0.05in}

\[\mathsf{G^{\wedge}\,A\,Q_A\,(c\,\ol{B}\,e\,x)
= \exists [\ol{Q_B}]\,
Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
\times F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}\]

\noindent
where $\mathsf{Q_A : A \to Set}$, $\ol{\mathsf{Q_B : B \to Set}}$,
$\mathsf{e : Equal\,A\,(K\,\ol{B})}$, and $\mathsf{x : F\,G\,\ol{B}}$.
If we have predicate liftings $\mathsf{F^{\wedge}} : \mathsf{\forall
  (G : Set^{\alpha} \to Set) (\ol{B : Set}) \to (\forall (A : Set) \to
  (A \to Set) \to G\,A \to Set) \to (\ol{B \to Set}) \to F\,G\,\ol{B}
  \to Set}$ for $\mathsf{F}$ and $\mathsf{K^{\wedge}} :
\mathsf{\forall (\ol{B : Set}) \to (\ol{B \to Set}) \to K\,\ol{B} \to
  Set}$
%\[\begin{array}{lll}
%\mathsf{F^{\wedge}} & : & \mathsf{\forall (G : Set^{\alpha} \to Set) (\ol{B : Set})
%\to (\forall (A : Set) \to (A \to Set) \to G\,A \to Set)
%\to (\ol{B \to Set}) \to F\,G\,\ol{B} \to Set}\\ 
%\mathsf{K^{\wedge}} & : & \mathsf{\forall (\ol{B : Set}) \to (\ol{B \to Set}) \to
%  K\,\ol{B} \to Set}
%\end{array}\]
for $\mathsf{K}$, then the induction hypothesis $\mathsf{dIndC}$
associated with $\mathsf{c}$ is
\[\begin{array}{l}
\mathsf{dIndC = \lambda (P : \forall (A : Set) \to (A \to Set) \to
  G\,A \to Set) \to}\\
\quad \mathsf{\forall (A : Set)\, (\ol{B : Set})\, (Q_A : A \to Set)\,
  (\ol{Q_B : B \to Set})\, (e : Equal\,A\,(K\,\ol{B}))\, (x :
  F\,G\,\ol{B}) \to} \\ 
\quad \mathsf{
  Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
  \to F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x \to
  P\,A\,Q_A\,(c\,\ol{B}\,e\,x)} 
\end{array}\]
and the induction rule for $\mathsf{G}$ is
\[
\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to G\,A \to
  Set) \to dIndC\,P \to \forall (A : Set)(Q_A : A \to Set)(y : G\,A)
\to G^{\wedge}\,A\,Q_A\,y \to P\,A\,Q_A\,y} \]

To prove that this rule is sound we define a witness
$\mathsf{dIndG}$ inhabiting this type by
\[\mathsf{dIndG\,P\,cc\,A\,Q_A\,(c\,\ol{B}\,e\,x)\,(\ol{Q_B}, liftE, liftF)
= cc\,A\,\ol{B}\,Q_A\,\ol{Q_B}\,e\,x\,liftE\,(p\,x\,liftF)}\] Here,
$\mathsf{cc : dIndC\,P}$, $\mathsf{e : Equal\,A\,(K\,\ol{B})}$,
$\mathsf{x : F\,G\,\ol{B}}$, $\mathsf{Q_A : A \to Set}$,
$\ol{\mathsf{Q_B : B \to Set}}$, $\mathsf{liftE :
  Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,
  (K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e}$, and $\mathsf{liftF :
  F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}$. As a result,
$\mathsf{(\ol{Q_B}, liftE, liftF) : G^{\wedge}\,A\,Q_A (c\,\ol{B}\,
  e\,x)}$ as expected.  Finally, the morphism of predicates $\mathsf{p
  :
  PredMap\,(F\,G\,\ol{B})\,(F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B})
  (F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}$ is defined by structural
induction on $\mathsf{F}$ as follows:

\begin{itemize}
\item 
If $\mathsf{F\,G\,\ol{B} = F_1\,G\,\ol{B} \times F_2\,G\,\ol{B}}$ then
$\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} =
  Pair^{\wedge}\,(F_1\,G\,\ol{B})\, (F_2\,G\,\ol{B})\,
  (F_1^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})\,
  (F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}$. The induction hypothesis
  ensures
%{\color{red} liftings $\mathsf{F_1}$ and $\mathsf{F_2}$ (with
%  types) and}
morphisms of predicates $\mathsf{p_1 :
  PredMap\,(F_1\,G\,\ol{B})\,(F_1^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_BQ})
  (F_1^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}$ and $\mathsf{p_2 :
  PredMap\,(F_2\,G\,\ol{B})\,(F_2^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B})
  (F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}$.  For $\mathsf{x_1 :
  F_1\,G\,\ol{B}}$, $\mathsf{liftF_1 :
  F_1^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x_1}$, $\mathsf{x_2 :
  F_2\,G\,\ol{B}}$ and $\mathsf{liftF_2 :
  F_2^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x_2}$ we therefore
define $\mathsf{p\, (x_1, x_2)\, (liftF_1, liftF_2) =
  (p_1\,x_1\,liftF_1,\, p_2\,x_2\,liftF_2)}$.
\item The case $\mathsf{F\,G\,\ol{B} = F_1\,G\,\ol{B} +
  F_2\,G\,\ol{B}}$ is analogous.
\item If $\mathsf{F\,G\,\ol{B} = F_1\,\ol{B} \to F_2\,G\,\ol{B}}$ then
  $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x = \forall (z :
  F_1\,\ol{B}) \to F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z \to
  F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,(x\,z)}$, where $\mathsf{x :
  F\,G\,\ol{B}}$. The induction hypothesis ensures
%  {\color{red}
% liftings $\mathsf{F_1}$ and $\mathsf{F_2}$ (with types) and}
a morphism of predicates $\mathsf{p_2 : PredMap\,(F_2\,G\,\ol{B})\,
  (F_2^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B})\,
  (F_2^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}$.  We define
$\mathsf{p\;x\;liftF : F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x}$, where
$\mathsf{liftF : F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x}$, to
be $\mathsf{p\;x\;liftF\;z\;liftF_1 = p_2\, (x\,z)\,
  (liftF\,z\,liftF_1)}$ for $\mathsf{z : F_1\,\ol{B}}$ and
$\mathsf{liftF_1 : F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z}$. Note that
$\mathsf{F_1}$ not containing $\mathsf{G}$ is a necessary restriction
since the proof relies on
$\mathsf{F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}\,x}$ and
$\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x}$ having the same
domain $\mathsf{F_1^{\wedge}\,\ol{B}\,\ol{Q_B}\,z}$.
\item If $\mathsf{F\,G\,\ol{B} = G\,(F_1\,\ol{B})}$ and $\mathsf{F_1}$
  does not contain $\mathsf{G}$, then
  $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} = P \,(F_1\,\ol{B})\,
    (F_1^{\wedge}\,\ol{B}\,\ol{Q_B})}$ for all $\mathsf{P : \forall (A
    : Set)\to(A \to Set) \to G\,A \to Set}$.  We therefore define
  $\mathsf{p = dIndG\,P\,
    cc\,(F_1\,\ol{B})\,(F_1^{\wedge}\,\ol{B}\,\ol{Q_B})}$.
\item If $\mathsf{F\,G\,\ol{B} = H\,\ol{B}}$ and $\mathsf{H}$ does not
  contain $\mathsf{G}$, then
  $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} = H^\wedge\,\ol{Q_B}}$
  for all $\mathsf{P : \forall (A : Set)\to(A \to Set) \to G\,A \to
    Set}$.  We therefore define $\mathsf{p : PredMap\, (H\,\ol{B})\,
    (H^{\wedge}\,\ol{B}\,\ol{Q_B})\, (H^{\wedge}\,\ol{B}\,\ol{Q_B})}$
  to be the identity morphism on predicates.
\item If $\mathsf{F\,G\,\ol{B} = H\, (\ol{F_k\,G\,\ol{B}})}$ and
  $\mathsf{H}$ does not contain $\mathsf{G}$, then
  $\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B} = H^{\wedge}\,
  (\ol{F_k\,G\,\ol{B}})\,
  (\ol{F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}})}$ for all $\mathsf{P :
  \forall (A : Set)\to(A \to Set) \to G\,A \to Set}$.  Since
  $\mathsf{H}$ is not a GADT, $\mathsf{H^\wedge}$ has a map function
  $\mathsf{H^\wedge Map}$ as in~\eqref{eq:hliftmap}.  The induction
  hypothesis ensures morphisms of predicates $\ol{\mathsf{p_k :
      PredMap\,(F_k\,G\,\ol{B})
      (F_k^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B})
      (F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B})}}$. We therefore define
  $\mathsf{p = H^\wedge Map\,(\ol{F_k\,G\,\ol{B}})\,
    (\ol{F_k^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,\ol{Q_B}})\,
    (\ol{F_k^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}})\,\ol{p_k}}$.
\end{itemize}

\section{Truly Nested GADTs Need Not Admit Deep Induction
    Rules}\label{sec:GADT-nested}

In Sections~\ref{sec:deep-ind-GADTs} and~\ref{sec:framework} we
derived deep induction rules for GADTs that are not truly
nested. Since both nested types and GADTs without true nesting admit
deep induction rules, we might expect truly nested GADTs to admit them
as well. Surprisingly, however, they do not. That is, our results from
the previous section are the strongest possible. Indeed, the induction
rule for a data type generally relies on (unary) parametricity of the
model interpreting it, and deep induction for a truly nested type or a
truly nested GADT also relies on this interpretation being functorial.
But, whereas ADTs and nested types both admit functorial parametric
semantics, proper GADTs cannot admit both functorial and parametric
semantics at the same time~\cite{jgj21}. In this section we show how
deep induction for truly nested GADTs goes wrong by analyzing the
following very simple nested proper GADT:\looseness=-1

\vspace*{-0.1in}

\begin{equation*}\label{gadt-nested}
\begin{array}{l}
\mathsf{data\ G : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;\;c :\,  \forall \{A : Set\} \to G\,(G\,A) \to G\,(A \times A)}
\end{array}
\end{equation*}
The constructor $\mathsf{c}$ can be rewritten as $\mathsf{c :
  \forall\, (B : Set) \to Equal\,A\,(B \times B) \to G\,(G\,B) \to
  G\,A}$, so the predicate lifting $\mathsf{G^{\wedge} : \forall\, (A
  : Set) \to (A \to Set) \to G\,A \to Set}$ for $\mathsf{G}$ is
\[
\mathsf{G^{\wedge}\,A\,Q\,(c\,B\,e\,x)
= \exists\, [Q_A]\,
Equal^{\wedge}\,A\,(B \times B)\,Q_A\,(Pair^{\wedge}\,B\,B\,Q_B\,Q_B)\,e
\times G^{\wedge}\,(G\,B)\,(G^{\wedge}\,B\,Q_B)\,x}
\]
where $\mathsf{Q_A : A \to Set}$, $\mathsf{Q_B : B \to Set}$, $\mathsf{e
  : Equal\,A\,(B \times B)}$, and $\mathsf{x : G\,(G\,B)}$.
The induction hypothesis $\mathsf{dIndC}$ for $\mathsf{c}$ is
\[\begin{array}{l}
\mathsf{\lambda\, (P : \forall\, (A : Set) \to (A \to Set) \to G\,A
  \to Set)\to} \\ 
\quad\mathsf{ \forall\, (A\;B : Set)\, (Q_A : A \to Set)\, (Q_B : B
  \to Set)\, (e : Equal\,A\,(B \times B))\, (x : G\,(G\,B))\to} \\ 
\quad\mathsf{Equal^{\wedge}\,A\,(B \times
  B)\,Q_A\,(Pair^{\wedge}\,B\,B\,Q_B\,Q_B)\,e \to P\,(G\,B)\,(P\,B\,Q_B)\,x 
	\to P\,A\,Q_A\,(c\,B\, e\,x)} 
\end{array}\]
so the deep induction rule for $\mathsf{G}$ is
\[\mathsf{\forall\, (P : \forall\, (A : Set) \to (A \to Set) \to G\,A \to Set)
\to dIndC\,P \to \forall\, (A : Set)\, (Q : A \to Set)\, (y : G\,a)
\to G^{\wedge}\,A\,Q\,y \to P\,A\,Q\,y}\] But if we now try to show
that this rule is sound by constructing a witness $\mathsf{dIndG}$
inhabiting this type we run into problems. We can define
$\mathsf{dIndG\,P\,cc\,A\,Q\,(c\,B\,e\,x)\,(Q', liftE, liftG) =
  cc\,A\,B\,Q\,Q'\,e\,x\,liftE\,p}$, where $\mathsf{cc : dIndC\,P}$,
$\mathsf{Q : A \to Set}$, $\mathsf{Q' : B \to Set}$, $\mathsf{e :
  Equal\,A\,(B \times B)}$, $\mathsf{x : G\,(G\,B)}$, $\mathsf{liftG :
  G^{\wedge}\,(G\,B)\,(G^{\wedge}\,B\,Q')\,x}$, and $\mathsf{liftE :
  Equal^{\wedge}\,A\,(B \times
  B)\,Q\,(Pair^{\wedge}\,B\,B\,Q'\,Q')\,e}$, but we still need to
define $\mathsf{p : P\,(G\,B)\,(P\,B\,Q')\,x}$.  For this we can use
the induction rule and let $\mathsf{p =
  dIndG\,P\,cc\,(G\,B)\,(P\,B\,Q')\,x\,q}$, but we still need to
define $\mathsf{q : G^{\wedge}\,(G\,B)\,(P\,B\,Q')\,x}$.  If we had
the map function $\mathsf{G^\wedge Map : \forall\, (A : Set)\, (Q\;Q'
  : A \to Set) \to PredMap\,A\,Q\,Q' \to
  PredMap\,(G\,A)\,(G^{\wedge}\,A\,Q)\,(G^{\wedge}\,A\,Q')}$ for
$\mathsf{G^{\wedge}}$, then we could define $\mathsf{q = G^\wedge
  Map\,(G\,B)\,(G^{\wedge}\,B\,Q')\,(P\,B\,Q')\,
  (dIndG\,P\,cc\,B\,Q')\,x\,liftG}$.  Unfortunately, however, we
cannot define $\mathsf{G^\wedge Map}$. Indeed, its definition would
have to be $ \mathsf{G^\wedge Map\,A\,Q\,Q'\,m\,(c\,B\,e\,x)\,(Q_B,
  liftE, liftG) = (Q'_B, liftE', liftG')}$ for some $\mathsf{Q'_B : B
  \to Set}$, $\mathsf{liftE' : Equal^{\wedge}\,A\,(B \times
  B)\,Q'\,(Pair^{\wedge}\,B\,B\,Q'_B\,Q'_B)\,e}$, and $\mathsf{liftG'
  : G^{\wedge}\,(G\,B)\,(G^{\wedge}\,B\,Q'_B)\,x}$, where $\mathsf{Q :
  A \to Set}$, $\mathsf{Q' : A \to Set}$, $\mathsf{Q_B : B \to Set}$,
$\mathsf{m : PredMap\,A\,Q\,Q'}$, $\mathsf{e : Equal\,A\,(B \times
  B)}$, $\mathsf{x : G\,(G\,B)}$, $\mathsf{liftE :
  Equal^{\wedge}\,A\,(B \times
  B)\,Q\,(Pair^{\wedge}\,B\,B\,Q_B\,Q_B)\,e}$, and $\mathsf{liftG :
  G^{\wedge}\,(G\,B)\,(G^{\wedge}\,B\,Q_B)\,x}$. In other words, we
would need to produce a proof $\mathsf{liftE'}$ of the (extensional)
equality of the predicates $\mathsf{Q'}$ and
$\mathsf{Pair^{\wedge}\,B\,B\,Q'_B\,Q'_B}$ from just a proof
$\mathsf{liftE}$ of the (extensional) equality of the predicates
$\mathsf{Q}$ and $\mathsf{Pair^{\wedge}\,B\,B\,Q_B\,Q_B}$ and a
morphism of predicates $\mathsf{m}$ from $\mathsf{Q}$ to
$\mathsf{Q_2'}$.
%a proof $\mathsf{liftE}$ of the (extensional) equality of the
%predicates $\mathsf{Q}$ and $\mathsf{Pair^{\wedge}\,B\,B\,Q_1\,Q_1}$
%and a morphism of predicates $\mathsf{m}$ from $\mathsf{Q}$ to
%$\mathsf{Q_2}$, and we need to use those data to deduce a proof of the
%(extensional) equality of the predicates $\mathsf{Q_2}$ and
%$\mathsf{Pair^{\wedge}\,B\,B\,Q_3\,Q_3}$, for some predicate
%$\mathsf{Q_3}$ on $\mathsf{B}$.
But this will not be possible in general: the facts that $\mathsf{Q}$
is equal to $\mathsf{Pair^{\wedge}\,B\,B\,Q_B\,Q_B}$ and that there is
a morphism of predicates $\mathsf{m}$ from $\mathsf{Q}$ to
$\mathsf{Q'}$ do not guarantee that there exists a predicate
$\mathsf{Q'_B}$ such that $\mathsf{Q'}$ is equal to
$\mathsf{Pair^{\wedge}\,B\,B\,Q'_B\,Q'_B}$.

At a deeper level, the fundamental issue is that the $\mathsf{Equal}$
type does not have functorial semantics~\cite{jgj21}, so that having
morphisms $\mathsf{A \to A'}$ and $\mathsf{B \to B'}$ (for any type
$\mathsf{A, A', B}$ and $\mathsf{B'}$) and a proof that $\mathsf{A}$
is equal to $\mathsf{A'}$ does not provide a proof that $\mathsf{B}$
is equal to $\mathsf{B'}$. And not being able to define
$\mathsf{Equal^\wedge Map :}$ $\mathsf{\forall (A\, B : Set) (Q_A\,
  Q_A' : A \to Set) (Q_B\, Q_B' : B \to Set) \to PredMap\,A\,Q_A\,Q_A'
  \to PredMap\,B\,Q_B\,Q_B' \to
  PredMap\,(Equal\,A\,B)}\\ \mathsf{(Equal^\wedge
  A\,B\,Q_A\,Q_B)\,(Equal^\wedge A\,B\,Q_A'\,Q_B')}$ of course makes
it impossible to define $\mathsf{G^\wedge Map}$ for more general
$\mathsf{G}$.

\section{Case Study: Extracting Types of Lambda Terms}\label{sec:app}

In this section we use deep induction for the $\mathsf{LTerm}$ GADT
from~\eqref{eq:eq_lterm} to extract the type from a lambda term.  The
following predicate either returns the type of its input lambda term
if that type can be inferred or indicates that the type inference
fails:

\vspace*{-0.3in}

\begin{align*}
  &\mathsf{GetType : \forall \, (A : Set) \to LTerm\,A \to Set} \\
  &\mathsf{GetType \,A \,t = Maybe \, (LType \, A)}
\end{align*}
The predicate $\mathsf{GetType}$ uses the standard $\mathsf{Maybe}$
data type to represent failure of type inference. It is defined by:\looseness=-1

\vspace*{-0.1in}

\begin{equation}\label{eq:maybe}
\begin{array}{l}
\mathsf{data\ Maybe : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;nothing :\,  \forall \{A : Set\} \to Maybe\,A}\\
\mathsf{\;\;\;\;\;\;just\;\;\;\;\;\;\; :\,  \forall \{A : Set\} \to A \to Maybe\,A}
\end{array}
\end{equation}

\noindent
We want to show that $\mathsf{GetType\,A\,t}$ is satisfied by every
element $\mathsf{t}$ in $\mathsf{LTerm\,A}$, i.e., we want to prove:
\[ \mathsf{getTypeProof : \forall \, (A : Set)\, (t : LTerm\,A) \to
   GetType \,A \,t}\] This property can be proved with deep
induction, which is used to apply the induction hypothesis to the
individual terms in the list of terms that the data constructor
$\mathsf{list}$ takes as an argument. Indeed, using the deep induction
rule $\mathsf{dIndLTerm}$ from Section~\ref{sec:ind-lam} we can define
$\mathsf{getTypeProof}$ by
\[ \mathsf{getTypeProof \,A \,t = 
    dIndLTerm\, P \, cvar\, cabs\, capp\, clist\, A\, K_\top\, t\,
    (LTerm^\wedge KT\, A\, t) }\] where $\mathsf{t : LTerm\,A}$,
$\mathsf{P}$ is the polymorphic predicate $\mathsf{ \lambda \, (A:
  Set)\, (Q : A \to Set)\, (t : LTerm\,A)\, \to Maybe \, (LType \,
  A)}$, $\mathsf{K_\top}$ is the constantly $\mathsf{\top}$-valued
predicate on $\mathsf{A}$, and $\mathsf{LTerm^\wedge KT : \forall\, (A
  : Set)\, (t : LTerm A) \to LTerm^{\wedge}\, A\, K_\top\,t}$ is a
term, to be defined below, witnessing that $\mathsf{K_\top}$ can be
lifted to all terms. We also need the applications to $\mathsf{P}$ of
each of the induction hypotheses from Section~\ref{sec:ind-lam}. These
are:
\[\begin{array}{lll}
\mathsf{cvar} & : & \mathsf{\forall\, (A : Set)\, (Q_A : A \to Set)\,
  (s : String)\, (T_A : LType\, A) \to LType^{\wedge} \, A\, Q_A\,
  T_A\, \to Maybe\, (LType\, A)}\\
 \mathsf{cabs} & : & \mathsf{\forall\, (A \, B\, C: Set)\, (Q_A : A
   \to Set)\, (Q_B : B \to Set)\, (Q_C : C \to Set)} \\ 
& &    \quad\mathsf{(e : Equal\, A\, (B \to C))\, (s : String)\, (T_B
   : LType\, B)\, (t_C : LTerm \, C)  \to} \\  
& & \quad\mathsf{Equal^{\wedge}\, A\, (B \to C)\, Q_A\,
   (Arr^{\wedge}\, B\, C\, Q_B\, Q_C)\, e  \to LType^{\wedge} \, B\,
   Q_B\, T_B  \to Maybe\, (LType\,C) \to Maybe\, (LType\, A)}\\
\mathsf{capp} & : & \mathsf{\forall\, (A \, B : Set)\, (Q_A : A \to
  Set)\, (Q_B : B \to Set)\,  (t_{BA}: LTerm\, (B \to A))\, (t_B :
  LTerm\, B)\to} \\  
& & \quad\mathsf{ Maybe\, (LType\, (B \to A))\to Maybe\, (LType\,
      B) \to Maybe\, (LType\, A)}\\
\mathsf{clist} & : & \mathsf{\forall\, (A \, B : Set)\, (Q_A : A \to
  Set)\, (Q_B : B \to Set)\, (e : Equal\, A\, (List\, B))\, (ts : 
  List\, (LTerm\, B)) \to}\\  
& & \quad\mathsf{Equal^{\wedge}\, A\, (List\,B) \, Q_A\,
  (List^{\wedge}\, B\, Q_B)\, e \to List^{\wedge}\, (LTerm\, B)\,
  (GetType\, B)\, ts \to Maybe\, (LType\, A)}
\end{array}\]
In the first clause, $\mathsf{cvar}$ returns $\mathsf{just\,T_A}$.  In
the second clause, $\mathsf{cabs}$ returns $\mathsf{nothing}$ if its
final argument is $\mathsf{nothing}$ and $\mathsf{cabs\, A \,B \,C
  \,Q_A\, Q_B \,Q_C\, e \,s \,T_B\, t_C\, liftE\;
%  LLTy^\wedge B\,
lift_{T_B}\; (just \,T_C) = just \, (arr\, B\, C\, e\, T_B \,T_C)}$ otherwise.
In the third clause, $\mathsf{capp\, A \,B \, Q_A \, Q_B \, t_{BA}\,
  t_A \, (just \, (arr\, B \,A\, refl\, T_B\, T_A))\, mb = just\,
  T_A}$ and $\mathsf{capp}$ returns $\mathsf{nothing}$ otherwise.  In
the fourth clause, we must use $\mathsf{List^{\wedge}\, (LTerm\, B)\,
  (GetType\, B)\, ts}$ to extract the type of the head of
$\mathsf{ts}$ (from which we can deduce the type of the list). When
$\mathsf{ts} = \mathsf{nil}$ we define $\mathsf{clist\, A\, B\, Q\,
  Q'\, e\, nil \, liftE\, lift_{ts} = nothing}$, where $\mathsf{liftE
  : Equal^{\wedge}\, A\, (List\,B) \, Q\, (List^{\wedge}\, B\, Q')\,
  e}$, and $\mathsf{lift_{ts} : List^{\wedge}\, (LTerm\, B)\,
  (GetType\, B)\, ts}$.  When $\mathsf{ts} = \mathsf{cons\,t\,ts'}$
% The interesting case is
%  $\mathsf{clist}$, in which we have to use the results of
%  $\mathsf{List^{\wedge}\, (LTerm\, B)\, (GetType\, B)\, ts}$ in order
%  to extract the type of one of the terms in the list.  To define
%  $\mathsf{clist}$ we pattern-match on the list of terms
%  $\mathsf{ts}$.  If $\mathsf{ts}$ is the empty list $\mathsf{nil}$,
%  we cannot extract a type, so we return $\mathsf{nothing}$:
%\[
%  \mathsf{clist\, A\, B\, Q\, Q'\, e\, nil \, liftE\, lift_{ts} = nothing}
%\]
%where $\mathsf{liftE : Equal^{\wedge}\, A\, (List\,B) \, Q\,
%  (List^{\wedge}\, B\, Q')\, e}$ and $\mathsf{lift_{ts} :
%  List^{\wedge}\, (LTerm\, B)\, (GetType\, B)\, ts}$.
%If $\mathsf{ts}$ is a non-empty list $\mathsf{cons\,t\,ts'}$, we
%pattern match on
%$\mathsf{list_{ts}}$ and use the result to construct
%the type we need.  The type of $\mathsf{list_{ts}}$ is
the type of $\mathsf{list_{ts}}$ becomes $\mathsf{List^{\wedge}\,
  (LTerm\, B)\, (GetType\, B)\, (cons\, t\, ts')} = \mathsf{GetType\,
  B\, t \times List^{\wedge}\, (LTerm\, B)\, (GetType\, B)\, ts'} =
\mathsf{Maybe\, (LType\, B) \times List^{\wedge}\, (LTerm\, B)\,
  (GetType\, B)\, ts'}$. We pattern match on the first component
of the pair to define\looseness=-1
%\[\begin{array}{lll}
%\mathsf{list_{ts}} & : &  
%  \mathsf{List^{\wedge}\, (LTerm\, B)\, (GetType\, B)\, (cons\, t\, ts')}\\
%& = & \mathsf{GetType\, B\, t \times List^{\wedge}\, (LTerm\, B)\,
%    (GetType\, B)\, ts'}\\ 
%&= & \mathsf{Maybe\, (LType\, B) \times List^{\wedge}\, (LTerm\, B)\,
%    (GetType\, B)\, ts'} 
%\end{array}\]
\begin{align*}
  &\mathsf{clist\, A\, B\, Q\, Q'\, e\, (cons\, t\, ts') \, liftE\,
    (nothing , lift_{ts'}) = nothing} \\ 
  &\mathsf{clist\, A\, B\, Q\, Q'\, e\, (cons\, t\, ts') \, liftE\,
    (just\, T' , lift_{ts'}) = just \, (list\, B\, e\, T')} 
\end{align*}
Here $\mathsf{e : Equal\, A\, (List\,B)}$, $\mathsf{T' : LType\, B}$,
and $\mathsf{lift_{ts'} : List^{\wedge}\, (LTerm\, B)\, (GetType\,
  B)\, ts'}$.

\pagebreak

To finish defining $\mathsf{getTypeProof}$ we still need a proof

\vspace*{-0.1in}

\[ \mathsf{LTerm^\wedge KT : \forall\, (A : Set)\, (t : LTerm\, A) \to
  LTerm^{\wedge}\, A\, K_\top\,t} \]

\vspace*{-0.05in}

\noindent
Since $\mathsf{LTerm^\wedge}$ is
defined in terms of $\mathsf{LType^\wedge}$ and $\mathsf{Arr^\wedge}$,
and since $\mathsf{LType^\wedge}$ is also defined in terms of
$\mathsf{List^\wedge}$, we need analogous functions
$\mathsf{LType^\wedge KT}$, $\mathsf{Arr^\wedge KT}$ and
$\mathsf{List^\wedge KT}$, respectively, for each of these liftings as
well.  We only give the definition of $\mathsf{LTerm^\wedge KT}$ here
since $\mathsf{LType^\wedge KT}$, $\mathsf{Arr^\wedge KT}$, and
$\mathsf{List^\wedge KT}$ are defined analogously. We have:
\begin{itemize}
\item If $\mathsf{s : String}$ and $\mathsf{T : LType\,A}$ we define
  $\mathsf{LTerm^\wedge KT\,A\,(var\,s\,T)} = \mathsf{LType^\wedge
  KT\,A\,T}$.
\item If $\mathsf{e : Equal\,A\,(B \to C)}$, $\mathsf{s : String}$,
  $\mathsf{T : LType\,B}$, and $\mathsf{t' : LTerm\,C}$ we need to
  define $\mathsf{LTerm^\wedge KT\,A\, (abs\, B \,C \, e \,s \,T \,
    t')}$ of type

\vspace*{-0.15in}

  \[\begin{array}{l}
  \mathsf{LTerm^{\wedge}\,A\,K_\top\, (abs \,B \,C \,e \,s \,T
  \,t')}\\
\quad \mathsf{ = \exists [Q_B]\, [Q_C]\, Equal^{\wedge} \, A\, (B \to
    C)\, K_\top\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e \times \,
    LType^{\wedge}\, B\, Q_B\, T \times \, LTerm^{\wedge}\, C\, Q_C\,
    t' }
\end{array}\]
where $\mathsf{K_\top : A \to Set}$, $\mathsf{Q_B : B \to Set}$, and
$\mathsf{Q_C : C \to Set}$.  The only reasonable choice is to let both
$\mathsf{Q_B}$ and $\mathsf{Q_C}$ be $\mathsf{K_\top}$, which means we
need proofs of $\mathsf{Equal^{\wedge} \, A\, (B \to C)\, K_\top\,
  (Arr^{\wedge} \, B\, C\, K_\top \, K_\top)\, e}$,
$\mathsf{LType^{\wedge}\, B\, K_\top\, T}$ and
$\mathsf{LTerm^{\wedge}\, C\, K_\top\, t'}$.  We take
$\mathsf{LType^\wedge KT\, B\, T}$ and $\mathsf{LTerm^\wedge KT\, C\,
  t'}$ for the latter two proofs. For the former we note that, since
we are working with proof-relevant predicates, the lifting
$\mathsf{Arr^{\wedge} \, B\, C\, K_\top \, K_\top}$ of
$\mathsf{K_\top}$ to arrow types is not identical to $\mathsf{K_\top}$
on arrow types but rather (extensionally) isomorphic.  We discuss this
issue in more detail at the end of the section, but for now we simply
assume a proof $\mathsf{Equal^\wedge ArrKT : Equal^{\wedge} \, A\, (B
  \to C)\, K_\top\, (Arr^{\wedge} \, B\, C\, K_\top \, K_\top)\, e}$
and define $\mathsf{LTerm^\wedge KT\,A\, (abs\, B \,C \, e \,s \,T \,
  t') = (K_\top , K_\top , Equal^\wedge ArrKT , LType^\wedge KT\, B\,
  T , LTerm^\wedge KT\, C\, t') }$.
\item If $\mathsf{t_1 : LTerm\,(B \to A)}$ and $\mathsf{t_2 :
  LTerm\,B}$ then, by the same reasoning as in the previous case, we
  need to define $\mathsf{LTerm^\wedge KT\,A\, (app\, B \, t_1\,t_2)}
  : \mathsf{LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \, B\, A\,
    K_\top \, K_\top)\, t_1 \times LTerm^{\wedge}\, B\, K_\top\,
    t_2}$.  We define the second component of the pair to be
  $\mathsf{LTerm^\wedge KT\,B\,t_2}$. We can define the first
  component from a proof of $\mathsf{LTerm^{\wedge}\, (B \to A)\,
    K_\top\, t_1}$ and the function $\mathsf{LTerm^\wedge EqualMap :
    \forall\, (A : Set)\, (Q\,Q' : A \to Set) \to
    Equal^\wedge\,A\,A\,Q\,Q'\,refl \to}$
  $\mathsf{PredMap\,(LTerm\,A)\,
    (LTerm^{\wedge}\,A\,Q)\,(LTerm^{\wedge}\,A\,Q')}$ that takes two
  (extensionally) equal predicates with the same carrier and produces
  a morphism of predicates between their liftings. The definition of
  $\mathsf{LTerm^\wedge EqualMap}$ is straightforwardly given by
  pattern matching on the first two arguments to $\mathsf{PredMap}$ in
  its return type, using transitivity and symmetry of the type
  constructor $\mathsf{Equal}$, together with the analogously defined
  functions $\mathsf{LType^\wedge EqualMap}$ and $\mathsf{Arr^\wedge
    EqualMap}$ in the cases when the first argument to
  $\mathsf{PredMap}$ is constructed using $\mathsf{var}$ and
  $\mathsf{app}$, respectively. Taking $\mathsf{L_{K_\top} :
    LTerm^{\wedge}\, (B \to A) \, K_\top\, t_1}$ to be the proof
  $\mathsf{L_{K_\top} = LTerm^\wedge KT\, (B \to A)\, t_1}$ and taking
  $\mathsf{ LTerm^\wedge Arr : LTerm^{\wedge}\, (B \to A)\,
    (Arr^{\wedge} \, B\, A\, K_\top \, K_\top)\, t_1}$ to be the proof
  $\mathsf{LTerm^\wedge Arr =}$ $\mathsf{LTerm^\wedge
    EqualMap\,K_\top\,(Arr^{\wedge}\,B\,A\,K_\top\,K_\top)\,\,
    Equal^\wedge ArrKT\, t_1\, L_{K_\top}}$, we define
  $\mathsf{LTerm^\wedge KT\,A\, (app\, B\, t_1\, t_2) = (K_\top ,
    LTerm^\wedge Arr , LTerm^\wedge KT\,B\,t_2)}$.
\item If $\mathsf{e : Equal\,A\,(List\,B)}$ and $\mathsf{ts :
  List\,(LTerm\, B)}$ then, as above, we need to define
  $\mathsf{LTerm^\wedge KT\,A\, (list\, B \,e\, ts)}$ $\mathsf{:
  Equal^{\wedge} \, A\, (List\,B)\, K_\top\, (List^{\wedge} \, B\,
  K_\top) \, e \times List^{\wedge}\, (LTerm\,B) \, (LTerm^{\wedge} \,
  B\, K_\top) \, ts }$.  As in that case we assume a proof
  $\mathsf{Equal^\wedge ListKT : Equal^{\wedge} \, A\, (List\,B)\, K_\top\,
  (List^{\wedge} \, B\, K_\top) \, e}$ for the first component.  We
  can define the second component using $\mathsf{liftListMap}$ from
  Section~\ref{sec:ind-lam} to map a morphism $\mathsf{PredMap\,
    (LTerm\,B)\, (K_\top)\, (LTerm^{\wedge}\,B\,K_\top)}$ of
  predicates to a morphism $\mathsf{PredMap\,(List\,(LTerm\,B))}$
  $\mathsf{(List^{\wedge}\,(LTerm\,B)\,K_\top) \,
    (List^{\wedge}\,(LTerm\,B)\, (LTerm^{\wedge}\,B\,K_\top))}$ of
  lifted predicates.  Taking $\mathsf{m_{K_\top} : PredMap\,
    (LTerm\,B) \, (K_\top)\, (LTerm^{\wedge}\,B\,K_\top)}$ to be the
  proof $\mathsf{m_{K_\top} \, t'\, tt\, = LTerm^\wedge KT\, B\, t'}$,
  where $\mathsf{t' : LTerm\,B}$ and $\mathsf{tt}$ is the single
  element of $\mathsf{K_\top\, t'}$, and taking
  $\mathsf{L_{List^\wedge LTerm^\wedge KT} : List^{\wedge}\,
    (LTerm\,B) \, (LTerm^{\wedge} \, B\, K_\top) \, ts}$ to be the
  proof $\mathsf{ L_{List^\wedge LTerm^\wedge KT} =}$
  $\mathsf{liftListMap \, (LTerm\,B) \, K_\top \, (LTerm^{\wedge}\,
    B\, K_\top) \, m_{K_\top} \, ts \, (List^\wedge KT\, (LTerm\, B)\,
    ts) }$, we define $\mathsf{LTerm^\wedge KT\,A\, (list\, B\, e\,
    ts) = (K_\top , \, Equal^\wedge ListKT , \, L_{List^\wedge
      LTerm^\wedge KT} ) }$.
\end{itemize}

The above techniques can be used to define a function
$\mathsf{G^\wedge KT : \forall\, (A : Set)\, (x : G\,A) \to
  G^{\wedge}\, A\, K_\top\, x}$ for an arbitrary GADT $\mathsf{G}$ as
defined in Section~\ref{sec:framework}.  To provide a proof of
$\mathsf{G^{\wedge}\, A\, K_\top \, x}$ for every term $\mathsf{x :
  G\, A}$, we need to know that, if $\mathsf{G}$ has a constructor
$\mathsf{c : H \, (\ol{F\,G\, \ol{B}}) \to G\, (\ol{K\,\ol{B}})}$,
then $\mathsf{H}$ cannot construct a GADT so the generalization
$\mathsf{H^\wedge Map}$ of $\mathsf{listLiftMap}$ in the final bullet
point above is guaranteed to exist. We also need to know that the
lifting of $\mathsf{K_\top}$ to types constructed by any nested type
constructor $\mathsf{F}$ is extensionally equal to $\mathsf{K_\top}$
on the types it constructs. For example, we might need a proof that
$\mathsf{Pair^{\wedge}\,A\,B\,K_\top\,K_\top}$ is equal to
$\mathsf{K_\top}$ on $\mathsf{A \times B}$.  Given a pair $\mathsf{(a
  , b) : A \times B}$, we have that
$\mathsf{Pair^{\wedge}\,A\,B\,K_\top\,K_\top (a, b) = K_\top \, a
  \times K_\top\, b = \top \times \top}$, whereas $\mathsf{K_\top\,
  (a, b) = \top}$. While these types are not equal, they are clearly
isomorphic. Similar isomorphisms between
$\mathsf{F^{\wedge}\,A\,K_\top}$ and $\mathsf{K_\top}$ hold for all
other nested type constructors $\mathsf{F}$ as well. These
isomorphisms can either be proved on an as-needed basis or, since
$\mathsf{F^\wedge\,A\,K_\top = K_\top}$ is the unary analogue of the
Identity Extension Lemma, be obtained at the meta-level as a
consequence of unary parametricity. At the object level, our Agda
code~\cite{web-page} simply postulates each isomorphism needed since
an Agda implementation of full parametricity for some relevant
calculus is beyond the scope of the present paper.

\section{Conclusion}\label{sec:conclusion}

This paper extends deep induction to deep GADTs that are not truly
nested. It also shows that truly nested GADTs, deep or not, do not
admit (deep) induction rules. Our development is implemented in Agda,
as is a case study showing how deep induction can prove properties of
GADTs that are not provable by structural induction.\looseness=-1

\pagebreak

\begin{thebibliography}{10}\label{bibliography}

\bibitem{atk12} Atkey, R.  {\em Relational parametricity for higher
  kinds}.  Proceedings, Computer Science Logic, pp.~46-61, 2012.

\bibitem{bfss90} Bainbridge, E. S., Freyd, P. Scedrov, A., and Scott,
  P. J. {\em Functorial polymorphism}. Theoretical Computer Science
  70(1), pp. 35-64, 1990.

\bibitem{bm98} Bird, R. and Meertens, L. {\em Nested
  datatypes}. Proceedings, Mathematics of Program Construction,
  pp. 52–67, 1998.

\bibitem{ch03} Cheney, J. and Hinze, R. {\em First-class phantom
 types}. CUCIS TR2003-1901, Cornell University, 2003.

\bibitem{ch88} Coquand, T. and Huet, G. {\em The calculus of
  constructions}. Information and Computation 76(2/3), 1988.

\bibitem{cis194}
Zilberstein, N. CIS194 homepage.
$\mathtt{https://www.seas.upenn.edu/\tilde~cis194/spring15/lectures/11\!\!-\!\!stlc.html}$

%\bibitem{chl} Chlipala, A. {\em Library Inductive
%  Types}. $\mathtt{http://adam.chlipala.net/cpdt/html/InductiveTypes.html}$

\bibitem{coq20} The Coq Development Team. {\em The Coq Proof
  Assistant}, version 8.11.0, January 2020.
  $\mathtt{https://doi.org/10.5281/zenodo.3744225}$

\bibitem{fs18} Fu, P. and Selinger, P.  Dependently typed folds for
  nested data types, 2018. $\mathtt{https://arxiv.org/abs/1806.05230}$

\bibitem{gjfor15} Ghani, N., Johann, P., Nordvall Forsberg, F.,
  Orsanigo, F., and Revell, T. {\em Bifibrational functorial semantics
    for parametric polymorphism}. Proceedings, Mathematical
  Foundations of Program Semantics, pp. 165-181, 2015.

\bibitem{hin03} Hinze, R. {\em Fun with phantom types}. Proceedings,
 The Fun of Programming, pp. 245–262, 2003.

\bibitem{web-page} Johann, P., Ghiorzi, E., and Jeffries,
  D. Accompanying Agda code for this paper.
    $\mathtt{https://cs.appstate.edu/\tilde~johannp/FoSSaCS21Code.html}$

\bibitem{jgj21f} Johann, P., Ghiorzi, E., and Jeffries, D. {\em
  Parametricity for primitive nested types}. Proceedings, Foundations
  of Software Science and Computation Structures, pp. 324-343, 2021.

\bibitem{jgj21} Johann, P., Ghiorzi, E., and Jeffries, D. {\em
  GADTs, Functoriality, Parametricity: Pick Two}. Submitted, 2021.

\bibitem{jp19} Johann, P. and Polonsky, A. {\em Higher-kinded data
  types: Syntax and semantics} Proceedings, Logic in Computer Science,
  pp. 1-13, 2019.

\bibitem{jp20} Johann, P. and Polonsky, A. {\em Deep induction:
  Induction rules for (truly) nested types}.  Proceedings, Foundations
  of Software Science and Computation Structures, pp. 339-358, 2020.

\bibitem{mac71} MacLane, S. {\em Categories for the Working
  Mathematician}. Springer, 1971.

\bibitem{mcb99} McBride, C. {\em Dependently Typed Programs and their
  Proofs}. PhD thesis, University of Edinburgh, 1999.

%\bibitem{min15} Minsky, Y.  {\em Why {GADT}s matter for performance}.
%  $\mathtt{https://blog.janestreet.com/why-gadts-matter-for-performance/}$,
%  2015.

%\bibitem{pl04} Pasalic, E., and Linger, N.  {\em Meta-programming with
%  typed object-language representations}.  Generic Programming and
%  Component Engineering, pp.~136-167, 2004.

%\bibitem{pen20} Penner, C.  {\em Simpler and safer {API} design using
%  {GADT}s}.  $\mathtt{https://chrispenner.ca/posts/gadt-design}$,
%  2020.

\bibitem{pvww06} Peyton Jones, S., Vytiniotis, D., Weirich, S., and
  Washburn, G. {\em Simple unification-based type inference for
    GADTs}. Proceedings, International Conference on Functional
  Programming, pp. 50-61, 2006.

%\bibitem{pr06} Pottier, F., and R{\'e}gis-Gianas, Y.  {\em Stratified
%  type inference for generalized algebraic data types}.  Principles of
%  Programming Languages, pp.~232-244, 2006.

\bibitem{sjsv09} Schrijvers, T., Peyton Jones, S. L., Sulzmann, M., and
  Vytiniotis, D. {\em Complete and decidable type inference for
    GADTs}. Proceedings, International Conference on Functional
  Programming, pp. 341– 352, 2009.

\bibitem{sp04} Sheard, T., and Pasalic, E. {\em Meta-programming with
  built-in type equality}. Proceedings, Workshop on Logical Frameworks
  and Meta-languages, pp. 106-124, 2004.

\bibitem{tas19} Tassi, E.: Deriving proved equality tests in Coq-elpi:
  Stronger induction principles for containers in Coq. Proceedings,
  Interactive Theorem Proving, pp. 1-18, 2019.

\bibitem{vw10} Vytiniotis, D., and Weirich, S.  {\em Parametricity,
  type equality, and higher-order polymorphism}.  Journal of
  Functional Programming 20(2), pp.~175--210, 2010.

\bibitem{xcc03} Xi, H., Chen, C. and Chen, G. {\em Guarded recursive
  datatype constructors}. Proceedings, Principles of Programming
  Languages, pp. 224–235, 2003.

\bibitem{ull20} Ullrich, M. {\em Generating Induction Principles for
Nested Induction Types in MetaCoq}. PhD thesis, Saarland University,
  2020.

\end{thebibliography}

\end{document}